---
title: "Intermediate R Practice"
author: "William Surles"
date: "2017-07-04"
output: 
 html_document:
  self_contained: yes
  theme: flatly
  highlight: tango
---

***
> This document was updated on **`r Sys.Date()`**.

***

```{r setup, include=FALSE}
options(width = 100)
knitr::opts_chunk$set(eval=T, echo=T, cache=T, message=F, warning=F)
```

```{r}
library(dplyr)
```

# Introduction

  - Course notes for [Intermediate R Practice](https://www.datacamp.com/courses/intermediate-r-practice)
  - This course covers the topics from the Intermediate R class again
    - It focuses on application of the skills to answer questions about data
    - It feels a little forced in some cases as I would normally not use loops to analyze data
      - I would get my data out of list format first, as I will show here
    - But it is a really good review of the skills and I feel much more confident in the skills covered after this practice
      - If nothing else this course highlights how some of these skills can be very usefule, especially the regular expressions

## Whats Covered

  - Conditionals and Control Flow
    - Relational Operators. `==` `<` `>` 
    - Logical Operators. `&&` `||` `!`
    - Conditional Statements. `if` `else`
  - While and For loops
  - Functions
  - Apply statements
    - lapply, sapply, vapply
  - Utilities
    - Regular Expressions and grep/sub
    - Times and Dates
    
&nbsp; &nbsp;

***  

# Conditionals and Control Flow

  - I have loaded the data into the notebook manually

```{r, echo=F, results='hide'}

me <- 89

other_199 <- c(
 66,  81,  77,  90,  95,  54,  78,  90,  71,  96,  70,  81,  71,  60,  91,  62,  54,
 70,  96,  90,  82,  78, 100,  79,  84,  73,  66,  63,  97,  91,  82,  81,  52,  73,
 89,  70,  69,  60,  63,  78,  77,  73,  62,  61,  72,  64,  87,  54,  80,  82,  61,
 70,  69,  61,  88,  90,  74,  80,  59,  65,  83,  70,  83,  83,  81,  80,  88,  77,
 84,  50,  73,  70,  74,  75,  71,  60,  80,  78,  81,  60,  80,  99,  90,  90,  66,
 61,  79,  71,  80,  69,  67,  80,  59,  76,  74,  70,  74,  96,  73,  90,  92,  75,
 77,  63,  94,  67,  56,  95,  85,  63,  96,  72,  77,  79,  69,  68,  70,  73,  99,
 64,  59,  63,  82,  76,  90,  80,  86,  80,  84,  80,  99,  80,  68,  77,  51,  67,
 86,  60,  61,  57,  62,  86,  86,  75,  62,  60,  75,  71,  70,  70,  70,  75,  72,
 79,  74,  72,  78,  87,  70,  77,  64,  67,  88,  87,  80,  76,  73,  78,  89,  72,
 86,  68,  84,  64,  73,  73,  64,  70,  91,  65,  69,  99,  76,  94,  72,  77,  80,
 64)

previous_4 <- structure(c(82L, 66L, 66L, 86L, 76L, 74L, 56L, 80L, 71L, 69L, 
69L, 72L, 68L, 63L, 83L, 55L, 87L, 79L, 75L, 100L, 61L, 70L, 
66L, 90L, 85L, 56L, 73L, 72L, 81L, 86L, 60L, 88L, 77L, 80L, 57L, 
69L, 73L, 61L, 50L, 67L, 98L, 80L, 82L, 69L, 89L, 74L, 89L, 73L, 
67L, 67L, 90L, 68L, 64L, 85L, 70L, 78L, 96L, 69L, 74L, 71L, 55L, 
62L, 56L, 83L, 65L, 76L, 83L, 77L, 78L, 52L, 75L, 55L, 50L, 58L, 
68L, 94L, 90L, 67L, 70L, 58L, 80L, 93L, 65L, 73L, 64L, 56L, 73L, 
57L, 65L, 56L, 83L, 72L, 61L, 58L, 56L, 92L, 99L, 68L, 61L, 60L, 
97L, 96L, 51L, 95L, 90L, 89L, 72L, 67L, 50L, 80L, 74L, 72L, 70L, 
88L, 56L, 79L, 98L, 51L, 66L, 79L, 53L, 53L, 72L, 66L, 87L, 61L, 
54L, 70L, 54L, 77L, 60L, 88L, 71L, 90L, 51L, 74L, 74L, 86L, 90L, 
80L, 68L, 75L, 63L, 78L, 65L, 73L, 70L, 69L, 72L, 65L, 88L, 65L, 
67L, 76L, 68L, 88L, 69L, 83L, 76L, 75L, 51L, 66L, 61L, 58L, 73L, 
80L, 96L, 61L, 52L, 74L, 72L, 76L, 78L, 54L, 80L, 79L, 66L, 94L, 
63L, 74L, 67L, 84L, 93L, 70L, 94L, 79L, 60L, 60L, 70L, 60L, 63L, 
68L, 72L, 90L, 96L, 61L, 66L, 70L, 61L, 63L, 89L, 89L, 65L, 79L, 
83L, 73L, 79L, 87L, 59L, 77L, 84L, 75L, 70L, 63L, 74L, 70L, 61L, 
85L, 75L, 72L, 53L, 80L, 80L, 85L, 75L, 77L, 89L, 73L, 72L, 68L, 
74L, 85L, 81L, 76L, 90L, 66L, 87L, 61L, 73L, 60L, 68L, 70L, 84L, 
90L, 68L, 74L, 88L, 93L, 66L, 79L, 83L, 77L, 78L, 66L, 56L, 66L, 
77L, 99L, 91L, 67L, 63L, 87L, 78L, 70L, 68L, 84L, 70L, 65L, 85L, 
70L, 99L, 60L, 100L, 77L, 92L, 82L, 79L, 89L, 50L, 64L, 87L, 
63L, 99L, 73L, 85L, 82L, 62L, 64L, 84L, 77L, 71L, 87L, 58L, 70L, 
87L, 60L, 90L, 68L, 55L, 85L, 96L, 79L, 80L, 71L, 59L, 80L, 70L, 
79L, 75L, 84L, 77L, 70L, 59L, 74L, 73L, 90L, 69L, 80L, 94L, 57L, 
99L, 73L, 80L, 74L, 54L, 77L, 84L, 79L, 53L, 89L, 73L, 83L, 85L, 
79L, 99L, 54L, 90L, 78L, 70L, 79L, 70L, 98L, 60L, 70L, 70L, 92L, 
86L, 59L, 73L, 98L, 72L, 76L, 62L, 82L, 77L, 78L, 84L, 82L, 78L, 
76L, 51L, 55L, 82L, 94L, 62L, 81L, 81L, 73L, 99L, 80L, 77L, 80L, 
79L, 84L, 76L, 84L, 88L, 75L, 52L, 82L, 73L, 76L, 99L, 73L, 70L, 
75L, 67L, 70L, 83L, 88L, 83L, 78L, 57L, 87L, 63L, 60L, 76L, 76L, 
77L, 78L, 53L, 66L, 85L, 65L, 75L, 69L, 98L, 66L, 100L, 76L, 
70L, 89L, 60L, 52L, 80L, 84L, 95L, 68L, 92L, 78L, 55L, 69L, 64L, 
95L, 71L, 55L, 80L, 70L, 63L, 80L, 73L, 93L, 62L, 80L, 90L, 61L, 
64L, 75L, 92L, 100L, 76L, 85L, 97L, 67L, 72L, 60L, 89L, 91L, 
60L, 73L, 87L, 92L, 71L, 72L, 82L, 90L, 63L, 70L, 56L, 75L, 79L, 
62L, 98L, 63L, 88L, 66L, 58L, 73L, 67L, 80L, 69L, 52L, 77L, 60L, 
70L, 76L, 84L, 98L, 62L, 70L, 79L, 90L, 88L, 60L, 67L, 75L, 80L, 
77L, 69L, 72L, 68L, 54L, 90L, 79L, 65L, 71L, 83L, 68L, 60L, 77L, 
79L, 62L, 69L, 52L, 64L, 77L, 89L, 80L, 71L, 60L, 77L, 89L, 77L, 
82L, 90L, 73L, 72L, 50L, 50L, 80L, 61L, 76L, 89L, 91L, 65L, 86L, 
92L, 62L, 87L, 97L, 52L, 76L, 65L, 80L, 71L, 69L, 74L, 98L, 79L, 
75L, 50L, 80L, 66L, 55L, 85L, 75L, 78L, 97L, 58L, 86L, 74L, 70L, 
67L, 68L, 58L, 80L, 73L, 74L, 67L, 54L, 62L, 76L, 66L, 83L, 72L, 
82L, 66L, 66L, 75L, 84L, 81L, 70L, 99L, 84L, 98L, 64L, 68L, 76L, 
74L, 90L, 99L, 69L, 100L, 51L, 81L, 99L, 70L, 62L, 52L, 70L, 
70L, 73L, 80L, 64L, 73L, 98L, 74L, 61L, 70L, 80L, 70L, 76L, 73L, 
64L, 56L, 71L, 96L, 78L, 77L, 99L, 89L, 56L, 92L, 73L, 85L, 71L, 
67L, 75L, 65L, 90L, 87L, 91L, 77L, 81L, 99L, 98L, 59L, 73L, 56L, 
70L, 97L, 50L, 71L, 75L, 95L, 89L, 88L, 93L, 88L, 74L, 61L, 96L, 
97L, 82L, 64L, 85L, 71L, 87L, 51L, 68L, 65L, 79L, 96L, 99L, 94L, 
80L, 71L, 94L, 64L, 59L, 86L, 70L, 55L, 87L, 75L, 83L, 72L, 69L, 
63L, 80L, 85L, 67L, 60L, 70L, 82L, 82L, 90L, 51L, 92L, 86L, 85L, 
62L, 90L, 90L, 68L, 98L, 93L, 80L, 75L, 70L, 66L, 87L, 100L, 
90L, 85L, 76L, 100L, 52L, 75L, 56L, 74L, 91L, 84L, 81L, 89L, 
89L, 80L, 70L, 88L, 77L, 59L, 50L, 56L, 83L, 88L, 77L, 95L, 90L, 
83L, 75L, 78L, 76L, 55L, 73L, 82L, 92L, 90L, 71L, 86L, 66L, 73L, 
86L, 64L, 71L, 80L, 95L, 69L, 82L, 84L, 84L, 53L, 97L, 78L, 76L, 
80L, 61L, 70L, 78L, 70L, 79L, 76L, 86L, 69L, 95L, 55L, 74L, 96L, 
86L, 57L, 79L, 70L, 80L, 73L, 84L, 67L, 89L, 88L, 85L, 64L, 70L, 
77L, 72L, 55L, 77L, 72L, 73L, 86L, 94L, 50L, 71L, 80L, 74L, 94L, 
90L, 84L, 88L, 100L, 66L, 68L, 62L, 71L, 60L, 67L, 88L, 78L, 
87L, 78L, 86L, 76L, 94L), .Dim = c(200L, 4L))
```

```{r}

## my grade in the class (hypothetically)
me 

## the other students grades
glimpse(other_199)

## grades for students over the past 4 years
glimpse(previous_4)
head(previous_4)

```

## Grades analysis in R

  - First, we just join all the data to get a matrix of 5 years of data
  - Normally there would be different numbers each year, but whatev
  
```{r}

# Merge me and other_199: my_class
my_class <- c(me, other_199)
head(my_class)

# cbind() my_class and previous_4: last_5
last_5 <- cbind(my_class, previous_4)
head(last_5)

# Name last_5 appropriately
nms <- paste0("year_", 1:5)

colnames(last_5) <- nms
head(last_5)
```

## Explore your data

  - There is an entire class on this later so this is just the start, 
  - These 3 functions are pretty useful
  
```{r}

# Build histogram of my_class
hist(my_class)

# Generate summary of last_5
summary(last_5)

# Build boxplot of last_5
boxplot(last_5)
```

## Basic queries 
  
  - This uses relational operators to answer very simple questions
  - I am wrapping some of these in head, just to shorten the results

```{r}

# Is your grade equal to 72?
me == 72

# Which grades in your class are higher than 75?
head(my_class > 75, 10)

# Which grades in the last 5 years are below or equal to 64?
head(last_5 <= 64, 10)

```

## Build aggregates

  - Becuse the boolean arrays are only so useful. Normall a count or porportion is more helpful.

```{r}

# How many grades in your class are higher than 75?
sum(my_class > 75)

# How many students in your class scored strictly higher than you?
sum(my_class > me)

# What's the proportion of grades below or equal to 64 in the last 5 years?
mean(last_5 <= 64)

```

## Logical operator

```{r}

# Is your grade greater than 87 and smaller than or equal to 89?
me > 87 & me <= 89

# Which grades in your class are below 60 or above 90?
head(my_class < 60 | my_class > 90, 20)

```

## Build aggregates(2)

```{r}

# What's the proportion of grades in your class that is average?
mean(my_class >= 70 & my_class <= 85)

# How many students in the last 5 years had a grade of 80 or 90?
sum(last_5 == 80 | last_5 == 90)

```

## if, else

```{r}

if (me > 80) {
  print("Good student!")
} else {
  print("Better luck next year!")
}

```

## if, else: DIY

```{r}

# Define n_smart
n_smart <- sum(my_class >= 80)

# Code the if-else construct
if (n_smart > 50) {
  print("smart class")
} else {
  print("rather average")
}

```

## else if

```{r}

# Define prop_less
prop_less <- mean(my_class < me)

# Code the control construct
if(prop_less > .90) {
  print("you're among the best 10 percent")
} else if (prop_less > .80) {
  print("you're among the best 20 percent")
} else {
  print("need more analysis")
}


```

## Embed if-else clauses

```{r}

# Embedded control structure: fix the error
if (mean(my_class) < 75) {
  if (mean(my_class) > me) {
    print("average year, but still smarter than me")
  } else {
    print("average year, but I'm not that bad")
  }
} else {
  if (mean(my_class) > me) {
    print("smart year, even smarter than me")
  } else {
    print("smart year, but I am smarter")
  }
}

```

## Operations and controls expertise

```{r}

# Create top_grades
top_grades <- my_class[my_class >= 85]

# Create worst_grades
worst_grades <- my_class[my_class < 65]

# Write conditional statement
if (length(top_grades) > length(worst_grades)) {
  print("top grades prevail")
} 
```

&nbsp; &nbsp;

***  

# Loops

  - I have loaded the data into the notebook manually

```{r, echo=F, results='hide'}

logs <- list(structure(list(success = TRUE, details = structure(list(
    message = "check"), .Names = "message"), timestamp = structure(1442271667.43207, class = c("POSIXct", 
"POSIXt"))), .Names = c("success", "details", "timestamp")), 
    structure(list(success = TRUE, details = structure(list(message = "all good"), .Names = "message"), 
        timestamp = structure(1442275213.39346, class = c("POSIXct", 
        "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "check"), .Names = "message"), timestamp = structure(1442278843.47229, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "check"), .Names = "message"), timestamp = structure(1442282478.63858, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "ok"), .Names = "message"), timestamp = structure(1442285999.74305, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "all good"), .Names = "message"), timestamp = structure(1442289668.86018, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "check"), .Names = "message"), timestamp = structure(1442293400.3281, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "all good"), .Names = "message"), timestamp = structure(1442296765.63371, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "all good"), .Names = "message"), timestamp = structure(1442300369.37386, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = FALSE, details = structure(list(
        message = "stack overflow", location = "control room"), .Names = c("message", 
    "location")), timestamp = structure(1442304053.61872, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "ok"), .Names = "message"), timestamp = structure(1442307594.93076, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "ok"), .Names = "message"), timestamp = structure(1442311239.45235, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "all good"), .Names = "message"), timestamp = structure(1442314998.8012, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "check"), .Names = "message"), timestamp = structure(1442318509.16918, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "check"), .Names = "message"), timestamp = structure(1442322114.63074, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "ok"), .Names = "message"), timestamp = structure(1442325747.80094, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "check"), .Names = "message"), timestamp = structure(1442329580.85214, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "check"), .Names = "message"), timestamp = structure(1442333148.23046, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "all good"), .Names = "message"), timestamp = structure(1442336830.85628, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = FALSE, details = structure(list(
        message = "segmentation fault", location = "waste"), .Names = c("message", 
    "location")), timestamp = structure(1442340478.48111, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "check"), .Names = "message"), timestamp = structure(1442344104.31929, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "all good"), .Names = "message"), timestamp = structure(1442347808.63364, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = FALSE, details = structure(list(
        message = "human error", location = "reactor"), .Names = c("message", 
    "location")), timestamp = structure(1442351656.34473, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "check"), .Names = "message"), timestamp = structure(1442355229.92537, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "ok"), .Names = "message"), timestamp = structure(1442358897.12962, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "all good"), .Names = "message"), timestamp = structure(1442362564.81979, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "check"), .Names = "message"), timestamp = structure(1442366247.3068, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "ok"), .Names = "message"), timestamp = structure(1442369824.54196, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "ok"), .Names = "message"), timestamp = structure(1442373613.23665, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "all good"), .Names = "message"), timestamp = structure(1442377367.33987, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "ok"), .Names = "message"), timestamp = structure(1442381036.07308, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "ok"), .Names = "message"), timestamp = structure(1442384586.79444, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "ok"), .Names = "message"), timestamp = structure(1442388126.81468, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "ok"), .Names = "message"), timestamp = structure(1442391716.33624, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "check"), .Names = "message"), timestamp = structure(1442395200.71391, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "all good"), .Names = "message"), timestamp = structure(1442398873.1383, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "ok"), .Names = "message"), timestamp = structure(1442402487.24673, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "ok"), .Names = "message"), timestamp = structure(1442406138.39827, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "ok"), .Names = "message"), timestamp = structure(1442409894.00923, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "ok"), .Names = "message"), timestamp = structure(1442413502.7706, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "ok"), .Names = "message"), timestamp = structure(1442416979.75338, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "ok"), .Names = "message"), timestamp = structure(1442420588.43106, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "check"), .Names = "message"), timestamp = structure(1442424320.35423, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "all good"), .Names = "message"), timestamp = structure(1442427945.2025, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "all good"), .Names = "message"), timestamp = structure(1442431539.98714, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "check"), .Names = "message"), timestamp = structure(1442435144.14022, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "check"), .Names = "message"), timestamp = structure(1442438737.23849, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "check"), .Names = "message"), timestamp = structure(1442442496.58945, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "check"), .Names = "message"), timestamp = structure(1442446128.80461, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "check"), .Names = "message"), timestamp = structure(1442449737.47012, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = FALSE, details = structure(list(
        message = "stack overflow", location = "tubes"), .Names = c("message", 
    "location")), timestamp = structure(1442453511.53602, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "ok"), .Names = "message"), timestamp = structure(1442457227.26855, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "ok"), .Names = "message"), timestamp = structure(1442460737.11087, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "all good"), .Names = "message"), timestamp = structure(1442464501.33007, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "check"), .Names = "message"), timestamp = structure(1442468239.02558, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "ok"), .Names = "message"), timestamp = structure(1442471720.55968, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "check"), .Names = "message"), timestamp = structure(1442475370.43848, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "ok"), .Names = "message"), timestamp = structure(1442478869.32623, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "check"), .Names = "message"), timestamp = structure(1442482442.18562, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "ok"), .Names = "message"), timestamp = structure(1442485999.92793, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "ok"), .Names = "message"), timestamp = structure(1442489681.0897, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "ok"), .Names = "message"), timestamp = structure(1442493221.61982, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = FALSE, details = structure(list(
        message = "segmentation fault", location = "waste"), .Names = c("message", 
    "location")), timestamp = structure(1442496823.20539, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "check"), .Names = "message"), timestamp = structure(1442500519.2637, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = FALSE, details = structure(list(
        message = "human error", location = "waste"), .Names = c("message", 
    "location")), timestamp = structure(1442504062.01691, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "ok"), .Names = "message"), timestamp = structure(1442507675.88601, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "ok"), .Names = "message"), timestamp = structure(1442511300.87333, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "all good"), .Names = "message"), timestamp = structure(1442514920.40155, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "ok"), .Names = "message"), timestamp = structure(1442518448.7965, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "all good"), .Names = "message"), timestamp = structure(1442522023.70184, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "check"), .Names = "message"), timestamp = structure(1442525547.24996, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "check"), .Names = "message"), timestamp = structure(1442529154.02717, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "ok"), .Names = "message"), timestamp = structure(1442532866.13324, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "check"), .Names = "message"), timestamp = structure(1442536496.44837, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "all good"), .Names = "message"), timestamp = structure(1442540274.29455, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "ok"), .Names = "message"), timestamp = structure(1442543775.11426, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = FALSE, details = structure(list(
        message = "human error", location = "waste"), .Names = c("message", 
    "location")), timestamp = structure(1442547438.416, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "ok"), .Names = "message"), timestamp = structure(1442550882.44039, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "check"), .Names = "message"), timestamp = structure(1442554480.0724, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "all good"), .Names = "message"), timestamp = structure(1442558130.89049, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "check"), .Names = "message"), timestamp = structure(1442561734.49783, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "check"), .Names = "message"), timestamp = structure(1442565493.2596, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "check"), .Names = "message"), timestamp = structure(1442569141.35461, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "check"), .Names = "message"), timestamp = structure(1442572776.04087, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "check"), .Names = "message"), timestamp = structure(1442576397.55859, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "ok"), .Names = "message"), timestamp = structure(1442580077.17186, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "ok"), .Names = "message"), timestamp = structure(1442583672.2674, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "all good"), .Names = "message"), timestamp = structure(1442587329.75573, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "ok"), .Names = "message"), timestamp = structure(1442591011.3022, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "all good"), .Names = "message"), timestamp = structure(1442594690.06613, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "all good"), .Names = "message"), timestamp = structure(1442598327.47824, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "all good"), .Names = "message"), timestamp = structure(1442602123.66626, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "all good"), .Names = "message"), timestamp = structure(1442605778.01847, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "ok"), .Names = "message"), timestamp = structure(1442609424.24531, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "check"), .Names = "message"), timestamp = structure(1442613054.34025, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )), structure(list(success = TRUE, details = structure(list(
        message = "check"), .Names = "message"), timestamp = structure(1442616469.25693, class = c("POSIXct", 
    "POSIXt"))), .Names = c("success", "details", "timestamp"
    )))

```

  - logs has 96 entries 
  - showing the `str()` of logs will print the str of all entries
  - I shorten here to save space
  
```{r}

length(logs)
logs[1:3]
str(logs[1:3])

```
## Scanning Logs in R

```{r}

# Print the structure of logs
# str(logs)

# Use list subsetting to print the details part of 11th logs entry
logs[11]

# Print the class of the timestamp component of the first entry
class(logs[[1]]$timestamp)

```

## Understand the logs

```{r}
## You can index into a list like this
logs[[2]][[2]][[1]]

## You can also use an array to index
logs[[c(2, 2, 1)]]

## Note that if success == FALSE then there are two elements in the details

```

## While: start easy

```{r}

# Initialize the iterator i to be 1
i <- 1

# Code the while loop
while ( logs[[i]]$success == TRUE ) {
  print(i)
  i <- i + 1
}

```


## Adapt the while loop

```{r}

# Adapt the while loop
i <- 1
while (logs[[i]]$success) {
  print(logs[[i]]$details$message)
  i <- i + 1
}

```

## While: different approach

  - This approach set a value to false initially, then updates when a specific condition is met which will then top the loop

```{r}

# Initialize i and found
i <- 1
found <- FALSE

# Code the while loop
while (found == FALSE) {
  if (logs[[i]]$success == FALSE && logs[[i]]$details$location == "waste") {
    print("found")
    found <- TRUE
  } else {
    print("still looking")
    i <- i + 1
  }
}

```

## The for loop
  
  - I shorten this up to just the first 10 logs
  
```{r}

# Code a for loop that prints the timestamp of each log
for (log in logs[1:10]) {
  print(log$timestamp)
}

```

## Going through the list

```{r}

# Make the printout conditional: only if success
for (log in logs[1:20]) {
  if (log$success == TRUE) {
    print(log$timestamp)  
  }
}

```

## Adapt the logs list

```{r}

# Finish the for loop: add date element for each entry
for (i in 1:length(logs)) {
  logs[[i]]$date <- as.Date(logs[[i]]$timestamp)
}

# Print first 6 elements in logs
head(logs, 3)

```

## Collect all failures

```{r}

# Intialize empty list: failures
failures <- list()

# Finish the for loop: add each failure to failures
for (log in logs) {
  if (log$success == FALSE) {
    failures <- c(failures, list(log))
  }
}

# Display the structure of failures
str(failures)

```

## Without Loops

  - The looping code above is good practice but normally you should not use loops in R. 
  - There are much better ways to work with data. 
  - I will show how to pull this list data into a dataframe to make life much easier while doing analysis. 
  - This is important because you can waste a lot of time not just in execution, but in writing code if you don't learn `dplyr` and other libraries that are specifically made to crunch data. 
  - Loops are good for other programming routines or low level operations but not for crunching data. 
  
### Pulling the list into a clean data frame

  - Ideally the log data would be consistent and not have different elements in different situations.
  - That makes it tricky to clean this up into a dataframe
  - But its still possible and once clean is much easier to work with!
  - Note: I am using the dplyr chain function to make these read well

Convert the variable level list to a dataframe:
  
  - First, I convert each log into a data frame
    - I want to apply the `as.data.frame` function to each log record
    - I can't name the columns here becasue `details.location` is not always present.
    - So, I will clean that up later
  - Then you can use `bind_rows` (from `dplyr` package) to combine the list of dataframes
    - If the columns are missing in some of the dataframes (and they are) that is okay, it will just fill in with NA
  - Now I will clean up the column names so `message` and `details.message` are the same column
    - I just set `details.message` to be `message` when its NA 
    - I wish `as.data.frame` would always pull in the full name path from a list as the column name
    - Unfortunately, If there is just one nested record it will just use the last name as the column name, so we have `message` when its the only thing in details or `details.message` when there are multiple recoreds under details
    - This is lame but its how it works. 
  - And last thing is to remove the original `message` column since we have moved it into `details.message`
    
```{r}

df_logs <- logs %>%
  sapply(as.data.frame) %>%
  bind_rows() %>%
  mutate(
    details.message = ifelse(is.na(details.message), message, details.message)
  ) %>%
  select(-message)

head(df_logs, 12)

```

### Analysing the data

  - Now we can see all the data so much easier. 
  - Really, you always want your data in a dataframe like this. 
  - These few lines of code pretty much cover all the tasks we just did with the loops above. 

```{r, opt.tidy=T}

## subset the data
df_logs %>% filter(success==FALSE)
df_logs %>% filter(success==FALSE & details.location == 'waste')

## add data column
df_logs %>% mutate(date = as.Date(timestamp)) %>% head(10)

```

&nbsp; &nbsp;

***  

# Functions

  - I have loaded the `timestamps` data into the notebook manually
  - This is good practive for writting functions
  - I should note that using dataframes instead of a list makes all of these next functions unnecessary
  - Normall I use functions for more complex data conversions that I want to call from a dplyr chain
    - Or I may use a function to do a data crunching or charting process that I use often
  - I don't use functions to extract data. Just put it in a dataframe and you are good to go. : ) 
  
```{r, echo=F, results='hide',}

timestamps <- structure(c(1442304053.61872, 1442340478.48111, 1442351656.34473, 
1442453511.53602, 1442496823.20539, 1442504062.01691, 1442547438.416
), class = c("POSIXct", "POSIXt"))

```

## Using functions

```{r}

# Call max() on timestamps
max(timestamps)

# What is the date of the latest timestamp?
as.Date(max(timestamps))

```

## Optional Arguments

```{r}

# Print out timestamps
print(timestamps)

# Call max() on timestamps, no additional arguments
max(timestamps)

# Call max() on timestamps, specify na.rm
max(timestamps, na.rm=T)


```

## Extracting log information (1)

```{r}

# Build a function extract_info(): use for loop, add return statement
extract_info <- function(x) {
  info <- c()
  for (log in x) {
    info <- c(info, log$timestamp)
  }
  return(info)
}

# Call extract_info() on logs
extract_info(logs)

```

## Extracting log information (2)

```{r}

# Adapt the extract_info() function.
extract_info <- function(x, property) {
  
  info <- c()
  
  for (log in x) {
   info <- c(info, log[[property]])
  }
  
  return(info)
}

# Call extract_info() on logs, set property to "timestamp"
extract_info(logs, "timestamp")

# Call extract_info() on logs, set property to "success"
extract_info(logs, "success")
```

## Extracting log information (3)

```{r}

# Add default value for property argument
extract_info <- function(x, property="success") {
  info <- c()
  for (log in x) {
   info <- c(info, log[[property]])
  }
  return(info)
}

# Call extract_info() on logs, don't specify property
extract_info(logs)

# Call extract_info() on logs, set property to "timestamp"
extract_info(logs, "timestamp")

```

## Extracting log information (4)

```{r}

# Adapt extract_info():
# - add argument with default value
# - change function body
extract_info <- function(x, property = "success", include_all=T) {
  info <- c()
  for (log in x) {

   # add if construct around the line below
   if (include_all || !log$success) {
      
      info <- c(info, log[[property]])
      
    }

  }
  return(info)
}

# Call extract_info() on logs, no additional arguments
extract_info(logs)

# Call extract_info() on logs, set include_all to FALSE
extract_info(logs, include_all=F)
```

## Extracting log information (5)

```{r}

# Defition of the extract_info() function
extract_info <- function(x, property = "success", include_all = TRUE) {
  info <- c()
  for (log in x) {
    if (include_all || !log$success) {
     info <- c(info, log[[property]])
    }
  }
  return(info)
}

# Generate vector of messages
extract_info(logs, property = c('details', 'message'))

# Generate vector of locations for failed log entries
extract_info(logs, property = c('details','location'), include_all = F)
```

## Over to you

```{r}

# Write the function compute_fail_pct

compute_fail_pct <- function(logs) {
  
  is_fail <- c()
  
  for (i in 1:length(logs)) {
    is_fail[i] <- logs[[i]]$success == FALSE
  }
  
  fail_pct <- sum(is_fail)/length(is_fail) * 100
  
  return(fail_pct)
}


# Call compute_fail_pct on logs
compute_fail_pct(logs)

```

&nbsp; &nbsp;

***  

# The apply family

  - I have rarely use the apply family after learning dplyr
  - But there are some times when its helpful, like above where I used it to convert the list elements to a dataframe before row binding them
  - These are also good to know because you may need to do some data crunching in another language and this would be a good pattern to follow if they have a `map` function or something like the `apply` functions. 

## lapply refresher
  
```{r}

# Call length() on each element of logs
lapply(logs,length) %>% head()

# Call class() on each element of logs
lapply(logs,class) %>% head()

```

## lapply on logs (1)

```{r}

# Define get_timestamp()
get_timestamp <- function(x) {
  return(x$timestamp)
}

# Apply get_timestamp() over all elements in logs
lapply(logs, get_timestamp) %>% head()

```

## lapply on logs (2)

```{r}

# Have lapply() use an anonymous function
lapply(logs, function(x) {x$timestamp}) %>% head()

```

## lapply on logs (3)

```{r}

# Replace the anonymous function with `[[` 
lapply(logs, `[[`, 'timestamp') %>% head()

```

## sapply refresher

```{r}

# Call length() on each element of logs using sapply()
sapply(logs, length)

# Definition of get_timestamp
get_timestamp <- function(x) {
  x$timestamp
}

# Get vector of log entries' timestamps
sapply(logs, get_timestamp)

```

## sapply on logs (1)

```{r}

# Use sapply() to select the success element from each log: results
results <- sapply(logs, `[[`, 'success')

# Call mean() on results
mean(results)

# Use sapply() to select the details element from each log
sapply(logs, `[[`, 'details') %>% head()

```

## sapply on logs (2)

```{r}

# Implement function get_failure_loc
get_failure_loc <- function(x) {
  if (x$success==T) {
    return(NULL)
  } else {
    return(x$details$location)
  }
}

# Use sapply() to call get_failure_loc on logs
sapply(logs, get_failure_loc) %>% head(11)

```

## vapply refresher

```{r}

# Convert the sapply call to vapply
vapply(logs, length, integer(1))

# Convert the sapply call to vapply
vapply(logs, `[[`, "success", FUN.VALUE = logical(1))

```

## vapply on logs (1)

```{r}

# Convert the sapply() call to a vapply() or lapply() call
vapply(logs, `[[`, c("details", "message"), FUN.VALUE=character(1))

# Convert the sapply() call to a vapply() or lapply() call
lapply(logs, function(x) { x$details }) %>% head()

```

## loop it the way you want it

```{r}

# Return vector with uppercase version of message elements in log entries
toupper(sapply(logs, `[[`, c('details','message')))

```

&nbsp; &nbsp;

***  

# Utilities

  - I have loaded the data into the notebook manually
  
```{r, echo=F, results='hide'}

titanic <- structure(list(PassengerId = 1:891, Survived = c(0L, 1L, 1L, 
1L, 0L, 0L, 0L, 0L, 1L, 1L, 1L, 1L, 0L, 0L, 0L, 1L, 0L, 1L, 0L, 
1L, 0L, 1L, 1L, 1L, 0L, 1L, 0L, 0L, 1L, 0L, 0L, 1L, 1L, 0L, 0L, 
0L, 1L, 0L, 0L, 1L, 0L, 0L, 0L, 1L, 1L, 0L, 0L, 1L, 0L, 0L, 0L, 
0L, 1L, 1L, 0L, 1L, 1L, 0L, 1L, 0L, 0L, 1L, 0L, 0L, 0L, 1L, 1L, 
0L, 1L, 0L, 0L, 0L, 0L, 0L, 1L, 0L, 0L, 0L, 1L, 1L, 0L, 1L, 1L, 
0L, 1L, 1L, 0L, 0L, 1L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 1L, 1L, 
0L, 0L, 0L, 0L, 0L, 0L, 0L, 1L, 1L, 0L, 1L, 0L, 0L, 0L, 0L, 0L, 
0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 1L, 0L, 1L, 0L, 1L, 1L, 0L, 0L, 
0L, 0L, 1L, 0L, 0L, 1L, 0L, 0L, 0L, 0L, 1L, 1L, 0L, 0L, 0L, 1L, 
0L, 0L, 0L, 0L, 1L, 0L, 0L, 0L, 0L, 1L, 0L, 0L, 0L, 0L, 1L, 0L, 
0L, 0L, 1L, 1L, 0L, 0L, 0L, 0L, 0L, 1L, 0L, 0L, 0L, 0L, 0L, 0L, 
0L, 0L, 0L, 0L, 1L, 1L, 0L, 1L, 1L, 0L, 0L, 1L, 0L, 1L, 1L, 1L, 
1L, 0L, 0L, 1L, 0L, 0L, 0L, 0L, 0L, 1L, 0L, 0L, 1L, 1L, 1L, 0L, 
1L, 0L, 0L, 0L, 1L, 1L, 0L, 1L, 0L, 1L, 0L, 0L, 0L, 1L, 0L, 1L, 
0L, 0L, 0L, 1L, 0L, 0L, 1L, 0L, 0L, 0L, 1L, 0L, 0L, 0L, 1L, 0L, 
0L, 0L, 0L, 0L, 1L, 1L, 0L, 0L, 0L, 0L, 0L, 0L, 1L, 1L, 1L, 1L, 
1L, 0L, 1L, 0L, 0L, 0L, 0L, 0L, 1L, 1L, 1L, 0L, 1L, 1L, 0L, 1L, 
1L, 0L, 0L, 0L, 1L, 0L, 0L, 0L, 1L, 0L, 0L, 1L, 0L, 1L, 1L, 1L, 
1L, 0L, 0L, 0L, 0L, 0L, 0L, 1L, 1L, 1L, 1L, 0L, 1L, 0L, 1L, 1L, 
1L, 0L, 1L, 1L, 1L, 0L, 0L, 0L, 1L, 1L, 0L, 1L, 1L, 0L, 0L, 1L, 
1L, 0L, 1L, 0L, 1L, 1L, 1L, 1L, 0L, 0L, 0L, 1L, 0L, 0L, 1L, 1L, 
0L, 1L, 1L, 0L, 0L, 0L, 1L, 1L, 1L, 1L, 0L, 0L, 0L, 0L, 0L, 0L, 
0L, 1L, 0L, 1L, 1L, 0L, 0L, 0L, 0L, 0L, 0L, 1L, 1L, 1L, 1L, 1L, 
0L, 0L, 0L, 0L, 1L, 1L, 0L, 0L, 0L, 1L, 1L, 0L, 1L, 0L, 0L, 0L, 
1L, 0L, 1L, 1L, 1L, 0L, 1L, 1L, 0L, 0L, 0L, 0L, 1L, 1L, 0L, 0L, 
0L, 0L, 0L, 0L, 1L, 0L, 0L, 0L, 0L, 1L, 0L, 1L, 0L, 1L, 1L, 0L, 
0L, 0L, 0L, 0L, 0L, 0L, 0L, 1L, 1L, 0L, 1L, 1L, 1L, 1L, 0L, 0L, 
1L, 0L, 1L, 0L, 0L, 1L, 0L, 0L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 0L, 
0L, 0L, 1L, 0L, 1L, 0L, 1L, 1L, 0L, 1L, 0L, 0L, 0L, 0L, 0L, 0L, 
0L, 0L, 1L, 0L, 0L, 1L, 1L, 0L, 0L, 0L, 0L, 0L, 1L, 0L, 0L, 0L, 
1L, 1L, 0L, 1L, 0L, 0L, 1L, 0L, 0L, 0L, 0L, 0L, 0L, 1L, 0L, 0L, 
0L, 0L, 0L, 0L, 0L, 1L, 0L, 1L, 1L, 0L, 1L, 1L, 0L, 1L, 1L, 0L, 
0L, 1L, 0L, 1L, 0L, 1L, 0L, 0L, 1L, 0L, 0L, 1L, 0L, 0L, 0L, 1L, 
0L, 0L, 1L, 0L, 1L, 0L, 1L, 0L, 1L, 1L, 0L, 0L, 1L, 0L, 0L, 1L, 
1L, 0L, 1L, 1L, 0L, 0L, 1L, 1L, 0L, 1L, 0L, 1L, 1L, 0L, 0L, 0L, 
0L, 0L, 0L, 0L, 0L, 0L, 1L, 1L, 1L, 1L, 1L, 0L, 0L, 1L, 1L, 0L, 
1L, 1L, 1L, 0L, 0L, 0L, 1L, 0L, 1L, 0L, 0L, 0L, 1L, 0L, 0L, 0L, 
0L, 1L, 0L, 0L, 1L, 1L, 0L, 0L, 0L, 1L, 0L, 0L, 1L, 1L, 1L, 0L, 
0L, 1L, 0L, 0L, 1L, 0L, 0L, 1L, 0L, 0L, 1L, 1L, 0L, 0L, 0L, 0L, 
1L, 0L, 0L, 1L, 0L, 1L, 0L, 0L, 1L, 0L, 0L, 0L, 0L, 0L, 1L, 0L, 
1L, 1L, 1L, 0L, 1L, 0L, 1L, 0L, 1L, 0L, 1L, 0L, 0L, 0L, 0L, 0L, 
0L, 1L, 0L, 0L, 0L, 1L, 0L, 0L, 0L, 0L, 1L, 1L, 0L, 0L, 1L, 0L, 
0L, 0L, 1L, 0L, 1L, 0L, 1L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 1L, 1L, 
1L, 1L, 0L, 0L, 0L, 0L, 1L, 0L, 0L, 1L, 1L, 0L, 0L, 0L, 0L, 1L, 
1L, 1L, 1L, 1L, 0L, 1L, 0L, 0L, 0L, 1L, 1L, 0L, 0L, 1L, 0L, 0L, 
0L, 1L, 0L, 1L, 1L, 0L, 0L, 1L, 0L, 0L, 0L, 0L, 0L, 0L, 1L, 0L, 
0L, 1L, 0L, 1L, 0L, 1L, 0L, 0L, 1L, 0L, 0L, 1L, 1L, 0L, 0L, 1L, 
1L, 0L, 0L, 0L, 1L, 0L, 0L, 1L, 1L, 0L, 1L, 0L, 0L, 0L, 0L, 0L, 
0L, 0L, 0L, 1L, 0L, 0L, 1L, 0L, 1L, 1L, 1L, 0L, 0L, 0L, 0L, 1L, 
0L, 1L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 1L, 1L, 0L, 0L, 0L, 1L, 1L, 
1L, 1L, 0L, 0L, 0L, 0L, 1L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 
0L, 1L, 1L, 0L, 1L, 0L, 0L, 0L, 1L, 1L, 1L, 1L, 1L, 0L, 0L, 0L, 
1L, 0L, 0L, 1L, 1L, 0L, 0L, 1L, 0L, 0L, 0L, 0L, 0L, 0L, 1L, 0L, 
0L, 0L, 1L, 0L, 1L, 1L, 1L, 1L, 0L, 0L, 0L, 1L, 0L, 0L, 1L, 1L, 
0L, 0L, 1L, 0L, 1L, 0L, 0L, 1L, 1L, 0L, 0L, 0L, 1L, 1L, 0L, 0L, 
0L, 0L, 0L, 0L, 1L, 0L, 1L, 0L), Pclass = c(3L, 1L, 3L, 1L, 3L, 
3L, 1L, 3L, 3L, 2L, 3L, 1L, 3L, 3L, 3L, 2L, 3L, 2L, 3L, 3L, 2L, 
2L, 3L, 1L, 3L, 3L, 3L, 1L, 3L, 3L, 1L, 1L, 3L, 2L, 1L, 1L, 3L, 
3L, 3L, 3L, 3L, 2L, 3L, 2L, 3L, 3L, 3L, 3L, 3L, 3L, 3L, 3L, 1L, 
2L, 1L, 1L, 2L, 3L, 2L, 3L, 3L, 1L, 1L, 3L, 1L, 3L, 2L, 3L, 3L, 
3L, 2L, 3L, 2L, 3L, 3L, 3L, 3L, 3L, 2L, 3L, 3L, 3L, 3L, 1L, 2L, 
3L, 3L, 3L, 1L, 3L, 3L, 3L, 1L, 3L, 3L, 3L, 1L, 1L, 2L, 2L, 3L, 
3L, 1L, 3L, 3L, 3L, 3L, 3L, 3L, 3L, 1L, 3L, 3L, 3L, 3L, 3L, 3L, 
2L, 1L, 3L, 2L, 3L, 2L, 2L, 1L, 3L, 3L, 3L, 3L, 3L, 3L, 3L, 3L, 
2L, 2L, 2L, 1L, 1L, 3L, 1L, 3L, 3L, 3L, 3L, 2L, 2L, 3L, 3L, 2L, 
2L, 2L, 1L, 3L, 3L, 3L, 1L, 3L, 3L, 3L, 3L, 3L, 2L, 3L, 3L, 3L, 
3L, 1L, 3L, 1L, 3L, 1L, 3L, 3L, 3L, 1L, 3L, 3L, 1L, 2L, 3L, 3L, 
2L, 3L, 2L, 3L, 1L, 3L, 1L, 3L, 3L, 2L, 2L, 3L, 2L, 1L, 1L, 3L, 
3L, 3L, 2L, 3L, 3L, 3L, 3L, 3L, 3L, 3L, 3L, 3L, 1L, 3L, 2L, 3L, 
2L, 3L, 1L, 3L, 2L, 1L, 2L, 3L, 2L, 3L, 3L, 1L, 3L, 2L, 3L, 2L, 
3L, 1L, 3L, 2L, 3L, 2L, 3L, 2L, 2L, 2L, 2L, 3L, 3L, 2L, 3L, 3L, 
1L, 3L, 2L, 1L, 2L, 3L, 3L, 1L, 3L, 3L, 3L, 1L, 1L, 1L, 2L, 3L, 
3L, 1L, 1L, 3L, 2L, 3L, 3L, 1L, 1L, 1L, 3L, 2L, 1L, 3L, 1L, 3L, 
2L, 3L, 3L, 3L, 3L, 3L, 3L, 1L, 3L, 3L, 3L, 2L, 3L, 1L, 1L, 2L, 
3L, 3L, 1L, 3L, 1L, 1L, 1L, 3L, 3L, 3L, 2L, 3L, 1L, 1L, 1L, 2L, 
1L, 1L, 1L, 2L, 3L, 2L, 3L, 2L, 2L, 1L, 1L, 3L, 3L, 2L, 2L, 3L, 
1L, 3L, 2L, 3L, 1L, 3L, 1L, 1L, 3L, 1L, 3L, 1L, 1L, 3L, 1L, 2L, 
1L, 2L, 2L, 2L, 2L, 2L, 3L, 3L, 3L, 3L, 1L, 3L, 3L, 3L, 3L, 1L, 
2L, 3L, 3L, 3L, 2L, 3L, 3L, 3L, 3L, 1L, 3L, 3L, 1L, 1L, 3L, 3L, 
1L, 3L, 1L, 3L, 1L, 3L, 3L, 1L, 3L, 3L, 1L, 3L, 2L, 3L, 2L, 3L, 
2L, 1L, 3L, 3L, 1L, 3L, 3L, 3L, 2L, 2L, 2L, 3L, 3L, 3L, 3L, 3L, 
2L, 3L, 2L, 3L, 3L, 3L, 3L, 1L, 2L, 3L, 3L, 2L, 2L, 2L, 3L, 3L, 
3L, 3L, 3L, 3L, 3L, 2L, 2L, 3L, 3L, 1L, 3L, 2L, 3L, 1L, 1L, 3L, 
2L, 1L, 2L, 2L, 3L, 3L, 2L, 3L, 1L, 2L, 1L, 3L, 1L, 2L, 3L, 1L, 
1L, 3L, 3L, 1L, 1L, 2L, 3L, 1L, 3L, 1L, 2L, 3L, 3L, 2L, 1L, 3L, 
3L, 3L, 3L, 2L, 2L, 3L, 1L, 2L, 3L, 3L, 3L, 3L, 2L, 3L, 3L, 1L, 
3L, 1L, 1L, 3L, 3L, 3L, 3L, 1L, 1L, 3L, 3L, 1L, 3L, 1L, 3L, 3L, 
3L, 3L, 3L, 1L, 1L, 2L, 1L, 3L, 3L, 3L, 3L, 1L, 1L, 3L, 1L, 2L, 
3L, 2L, 3L, 1L, 3L, 3L, 1L, 3L, 3L, 2L, 1L, 3L, 2L, 2L, 3L, 3L, 
3L, 3L, 2L, 1L, 1L, 3L, 1L, 1L, 3L, 3L, 2L, 1L, 1L, 2L, 2L, 3L, 
2L, 1L, 2L, 3L, 3L, 3L, 1L, 1L, 1L, 1L, 3L, 3L, 3L, 2L, 3L, 3L, 
3L, 3L, 3L, 3L, 3L, 2L, 1L, 1L, 3L, 3L, 3L, 2L, 1L, 3L, 3L, 2L, 
1L, 2L, 1L, 3L, 1L, 2L, 1L, 3L, 3L, 3L, 1L, 3L, 3L, 2L, 3L, 2L, 
3L, 3L, 1L, 2L, 3L, 1L, 3L, 1L, 3L, 3L, 1L, 2L, 1L, 3L, 3L, 3L, 
3L, 3L, 2L, 3L, 3L, 2L, 2L, 3L, 1L, 3L, 3L, 3L, 1L, 2L, 1L, 3L, 
3L, 1L, 3L, 1L, 1L, 3L, 2L, 3L, 2L, 3L, 3L, 3L, 1L, 3L, 3L, 3L, 
1L, 3L, 1L, 3L, 3L, 3L, 2L, 3L, 3L, 3L, 2L, 3L, 3L, 2L, 1L, 1L, 
3L, 1L, 3L, 3L, 2L, 2L, 3L, 3L, 1L, 2L, 1L, 2L, 2L, 2L, 3L, 3L, 
3L, 3L, 1L, 3L, 1L, 3L, 3L, 2L, 2L, 3L, 3L, 3L, 1L, 1L, 3L, 3L, 
3L, 1L, 2L, 3L, 3L, 1L, 3L, 1L, 1L, 3L, 3L, 3L, 2L, 2L, 1L, 1L, 
3L, 1L, 1L, 1L, 3L, 2L, 3L, 1L, 2L, 3L, 3L, 2L, 3L, 2L, 2L, 1L, 
3L, 2L, 3L, 2L, 3L, 1L, 3L, 2L, 2L, 2L, 3L, 3L, 1L, 3L, 3L, 1L, 
1L, 1L, 3L, 3L, 1L, 3L, 2L, 1L, 3L, 2L, 3L, 3L, 3L, 2L, 2L, 3L, 
2L, 3L, 1L, 3L, 3L, 3L, 1L, 3L, 1L, 1L, 3L, 3L, 3L, 3L, 3L, 2L, 
3L, 2L, 3L, 3L, 3L, 3L, 1L, 3L, 1L, 1L, 3L, 3L, 3L, 3L, 3L, 3L, 
1L, 3L, 2L, 3L, 1L, 3L, 2L, 1L, 3L, 3L, 3L, 2L, 2L, 1L, 3L, 3L, 
3L, 1L, 3L, 2L, 1L, 3L, 3L, 2L, 3L, 3L, 1L, 3L, 2L, 3L, 3L, 1L, 
3L, 1L, 3L, 3L, 3L, 3L, 2L, 3L, 1L, 3L, 2L, 3L, 3L, 3L, 1L, 3L, 
3L, 3L, 1L, 3L, 2L, 1L, 3L, 3L, 3L, 3L, 3L, 2L, 1L, 3L, 3L, 3L, 
1L, 2L, 3L, 1L, 1L, 3L, 3L, 3L, 2L, 1L, 3L, 2L, 2L, 2L, 1L, 3L, 
3L, 3L, 1L, 1L, 3L, 2L, 3L, 3L, 3L, 3L, 1L, 2L, 3L, 3L, 2L, 3L, 
3L, 2L, 1L, 3L, 1L, 3L), Name = structure(c(109L, 191L, 354L, 
273L, 16L, 555L, 516L, 625L, 413L, 577L, 728L, 96L, 730L, 29L, 
841L, 360L, 683L, 868L, 840L, 513L, 274L, 81L, 524L, 766L, 627L, 
45L, 241L, 261L, 606L, 814L, 829L, 777L, 290L, 857L, 536L, 373L, 
506L, 135L, 838L, 590L, 8L, 828L, 447L, 457L, 213L, 698L, 474L, 
605L, 727L, 39L, 630L, 595L, 334L, 247L, 621L, 874L, 708L, 596L, 
854L, 300L, 750L, 386L, 338L, 756L, 785L, 561L, 597L, 186L, 26L, 
441L, 397L, 301L, 377L, 162L, 88L, 547L, 779L, 564L, 128L, 224L, 
843L, 738L, 520L, 142L, 387L, 53L, 256L, 765L, 260L, 150L, 160L, 
32L, 151L, 210L, 185L, 741L, 293L, 309L, 220L, 423L, 657L, 659L, 
859L, 406L, 313L, 543L, 725L, 560L, 676L, 553L, 668L, 883L, 68L, 
418L, 47L, 646L, 179L, 827L, 71L, 25L, 363L, 552L, 576L, 849L, 
858L, 589L, 526L, 500L, 654L, 235L, 226L, 166L, 696L, 852L, 773L, 
686L, 586L, 272L, 619L, 283L, 100L, 598L, 320L, 117L, 33L, 587L, 
30L, 255L, 581L, 122L, 69L, 644L, 533L, 890L, 613L, 867L, 288L, 
182L, 768L, 717L, 187L, 848L, 83L, 130L, 629L, 294L, 158L, 761L, 
70L, 487L, 833L, 681L, 409L, 753L, 770L, 445L, 465L, 389L, 321L, 
475L, 718L, 651L, 42L, 77L, 442L, 701L, 602L, 699L, 102L, 824L, 
665L, 137L, 21L, 580L, 114L, 496L, 534L, 612L, 498L, 882L, 835L, 
722L, 404L, 879L, 167L, 790L, 52L, 10L, 141L, 95L, 13L, 132L, 
650L, 289L, 438L, 583L, 376L, 391L, 73L, 339L, 792L, 106L, 307L, 
582L, 380L, 84L, 532L, 493L, 244L, 468L, 340L, 461L, 755L, 44L, 
480L, 330L, 371L, 173L, 649L, 384L, 884L, 568L, 170L, 501L, 48L, 
540L, 482L, 323L, 79L, 148L, 674L, 791L, 783L, 488L, 703L, 820L, 
809L, 157L, 845L, 637L, 772L, 43L, 800L, 341L, 357L, 675L, 632L, 
653L, 306L, 92L, 126L, 818L, 531L, 578L, 351L, 34L, 483L, 635L, 
682L, 3L, 228L, 615L, 888L, 222L, 771L, 780L, 887L, 571L, 379L, 
178L, 66L, 91L, 478L, 316L, 542L, 479L, 325L, 18L, 715L, 72L, 
432L, 519L, 412L, 429L, 869L, 17L, 249L, 648L, 4L, 264L, 349L, 
711L, 450L, 356L, 343L, 591L, 424L, 556L, 860L, 776L, 212L, 201L, 
763L, 129L, 723L, 878L, 599L, 61L, 296L, 364L, 518L, 638L, 305L, 
839L, 266L, 211L, 643L, 118L, 193L, 94L, 579L, 259L, 171L, 734L, 
263L, 112L, 769L, 208L, 184L, 216L, 609L, 871L, 238L, 38L, 880L, 
836L, 105L, 271L, 523L, 546L, 760L, 889L, 65L, 41L, 600L, 6L, 
846L, 563L, 401L, 49L, 329L, 864L, 76L, 691L, 626L, 537L, 455L, 
863L, 86L, 315L, 87L, 573L, 812L, 374L, 666L, 206L, 299L, 119L, 
716L, 469L, 145L, 394L, 314L, 584L, 729L, 405L, 614L, 525L, 624L, 
822L, 593L, 7L, 419L, 319L, 617L, 275L, 862L, 688L, 89L, 466L, 
733L, 344L, 541L, 192L, 793L, 529L, 227L, 743L, 514L, 830L, 282L, 
156L, 885L, 199L, 702L, 873L, 163L, 663L, 250L, 664L, 93L, 811L, 
492L, 421L, 745L, 144L, 254L, 689L, 262L, 448L, 345L, 324L, 660L, 
680L, 403L, 217L, 530L, 736L, 56L, 662L, 855L, 317L, 258L, 291L, 
645L, 393L, 538L, 437L, 816L, 604L, 22L, 558L, 281L, 539L, 503L, 
297L, 133L, 767L, 732L, 55L, 431L, 125L, 856L, 402L, 788L, 165L, 
677L, 108L, 426L, 366L, 298L, 269L, 707L, 826L, 90L, 467L, 382L, 
436L, 775L, 183L, 318L, 872L, 548L, 40L, 782L, 881L, 243L, 739L, 
19L, 797L, 131L, 134L, 608L, 451L, 502L, 647L, 671L, 107L, 611L, 
456L, 196L, 850L, 522L, 706L, 169L, 844L, 473L, 710L, 36L, 642L, 
652L, 842L, 449L, 365L, 428L, 245L, 690L, 246L, 726L, 367L, 670L, 
819L, 237L, 655L, 124L, 342L, 121L, 462L, 693L, 267L, 189L, 27L, 
28L, 74L, 223L, 588L, 75L, 623L, 295L, 204L, 805L, 737L, 601L, 
464L, 610L, 875L, 229L, 694L, 801L, 886L, 559L, 752L, 594L, 747L, 
528L, 205L, 787L, 628L, 218L, 417L, 337L, 37L, 252L, 433L, 709L, 
640L, 278L, 746L, 136L, 420L, 161L, 806L, 225L, 704L, 641L, 799L, 
396L, 268L, 285L, 567L, 692L, 784L, 239L, 101L, 155L, 831L, 471L, 
410L, 99L, 230L, 392L, 762L, 336L, 817L, 375L, 484L, 425L, 200L, 
459L, 742L, 31L, 395L, 569L, 378L, 111L, 358L, 198L, 489L, 78L, 
280L, 876L, 440L, 574L, 328L, 104L, 795L, 443L, 491L, 97L, 603L, 
67L, 495L, 778L, 636L, 758L, 207L, 470L, 174L, 634L, 810L, 398L, 
724L, 759L, 253L, 54L, 333L, 181L, 748L, 866L, 781L, 545L, 219L, 
422L, 607L, 353L, 361L, 672L, 103L, 233L, 585L, 265L, 58L, 172L, 
168L, 485L, 362L, 120L, 700L, 180L, 803L, 115L, 203L, 544L, 865L, 
847L, 232L, 731L, 825L, 303L, 138L, 656L, 346L, 616L, 302L, 113L, 
458L, 633L, 195L, 248L, 499L, 214L, 427L, 453L, 714L, 851L, 154L, 
434L, 566L, 804L, 383L, 46L, 744L, 64L, 276L, 327L, 557L, 435L, 
127L, 164L, 562L, 515L, 444L, 802L, 460L, 308L, 774L, 242L, 823L, 
521L, 411L, 332L, 400L, 287L, 369L, 152L, 618L, 678L, 508L, 116L, 
388L, 15L, 347L, 446L, 85L, 821L, 870L, 257L, 477L, 390L, 575L, 
348L, 149L, 712L, 527L, 789L, 188L, 2L, 749L, 512L, 177L, 853L, 
550L, 834L, 416L, 359L, 322L, 139L, 59L, 807L, 705L, 277L, 592L, 
63L, 147L, 234L, 370L, 110L, 507L, 554L, 310L, 481L, 399L, 497L, 
236L, 368L, 570L, 813L, 240L, 439L, 695L, 51L, 215L, 490L, 415L, 
14L, 331L, 754L, 684L, 209L, 311L, 430L, 279L, 720L, 381L, 202L, 
622L, 463L, 620L, 385L, 832L, 667L, 175L, 143L, 808L, 352L, 407L, 
35L, 661L, 535L, 153L, 11L, 476L, 764L, 24L, 815L, 270L, 355L, 
505L, 372L, 757L, 350L, 494L, 679L, 551L, 631L, 251L, 454L, 504L, 
517L, 786L, 877L, 687L, 713L, 50L, 20L, 176L, 639L, 751L, 159L, 
509L, 12L, 565L, 735L, 472L, 190L, 1L, 721L, 510L, 335L, 292L, 
23L, 796L, 98L, 486L, 146L, 9L, 861L, 197L, 57L, 673L, 326L, 
284L, 798L, 719L, 286L, 123L, 231L, 697L, 891L, 408L, 60L, 80L, 
140L, 837L, 5L, 572L, 312L, 658L, 452L, 669L, 740L, 511L, 194L, 
62L, 794L, 685L, 549L, 304L, 414L, 82L, 221L), .Label = c("Abbing, Mr. Anthony", 
"Abbott, Mr. Rossmore Edward", "Abbott, Mrs. Stanton (Rosa Hunt)", 
"Abelson, Mr. Samuel", "Abelson, Mrs. Samuel (Hannah Wizosky)", 
"Adahl, Mr. Mauritz Nils Martin", "Adams, Mr. John", "Ahlin, Mrs. Johan (Johanna Persdotter Larsson)", 
"Aks, Mrs. Sam (Leah Rosen)", "Albimona, Mr. Nassef Cassem", 
"Alexander, Mr. William", "Alhomaki, Mr. Ilmari Rudolf", "Ali, Mr. Ahmed", 
"Ali, Mr. William", "Allen, Miss. Elisabeth Walton", "Allen, Mr. William Henry", 
"Allison, Master. Hudson Trevor", "Allison, Miss. Helen Loraine", 
"Allison, Mrs. Hudson J C (Bessie Waldo Daniels)", "Allum, Mr. Owen George", 
"Andersen-Jensen, Miss. Carla Christine Nielsine", "Anderson, Mr. Harry", 
"Andersson, Master. Sigvard Harald Elias", "Andersson, Miss. Ebba Iris Alfrida", 
"Andersson, Miss. Ellis Anna Maria", "Andersson, Miss. Erna Alexandra", 
"Andersson, Miss. Ingeborg Constanzia", "Andersson, Miss. Sigrid Elisabeth", 
"Andersson, Mr. Anders Johan", "Andersson, Mr. August Edvard (\"Wennerstrom\")", 
"Andersson, Mrs. Anders Johan (Alfrida Konstantia Brogren)", 
"Andreasson, Mr. Paul Edvin", "Andrew, Mr. Edgardo Samuel", "Andrews, Miss. Kornelia Theodosia", 
"Andrews, Mr. Thomas Jr", "Angle, Mrs. William A (Florence \"Mary\" Agnes Hughes)", 
"Appleton, Mrs. Edward Dale (Charlotte Lamson)", "Arnold-Franchi, Mr. Josef", 
"Arnold-Franchi, Mrs. Josef (Josefine Franchi)", "Artagaveytia, Mr. Ramon", 
"Asim, Mr. Adola", "Asplund, Master. Clarence Gustaf Hugo", "Asplund, Master. Edvin Rojj Felix", 
"Asplund, Miss. Lillian Gertrud", "Asplund, Mrs. Carl Oscar (Selma Augusta Emilia Johansson)", 
"Astor, Mrs. John Jacob (Madeleine Talmadge Force)", "Attalah, Miss. Malake", 
"Attalah, Mr. Sleiman", "Aubart, Mme. Leontine Pauline", "Augustsson, Mr. Albert", 
"Ayoub, Miss. Banoura", "Backstrom, Mr. Karl Alfred", "Backstrom, Mrs. Karl Alfred (Maria Mathilda Gustafsson)", 
"Baclini, Miss. Eugenie", "Baclini, Miss. Helene Barbara", "Baclini, Miss. Marie Catherine", 
"Baclini, Mrs. Solomon (Latifa Qurban)", "Badt, Mr. Mohamed", 
"Bailey, Mr. Percy Andrew", "Balkic, Mr. Cerin", "Ball, Mrs. (Ada E Hall)", 
"Banfield, Mr. Frederick James", "Barah, Mr. Hanna Assi", "Barbara, Miss. Saiide", 
"Barbara, Mrs. (Catherine David)", "Barber, Miss. Ellen \"Nellie\"", 
"Barkworth, Mr. Algernon Henry Wilson", "Barton, Mr. David John", 
"Bateman, Rev. Robert James", "Baumann, Mr. John D", "Baxter, Mr. Quigg Edmond", 
"Baxter, Mrs. James (Helene DeLaudeniere Chaput)", "Bazzani, Miss. Albina", 
"Beane, Mr. Edward", "Beane, Mrs. Edward (Ethel Clarke)", "Beavan, Mr. William Thomas", 
"Becker, Master. Richard F", "Becker, Miss. Marion Louise", "Beckwith, Mr. Richard Leonard", 
"Beckwith, Mrs. Richard Leonard (Sallie Monypeny)", "Beesley, Mr. Lawrence", 
"Behr, Mr. Karl Howell", "Bengtsson, Mr. John Viktor", "Berglund, Mr. Karl Ivar Sven", 
"Berriman, Mr. William John", "Betros, Mr. Tannous", "Bidois, Miss. Rosalie", 
"Bing, Mr. Lee", "Birkeland, Mr. Hans Martin Monsen", "Bishop, Mr. Dickinson H", 
"Bishop, Mrs. Dickinson H (Helen Walton)", "Bissette, Miss. Amelia", 
"Bjornstrom-Steffansson, Mr. Mauritz Hakan", "Blackwell, Mr. Stephen Weart", 
"Blank, Mr. Henry", "Bonnell, Miss. Elizabeth", "Bostandyeff, Mr. Guentcho", 
"Boulos, Miss. Nourelain", "Boulos, Mr. Hanna", "Boulos, Mrs. Joseph (Sultana)", 
"Bourke, Miss. Mary", "Bourke, Mr. John", "Bourke, Mrs. John (Catherine)", 
"Bowen, Mr. David John \"Dai\"", "Bowerman, Miss. Elsie Edith", 
"Bracken, Mr. James H", "Bradley, Mr. George (\"George Arthur Brayton\")", 
"Braund, Mr. Lewis Richard", "Braund, Mr. Owen Harris", "Brewe, Dr. Arthur Jackson", 
"Brocklebank, Mr. William Alfred", "Brown, Miss. Amelia \"Mildred\"", 
"Brown, Mr. Thomas William Solomon", "Brown, Mrs. James Joseph (Margaret Tobin)", 
"Brown, Mrs. Thomas William Solomon (Elizabeth Catherine Ford)", 
"Bryhl, Mr. Kurt Arnold Gottfrid", "Burke, Mr. Jeremiah", "Burns, Miss. Elizabeth Margaret", 
"Buss, Miss. Kate", "Butler, Mr. Reginald Fenton", "Butt, Major. Archibald Willingham", 
"Byles, Rev. Thomas Roussel Davids", "Bystrom, Mrs. (Karolina)", 
"Cacic, Miss. Marija", "Cacic, Mr. Luka", "Cairns, Mr. Alexander", 
"Calderhead, Mr. Edward Pennington", "Caldwell, Master. Alden Gates", 
"Caldwell, Mrs. Albert Francis (Sylvia Mae Harbaugh)", "Calic, Mr. Jovo", 
"Calic, Mr. Petar", "Cameron, Miss. Clear Annie", "Campbell, Mr. William", 
"Canavan, Miss. Mary", "Cann, Mr. Ernest Charles", "Caram, Mrs. Joseph (Maria Elias)", 
"Carbines, Mr. William", "Cardeza, Mr. Thomas Drake Martinez", 
"Carlsson, Mr. August Sigfrid", "Carlsson, Mr. Frans Olof", "Carr, Miss. Helen \"Ellen\"", 
"Carrau, Mr. Francisco M", "Carter, Master. William Thornton II", 
"Carter, Miss. Lucile Polk", "Carter, Mr. William Ernest", "Carter, Mrs. Ernest Courtenay (Lilian Hughes)", 
"Carter, Mrs. William Ernest (Lucile Polk)", "Carter, Rev. Ernest Courtenay", 
"Cavendish, Mr. Tyrell William", "Celotti, Mr. Francesco", "Chaffee, Mr. Herbert Fuller", 
"Chambers, Mr. Norman Campbell", "Chambers, Mrs. Norman Campbell (Bertha Griggs)", 
"Chapman, Mr. Charles Henry", "Chapman, Mr. John Henry", "Charters, Mr. David", 
"Cherry, Miss. Gladys", "Chibnall, Mrs. (Edith Martha Bowerman)", 
"Chip, Mr. Chang", "Christmann, Mr. Emil", "Christy, Miss. Julie Rachel", 
"Chronopoulos, Mr. Apostolos", "Clarke, Mrs. Charles V (Ada Maria Winfield)", 
"Cleaver, Miss. Alice", "Clifford, Mr. George Quincy", "Coelho, Mr. Domingos Fernandeo", 
"Cohen, Mr. Gurshon \"Gus\"", "Coleff, Mr. Peju", "Coleff, Mr. Satio", 
"Coleridge, Mr. Reginald Charles", "Collander, Mr. Erik Gustaf", 
"Colley, Mr. Edward Pomeroy", "Collyer, Miss. Marjorie \"Lottie\"", 
"Collyer, Mr. Harvey", "Collyer, Mrs. Harvey (Charlotte Annie Tate)", 
"Compton, Miss. Sara Rebecca", "Connaghton, Mr. Michael", "Connolly, Miss. Kate", 
"Connors, Mr. Patrick", "Cook, Mr. Jacob", "Cor, Mr. Liudevit", 
"Corn, Mr. Harry", "Coutts, Master. Eden Leslie \"Neville\"", 
"Coutts, Master. William Loch \"William\"", "Coxon, Mr. Daniel", 
"Crease, Mr. Ernest James", "Cribb, Mr. John Hatfield", "Crosby, Capt. Edward Gifford", 
"Crosby, Miss. Harriet R", "Culumovic, Mr. Jeso", "Cumings, Mrs. John Bradley (Florence Briggs Thayer)", 
"Cunningham, Mr. Alfred Fleming", "Dahl, Mr. Karl Edwart", "Dahlberg, Miss. Gerda Ulrika", 
"Dakic, Mr. Branko", "Daly, Mr. Eugene Patrick", "Daly, Mr. Peter Denis ", 
"Danbom, Mr. Ernst Gilbert", "Danbom, Mrs. Ernst Gilbert (Anna Sigrid Maria Brogren)", 
"Daniel, Mr. Robert Williams", "Danoff, Mr. Yoto", "Dantcheff, Mr. Ristiu", 
"Davidson, Mr. Thornton", "Davies, Master. John Morgan Jr", "Davies, Mr. Alfred J", 
"Davies, Mr. Charles Henry", "Davis, Miss. Mary", "Davison, Mrs. Thomas Henry (Mary E Finck)", 
"Dean, Master. Bertram Vere", "Dean, Mr. Bertram Frank", "Denkoff, Mr. Mitto", 
"Dennis, Mr. Samuel", "Devaney, Miss. Margaret Delia", "Dick, Mr. Albert Adrian", 
"Dick, Mrs. Albert Adrian (Vera Gillespie)", "Dimic, Mr. Jovan", 
"Dodge, Master. Washington", "Doharr, Mr. Tannous", "Doling, Miss. Elsie", 
"Doling, Mrs. John T (Ada Julia Bone)", "Dooley, Mr. Patrick", 
"Dorking, Mr. Edward Arthur", "Douglas, Mr. Walter Donald", "Dowdell, Miss. Elizabeth", 
"Downton, Mr. William James", "Drazenoic, Mr. Jozef", "Drew, Mrs. James Vivian (Lulu Thorne Christian)", 
"Duane, Mr. Frank", "Duff Gordon, Lady. (Lucille Christiana Sutherland) (\"Mrs Morgan\")", 
"Duff Gordon, Sir. Cosmo Edmund (\"Mr Morgan\")", "Duran y More, Miss. Asuncion", 
"Edvardsson, Mr. Gustaf Hjalmar", "Eitemiller, Mr. George Floyd", 
"Eklund, Mr. Hans Linus", "Ekstrom, Mr. Johan", "Elias, Mr. Dibo", 
"Elias, Mr. Joseph Jr", "Elias, Mr. Tannous", "Elsbury, Mr. William James", 
"Emanuel, Miss. Virginia Ethel", "Emir, Mr. Farred Chehab", "Endres, Miss. Caroline Louise", 
"Eustis, Miss. Elizabeth Mussey", "Fahlstrom, Mr. Arne Jonas", 
"Farrell, Mr. James", "Farthing, Mr. John", "Faunthorpe, Mrs. Lizzie (Elizabeth Anne Wilkinson)", 
"Fischer, Mr. Eberhard Thelander", "Fleming, Miss. Margaret", 
"Flynn, Mr. James", "Flynn, Mr. John", "Flynn, Mr. John Irwin (\"Irving\")", 
"Foo, Mr. Choong", "Ford, Miss. Doolina Margaret \"Daisy\"", 
"Ford, Miss. Robina Maggie \"Ruby\"", "Ford, Mr. William Neal", 
"Ford, Mrs. Edward (Margaret Ann Watson)", "Foreman, Mr. Benjamin Laventall", 
"Fortune, Miss. Alice Elizabeth", "Fortune, Miss. Mabel Helen", 
"Fortune, Mr. Charles Alexander", "Fortune, Mr. Mark", "Fox, Mr. Stanley Hubert", 
"Francatelli, Miss. Laura Mabel", "Frauenthal, Dr. Henry William", 
"Frauenthal, Mrs. Henry William (Clara Heinsheimer)", "Frolicher, Miss. Hedwig Margaritha", 
"Frolicher-Stehli, Mr. Maxmillian", "Frost, Mr. Anthony Wood \"Archie\"", 
"Fry, Mr. Richard", "Funk, Miss. Annie Clemmer", "Futrelle, Mr. Jacques Heath", 
"Futrelle, Mrs. Jacques Heath (Lily May Peel)", "Fynney, Mr. Joseph J", 
"Gale, Mr. Shadrach", "Gallagher, Mr. Martin", "Garfirth, Mr. John", 
"Garside, Miss. Ethel", "Gaskell, Mr. Alfred", "Gavey, Mr. Lawrence", 
"Gee, Mr. Arthur H", "Gheorgheff, Mr. Stanio", "Giglio, Mr. Victor", 
"Giles, Mr. Frederick Edward", "Gilinski, Mr. Eliezer", "Gill, Mr. John William", 
"Gillespie, Mr. William Henry", "Gilnagh, Miss. Katherine \"Katie\"", 
"Givard, Mr. Hans Kristensen", "Glynn, Miss. Mary Agatha", "Goldenberg, Mr. Samuel L", 
"Goldenberg, Mrs. Samuel L (Edwiga Grabowska)", "Goldschmidt, Mr. George B", 
"Goldsmith, Master. Frank John William \"Frankie\"", "Goldsmith, Mr. Frank John", 
"Goldsmith, Mrs. Frank John (Emily Alice Brown)", "Goncalves, Mr. Manuel Estanslas", 
"Goodwin, Master. Harold Victor", "Goodwin, Master. Sidney Leonard", 
"Goodwin, Master. William Frederick", "Goodwin, Miss. Lillian Amy", 
"Goodwin, Mr. Charles Edward", "Goodwin, Mrs. Frederick (Augusta Tyler)", 
"Graham, Miss. Margaret Edith", "Graham, Mr. George Edward", 
"Graham, Mrs. William Thompson (Edith Junkins)", "Green, Mr. George Henry", 
"Greenberg, Mr. Samuel", "Greenfield, Mr. William Bertram", "Gronnestad, Mr. Daniel Danielsen", 
"Guggenheim, Mr. Benjamin", "Gustafsson, Mr. Alfred Ossian", 
"Gustafsson, Mr. Anders Vilhelm", "Gustafsson, Mr. Johan Birger", 
"Gustafsson, Mr. Karl Gideon", "Haas, Miss. Aloisia", "Hagland, Mr. Ingvald Olai Olsen", 
"Hagland, Mr. Konrad Mathias Reiersen", "Hakkarainen, Mr. Pekka Pietari", 
"Hakkarainen, Mrs. Pekka Pietari (Elin Matilda Dolck)", "Hale, Mr. Reginald", 
"Hamalainen, Master. Viljo", "Hamalainen, Mrs. William (Anna)", 
"Hampe, Mr. Leon", "Hanna, Mr. Mansour", "Hansen, Mr. Claus Peter", 
"Hansen, Mr. Henrik Juul", "Hansen, Mr. Henry Damsgaard", "Harder, Mr. George Achilles", 
"Harknett, Miss. Alice Phoebe", "Harmer, Mr. Abraham (David Lishin)", 
"Harper, Miss. Annie Jessie \"Nina\"", "Harper, Mr. Henry Sleeper", 
"Harper, Mrs. Henry Sleeper (Myna Haxtun)", "Harper, Rev. John", 
"Harrington, Mr. Charles H", "Harris, Mr. George", "Harris, Mr. Henry Birkhardt", 
"Harris, Mr. Walter", "Harris, Mrs. Henry Birkhardt (Irene Wallach)", 
"Harrison, Mr. William", "Hart, Miss. Eva Miriam", "Hart, Mr. Benjamin", 
"Hart, Mr. Henry", "Hart, Mrs. Benjamin (Esther Ada Bloomfield)", 
"Hassab, Mr. Hammad", "Hassan, Mr. Houssein G N", "Hawksford, Mr. Walter James", 
"Hays, Miss. Margaret Bechstein", "Hays, Mrs. Charles Melville (Clara Jennings Gregg)", 
"Healy, Miss. Hanora \"Nora\"", "Hedman, Mr. Oskar Arvid", "Hegarty, Miss. Hanora \"Nora\"", 
"Heikkinen, Miss. Laina", "Heininen, Miss. Wendla Maria", "Hendekovic, Mr. Ignjac", 
"Henry, Miss. Delia", "Herman, Miss. Alice", "Herman, Mrs. Samuel (Jane Laver)", 
"Hewlett, Mrs. (Mary D Kingcome) ", "Hickman, Mr. Leonard Mark", 
"Hickman, Mr. Lewis", "Hickman, Mr. Stanley George", "Hippach, Miss. Jean Gertrude", 
"Hippach, Mrs. Louis Albert (Ida Sophia Fischer)", "Hirvonen, Miss. Hildur E", 
"Hocking, Mr. Richard George", "Hocking, Mrs. Elizabeth (Eliza Needs)", 
"Hodges, Mr. Henry Price", "Hogeboom, Mrs. John C (Anna Andrews)", 
"Hold, Mr. Stephen", "Holm, Mr. John Fredrik Alexander", "Holverson, Mr. Alexander Oskar", 
"Holverson, Mrs. Alexander Oskar (Mary Aline Towner)", "Homer, Mr. Harry (\"Mr E Haven\")", 
"Honkanen, Miss. Eliina", "Hood, Mr. Ambrose Jr", "Horgan, Mr. John", 
"Hosono, Mr. Masabumi", "Hoyt, Mr. Frederick Maxfield", "Hoyt, Mr. William Fisher", 
"Hoyt, Mrs. Frederick Maxfield (Jane Anne Forby)", "Humblen, Mr. Adolf Mathias Nicolai Olsen", 
"Hunt, Mr. George Henry", "Ibrahim Shawah, Mr. Yousseff", "Icard, Miss. Amelie", 
"Ilett, Miss. Bertha", "Ilmakangas, Miss. Pieta Sofia", "Isham, Miss. Ann Elizabeth", 
"Ivanoff, Mr. Kanio", "Jacobsohn, Mr. Sidney Samuel", "Jacobsohn, Mrs. Sidney Samuel (Amy Frances Christy)", 
"Jalsevac, Mr. Ivan", "Jansson, Mr. Carl Olof", "Jardin, Mr. Jose Neto", 
"Jarvis, Mr. John Denzil", "Jenkin, Mr. Stephen Curnow", "Jensen, Mr. Hans Peder", 
"Jensen, Mr. Niels Peder", "Jensen, Mr. Svend Lauritz", "Jermyn, Miss. Annie", 
"Jerwan, Mrs. Amin S (Marie Marthe Thuillard)", "Johannesen-Bratthammer, Mr. Bernt", 
"Johanson, Mr. Jakob Alfred", "Johansson, Mr. Erik", "Johansson, Mr. Gustaf Joel", 
"Johansson, Mr. Karl Johan", "Johnson, Master. Harold Theodor", 
"Johnson, Miss. Eleanor Ileen", "Johnson, Mr. Alfred", "Johnson, Mr. Malkolm Joackim", 
"Johnson, Mr. William Cahoone Jr", "Johnson, Mrs. Oscar W (Elisabeth Vilhelmina Berg)", 
"Johnston, Miss. Catherine Helen \"Carrie\"", "Johnston, Mr. Andrew G", 
"Jonkoff, Mr. Lalio", "Jonsson, Mr. Carl", "Jussila, Miss. Katriina", 
"Jussila, Miss. Mari Aina", "Jussila, Mr. Eiriik", "Kallio, Mr. Nikolai Erland", 
"Kalvik, Mr. Johannes Halvorsen", "Kantor, Mr. Sinai", "Kantor, Mrs. Sinai (Miriam Sternin)", 
"Karaic, Mr. Milan", "Karlsson, Mr. Nils August", "Karun, Miss. Manca", 
"Kassem, Mr. Fared", "Keane, Miss. Nora A", "Keane, Mr. Andrew \"Andy\"", 
"Keefe, Mr. Arthur", "Kelly, Miss. Anna Katherine \"Annie Kate\"", 
"Kelly, Miss. Mary", "Kelly, Mr. James", "Kelly, Mrs. Florence \"Fannie\"", 
"Kent, Mr. Edward Austin", "Kenyon, Mrs. Frederick R (Marion)", 
"Kiernan, Mr. Philip", "Kilgannon, Mr. Thomas J", "Kimball, Mr. Edwin Nelson Jr", 
"Kink, Mr. Vincenz", "Kink-Heilmann, Miss. Luise Gretchen", "Kirkland, Rev. Charles Leonard", 
"Klaber, Mr. Herman", "Klasen, Mr. Klas Albin", "Knight, Mr. Robert J", 
"Kraeff, Mr. Theodor", "Kvillner, Mr. Johan Henrik Johannesson", 
"Lahoud, Mr. Sarkis", "Lahtinen, Mrs. William (Anna Sylfven)", 
"Laitinen, Miss. Kristina Sofia", "Laleff, Mr. Kristo", "Lam, Mr. Ali", 
"Lam, Mr. Len", "Landergren, Miss. Aurora Adelia", "Lang, Mr. Fang", 
"Laroche, Miss. Simonne Marie Anne Andree", "Laroche, Mr. Joseph Philippe Lemercier", 
"Laroche, Mrs. Joseph (Juliette Marie Louise Lafargue)", "Larsson, Mr. August Viktor", 
"Larsson, Mr. Bengt Edvin", "LeRoy, Miss. Bertha", "Leader, Dr. Alice (Farnham)", 
"Leeni, Mr. Fahim (\"Philip Zenni\")", "Lefebre, Master. Henry Forbes", 
"Lefebre, Miss. Ida", "Lefebre, Miss. Jeannie", "Lefebre, Miss. Mathilde", 
"Lehmann, Miss. Bertha", "Leinonen, Mr. Antti Gustaf", "Leitch, Miss. Jessie Wills", 
"Lemberopolous, Mr. Peter L", "Lemore, Mrs. (Amelia Milley)", 
"Lennon, Mr. Denis", "Leonard, Mr. Lionel", "Lester, Mr. James", 
"Lesurer, Mr. Gustave J", "Levy, Mr. Rene Jacques", "Lewy, Mr. Ervin G", 
"Leyson, Mr. Robert William Norman", "Lievens, Mr. Rene Aime", 
"Lindahl, Miss. Agda Thorilda Viktoria", "Lindblom, Miss. Augusta Charlotta", 
"Lindell, Mr. Edvard Bengtsson", "Lindqvist, Mr. Eino William", 
"Lines, Miss. Mary Conover", "Ling, Mr. Lee", "Lobb, Mr. William Arthur", 
"Lobb, Mrs. William Arthur (Cordelia K Stanlick)", "Long, Mr. Milton Clyde", 
"Longley, Miss. Gretchen Fiske", "Louch, Mrs. Charles Alexander (Alice Adelaide Slow)", 
"Lovell, Mr. John Hall (\"Henry\")", "Lulic, Mr. Nikola", "Lundahl, Mr. Johan Svensson", 
"Lurette, Miss. Elise", "Mack, Mrs. (Mary)", "Madigan, Miss. Margaret \"Maggie\"", 
"Madill, Miss. Georgette Alexandra", "Madsen, Mr. Fridtjof Arne", 
"Maenpaa, Mr. Matti Alexanteri", "Maioni, Miss. Roberta", "Maisner, Mr. Simon", 
"Mallet, Master. Andre", "Mallet, Mr. Albert", "Mamee, Mr. Hanna", 
"Mangan, Miss. Mary", "Mannion, Miss. Margareth", "Marechal, Mr. Pierre", 
"Markoff, Mr. Marin", "Markun, Mr. Johann", "Marvin, Mr. Daniel Warner", 
"Masselmani, Mrs. Fatima", "Matthews, Mr. William John", "Mayne, Mlle. Berthe Antonine (\"Mrs de Villiers\")", 
"McCarthy, Mr. Timothy J", "McCormack, Mr. Thomas Joseph", "McCoy, Miss. Agnes", 
"McCoy, Mr. Bernard", "McDermott, Miss. Brigdet Delia", "McEvoy, Mr. Michael", 
"McGough, Mr. James Robert", "McGovern, Miss. Mary", "McGowan, Miss. Anna \"Annie\"", 
"McKane, Mr. Peter David", "McMahon, Mr. Martin", "McNamee, Mr. Neal", 
"Meanwell, Miss. (Marion Ogden)", "Meek, Mrs. Thomas (Annie Louise Rowley)", 
"Mellinger, Miss. Madeleine Violet", "Mellinger, Mrs. (Elizabeth Anne Maidment)", 
"Mellors, Mr. William John", "Meo, Mr. Alfonzo", "Mernagh, Mr. Robert", 
"Meyer, Mr. August", "Meyer, Mr. Edgar Joseph", "Meyer, Mrs. Edgar Joseph (Leila Saks)", 
"Millet, Mr. Francis Davis", "Milling, Mr. Jacob Christian", 
"Minahan, Dr. William Edward", "Minahan, Miss. Daisy E", "Mineff, Mr. Ivan", 
"Mionoff, Mr. Stoytcho", "Mitchell, Mr. Henry Michael", "Mitkoff, Mr. Mito", 
"Mockler, Miss. Helen Mary \"Ellie\"", "Moen, Mr. Sigurd Hansen", 
"Molson, Mr. Harry Markland", "Montvila, Rev. Juozas", "Moor, Master. Meier", 
"Moor, Mrs. (Beila)", "Moore, Mr. Leonard Charles", "Moran, Miss. Bertha", 
"Moran, Mr. Daniel J", "Moran, Mr. James", "Moraweck, Dr. Ernest", 
"Morley, Mr. Henry Samuel (\"Mr Henry Marshall\")", "Morley, Mr. William", 
"Morrow, Mr. Thomas Rowan", "Moss, Mr. Albert Johan", "Moubarek, Master. Gerios", 
"Moubarek, Master. Halim Gonios (\"William George\")", "Moussa, Mrs. (Mantoura Boulos)", 
"Moutal, Mr. Rahamin Haim", "Mudd, Mr. Thomas Charles", "Mullens, Miss. Katherine \"Katie\"", 
"Murdlin, Mr. Joseph", "Murphy, Miss. Katherine \"Kate\"", "Murphy, Miss. Margaret Jane", 
"Myhrman, Mr. Pehr Fabian Oliver Malkolm", "Naidenoff, Mr. Penko", 
"Najib, Miss. Adele Kiamie \"Jane\"", "Nakid, Miss. Maria (\"Mary\")", 
"Nakid, Mr. Sahid", "Nankoff, Mr. Minko", "Nasser, Mr. Nicholas", 
"Nasser, Mrs. Nicholas (Adele Achem)", "Natsch, Mr. Charles H", 
"Navratil, Master. Edmond Roger", "Navratil, Master. Michel M", 
"Navratil, Mr. Michel (\"Louis M Hoffman\")", "Nenkoff, Mr. Christo", 
"Newell, Miss. Madeleine", "Newell, Miss. Marjorie", "Newell, Mr. Arthur Webster", 
"Newsom, Miss. Helen Monypeny", "Nicholls, Mr. Joseph Charles", 
"Nicholson, Mr. Arthur Ernest", "Nicola-Yarred, Master. Elias", 
"Nicola-Yarred, Miss. Jamila", "Nilsson, Miss. Helmina Josefina", 
"Nirva, Mr. Iisakki Antino Aijo", "Niskanen, Mr. Juha", "Norman, Mr. Robert Douglas", 
"Nosworthy, Mr. Richard Cater", "Novel, Mr. Mansouer", "Nye, Mrs. (Elizabeth Ramell)", 
"Nysten, Miss. Anna Sofia", "Nysveen, Mr. Johan Hansen", "O'Brien, Mr. Thomas", 
"O'Brien, Mr. Timothy", "O'Brien, Mrs. Thomas (Johanna \"Hannah\" Godfrey)", 
"O'Connell, Mr. Patrick D", "O'Connor, Mr. Maurice", "O'Driscoll, Miss. Bridget", 
"O'Dwyer, Miss. Ellen \"Nellie\"", "O'Leary, Miss. Hanora \"Norah\"", 
"O'Sullivan, Miss. Bridget Mary", "Odahl, Mr. Nils Martin", "Ohman, Miss. Velin", 
"Olsen, Mr. Henry Margido", "Olsen, Mr. Karl Siegwart Andreas", 
"Olsen, Mr. Ole Martin", "Olsson, Miss. Elina", "Olsson, Mr. Nils Johan Goransson", 
"Olsvigen, Mr. Thor Anderson", "Oreskovic, Miss. Marija", "Oreskovic, Mr. Luka", 
"Osen, Mr. Olaf Elon", "Osman, Mrs. Mara", "Ostby, Mr. Engelhart Cornelius", 
"Otter, Mr. Richard", "Padro y Manent, Mr. Julian", "Pain, Dr. Alfred", 
"Palsson, Master. Gosta Leonard", "Palsson, Miss. Stina Viola", 
"Palsson, Miss. Torborg Danira", "Palsson, Mrs. Nils (Alma Cornelia Berglund)", 
"Panula, Master. Eino Viljami", "Panula, Master. Juha Niilo", 
"Panula, Master. Urho Abraham", "Panula, Mr. Ernesti Arvid", 
"Panula, Mr. Jaako Arnold", "Panula, Mrs. Juha (Maria Emilia Ojala)", 
"Parkes, Mr. Francis \"Frank\"", "Parr, Mr. William Henry Marsh", 
"Parrish, Mrs. (Lutie Davis)", "Partner, Mr. Austen", "Pasic, Mr. Jakob", 
"Patchett, Mr. George", "Paulner, Mr. Uscher", "Pavlovic, Mr. Stefo", 
"Pears, Mr. Thomas Clinton", "Pears, Mrs. Thomas (Edith Wearne)", 
"Peduzzi, Mr. Joseph", "Pekoniemi, Mr. Edvard", "Penasco y Castellana, Mr. Victor de Satode", 
"Penasco y Castellana, Mrs. Victor de Satode (Maria Josefa Perez de Soto y Vallejo)", 
"Pengelly, Mr. Frederick William", "Perkin, Mr. John Henry", 
"Pernot, Mr. Rene", "Perreault, Miss. Anne", "Persson, Mr. Ernst Ulrik", 
"Peter, Miss. Anna", "Peter, Mrs. Catherine (Catherine Rizk)", 
"Peters, Miss. Katie", "Petranec, Miss. Matilda", "Petroff, Mr. Nedelio", 
"Petroff, Mr. Pastcho (\"Pentcho\")", "Petterson, Mr. Johan Emil", 
"Pettersson, Miss. Ellen Natalia", "Peuchen, Major. Arthur Godfrey", 
"Phillips, Miss. Kate Florence (\"Mrs Kate Louise Phillips Marshall\")", 
"Pickard, Mr. Berk (Berk Trembisky)", "Pinsky, Mrs. (Rosa)", 
"Plotcharsky, Mr. Vasil", "Ponesell, Mr. Martin", "Porter, Mr. Walter Chamberlain", 
"Potter, Mrs. Thomas Jr (Lily Alexenia Wilson)", "Quick, Miss. Phyllis May", 
"Quick, Mrs. Frederick Charles (Jane Richards)", "Radeff, Mr. Alexander", 
"Razi, Mr. Raihed", "Reed, Mr. James George", "Reeves, Mr. David", 
"Rekic, Mr. Tido", "Renouf, Mr. Peter Henry", "Renouf, Mrs. Peter Henry (Lillian Jefferys)", 
"Reuchlin, Jonkheer. John George", "Reynaldo, Ms. Encarnacion", 
"Rice, Master. Arthur", "Rice, Master. Eric", "Rice, Master. Eugene", 
"Rice, Master. George Hugh", "Rice, Mrs. William (Margaret Norton)", 
"Richard, Mr. Emile", "Richards, Master. George Sibley", "Richards, Master. William Rowe", 
"Richards, Mrs. Sidney (Emily Hocking)", "Ridsdale, Miss. Lucy", 
"Ringhini, Mr. Sante", "Rintamaki, Mr. Matti", "Risien, Mr. Samuel Beard", 
"Robbins, Mr. Victor", "Robert, Mrs. Edward Scott (Elisabeth Walton McMillan)", 
"Robins, Mrs. Alexander A (Grace Charity Laury)", "Roebling, Mr. Washington Augustus II", 
"Rogers, Mr. William John", "Romaine, Mr. Charles Hallace (\"Mr C Rolmane\")", 
"Rommetvedt, Mr. Knud Paust", "Rood, Mr. Hugh Roscoe", "Rosblom, Mr. Viktor Richard", 
"Rosblom, Mrs. Viktor (Helena Wilhelmina)", "Ross, Mr. John Hugo", 
"Rothes, the Countess. of (Lucy Noel Martha Dyer-Edwards)", "Rothschild, Mrs. Martin (Elizabeth L. Barrett)", 
"Rouse, Mr. Richard Henry", "Rugg, Miss. Emily", "Rush, Mr. Alfred George John", 
"Ryan, Mr. Patrick", "Ryerson, Miss. Emily Borie", "Ryerson, Miss. Susan Parker \"Suzette\"", 
"Saad, Mr. Amin", "Saad, Mr. Khalil", "Saalfeld, Mr. Adolphe", 
"Sadlier, Mr. Matthew", "Sage, Master. Thomas Henry", "Sage, Miss. Constance Gladys", 
"Sage, Miss. Dorothy Edith \"Dolly\"", "Sage, Miss. Stella Anna", 
"Sage, Mr. Douglas Bullen", "Sage, Mr. Frederick", "Sage, Mr. George John Jr", 
"Sagesser, Mlle. Emma", "Salkjelsvik, Miss. Anna Kristine", "Salonen, Mr. Johan Werner", 
"Samaan, Mr. Youssef", "Sandstrom, Miss. Marguerite Rut", "Sandstrom, Mrs. Hjalmar (Agnes Charlotta Bengtsson)", 
"Saundercock, Mr. William Henry", "Sawyer, Mr. Frederick Charles", 
"Scanlan, Mr. James", "Sdycoff, Mr. Todor", "Sedgwick, Mr. Charles Frederick Waddington", 
"Serepeca, Miss. Augusta", "Seward, Mr. Frederic Kimber", "Sharp, Mr. Percival James R", 
"Sheerlinck, Mr. Jan Baptist", "Shellard, Mr. Frederick William", 
"Shelley, Mrs. William (Imanita Parrish Hall)", "Shorney, Mr. Charles Joseph", 
"Shutes, Miss. Elizabeth W", "Silven, Miss. Lyyli Karoliina", 
"Silverthorne, Mr. Spencer Victor", "Silvey, Mr. William Baird", 
"Silvey, Mrs. William Baird (Alice Munger)", "Simmons, Mr. John", 
"Simonius-Blumer, Col. Oberst Alfons", "Sinkkonen, Miss. Anna", 
"Sirayanian, Mr. Orsen", "Sirota, Mr. Maurice", "Sivic, Mr. Husein", 
"Sivola, Mr. Antti Wilhelm", "Sjoblom, Miss. Anna Sofia", "Sjostedt, Mr. Ernst Adolf", 
"Skoog, Master. Harald", "Skoog, Master. Karl Thorsten", "Skoog, Miss. Mabel", 
"Skoog, Miss. Margit Elizabeth", "Skoog, Mr. Wilhelm", "Skoog, Mrs. William (Anna Bernhardina Karlsson)", 
"Slabenoff, Mr. Petco", "Slayter, Miss. Hilda Mary", "Slemen, Mr. Richard James", 
"Slocovski, Mr. Selman Francis", "Sloper, Mr. William Thompson", 
"Smart, Mr. John Montgomery", "Smiljanic, Mr. Mile", "Smith, Miss. Marion Elsie", 
"Smith, Mr. James Clinch", "Smith, Mr. Richard William", "Smith, Mr. Thomas", 
"Sobey, Mr. Samuel James Hayden", "Soholt, Mr. Peter Andreas Lauritz Andersen", 
"Somerton, Mr. Francis William", "Spedden, Mrs. Frederic Oakley (Margaretta Corning Stone)", 
"Spencer, Mrs. William Augustus (Marie Eugenie)", "Stahelin-Maeglin, Dr. Max", 
"Staneff, Mr. Ivan", "Stankovic, Mr. Ivan", "Stanley, Miss. Amy Zillah Elsie", 
"Stanley, Mr. Edward Roland", "Stead, Mr. William Thomas", "Stephenson, Mrs. Walter Bertram (Martha Eustis)", 
"Stewart, Mr. Albert A", "Stone, Mrs. George Nelson (Martha Evelyn)", 
"Stoytcheff, Mr. Ilia", "Strandberg, Miss. Ida Sofia", "Stranden, Mr. Juho", 
"Strom, Miss. Telma Matilda", "Strom, Mrs. Wilhelm (Elna Matilda Persson)", 
"Sunderland, Mr. Victor Francis", "Sundman, Mr. Johan Julian", 
"Sutehall, Mr. Henry Jr", "Sutton, Mr. Frederick", "Svensson, Mr. Johan", 
"Svensson, Mr. Olof", "Swift, Mrs. Frederick Joel (Margaret Welles Barron)", 
"Taussig, Miss. Ruth", "Taussig, Mr. Emil", "Taussig, Mrs. Emil (Tillie Mandelbaum)", 
"Taylor, Mr. Elmer Zebley", "Taylor, Mrs. Elmer Zebley (Juliet Cummins Wright)", 
"Thayer, Mr. John Borland", "Thayer, Mr. John Borland Jr", "Thayer, Mrs. John Borland (Marian Longstreth Morris)", 
"Theobald, Mr. Thomas Leonard", "Thomas, Master. Assad Alexander", 
"Thorne, Mrs. Gertrude Maybelle", "Thorneycroft, Mr. Percival", 
"Thorneycroft, Mrs. Percival (Florence Kate White)", "Tikkanen, Mr. Juho", 
"Tobin, Mr. Roger", "Todoroff, Mr. Lalio", "Tomlin, Mr. Ernest Portage", 
"Toomey, Miss. Ellen", "Torber, Mr. Ernst William", "Tornquist, Mr. William Henry", 
"Toufik, Mr. Nakli", "Touma, Mrs. Darwis (Hanne Youssef Razi)", 
"Troupiansky, Mr. Moses Aaron", "Trout, Mrs. William H (Jessie L)", 
"Troutt, Miss. Edwina Celia \"Winnie\"", "Turcin, Mr. Stjepan", 
"Turja, Miss. Anna Sofia", "Turkula, Mrs. (Hedwig)", "Turpin, Mr. William John Robert", 
"Turpin, Mrs. William John Robert (Dorothy Ann Wonnacott)", "Uruchurtu, Don. Manuel E", 
"Van Impe, Miss. Catharina", "Van Impe, Mr. Jean Baptiste", "Van Impe, Mrs. Jean Baptiste (Rosalie Paula Govaert)", 
"Van der hoef, Mr. Wyckoff", "Vande Velde, Mr. Johannes Joseph", 
"Vande Walle, Mr. Nestor Cyriel", "Vanden Steen, Mr. Leo Peter", 
"Vander Cruyssen, Mr. Victor", "Vander Planke, Miss. Augusta Maria", 
"Vander Planke, Mr. Leo Edmondus", "Vander Planke, Mrs. Julius (Emelia Maria Vandemoortele)", 
"Vestrom, Miss. Hulda Amanda Adolfina", "Vovk, Mr. Janko", "Waelens, Mr. Achille", 
"Walker, Mr. William Anderson", "Ward, Miss. Anna", "Warren, Mrs. Frank Manley (Anna Sophia Atkinson)", 
"Watson, Mr. Ennis Hastings", "Watt, Mrs. James (Elizabeth \"Bessie\" Inglis Milne)", 
"Webber, Miss. Susan", "Webber, Mr. James", "Weir, Col. John", 
"Weisz, Mrs. Leopold (Mathilde Francoise Pede)", "Wells, Miss. Joan", 
"West, Miss. Constance Mirium", "West, Mr. Edwy Arthur", "West, Mrs. Edwy Arthur (Ada Mary Worth)", 
"Wheadon, Mr. Edward H", "White, Mr. Percival Wayland", "White, Mr. Richard Frasar", 
"Wick, Miss. Mary Natalie", "Wick, Mrs. George Dennick (Mary Hitchcock)", 
"Widegren, Mr. Carl/Charles Peter", "Widener, Mr. Harry Elkins", 
"Wiklund, Mr. Jakob Alfred", "Wilhelms, Mr. Charles", "Willey, Mr. Edward", 
"Williams, Mr. Charles Duane", "Williams, Mr. Charles Eugene", 
"Williams, Mr. Howard Hugh \"Harry\"", "Williams, Mr. Leslie", 
"Williams-Lambert, Mr. Fletcher Fellows", "Windelov, Mr. Einar", 
"Wiseman, Mr. Phillippe", "Woolner, Mr. Hugh", "Wright, Mr. George", 
"Yasbeck, Mr. Antoni", "Yasbeck, Mrs. Antoni (Selini Alexander)", 
"Young, Miss. Marie Grice", "Youseff, Mr. Gerious", "Yousif, Mr. Wazli", 
"Yousseff, Mr. Gerious", "Yrois, Miss. Henriette (\"Mrs Harbeck\")", 
"Zabour, Miss. Hileni", "Zabour, Miss. Thamine", "Zimmerman, Mr. Leo", 
"de Messemaeker, Mrs. Guillaume Joseph (Emma)", "de Mulder, Mr. Theodore", 
"de Pelsmaeker, Mr. Alfons", "del Carlo, Mr. Sebastiano", "van Billiard, Mr. Austin Blyler", 
"van Melkebeke, Mr. Philemon"), class = "factor"), Sex = structure(c(2L, 
1L, 1L, 1L, 2L, 2L, 2L, 2L, 1L, 1L, 1L, 1L, 2L, 2L, 1L, 1L, 2L, 
2L, 1L, 1L, 2L, 2L, 1L, 2L, 1L, 1L, 2L, 2L, 1L, 2L, 2L, 1L, 1L, 
2L, 2L, 2L, 2L, 2L, 1L, 1L, 1L, 1L, 2L, 1L, 1L, 2L, 2L, 1L, 2L, 
1L, 2L, 2L, 1L, 1L, 2L, 2L, 1L, 2L, 1L, 2L, 2L, 1L, 2L, 2L, 2L, 
2L, 1L, 2L, 1L, 2L, 2L, 1L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 1L, 2L, 
2L, 1L, 2L, 1L, 1L, 2L, 2L, 1L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 
2L, 1L, 2L, 1L, 2L, 2L, 2L, 2L, 2L, 1L, 2L, 2L, 1L, 2L, 1L, 2L, 
1L, 1L, 2L, 2L, 2L, 2L, 1L, 2L, 2L, 2L, 1L, 2L, 2L, 2L, 2L, 1L, 
2L, 2L, 2L, 1L, 1L, 2L, 2L, 1L, 2L, 2L, 2L, 1L, 1L, 1L, 2L, 2L, 
2L, 2L, 1L, 2L, 2L, 2L, 1L, 2L, 2L, 2L, 2L, 1L, 2L, 2L, 2L, 2L, 
1L, 2L, 2L, 2L, 2L, 1L, 1L, 2L, 2L, 2L, 2L, 1L, 2L, 2L, 2L, 2L, 
1L, 2L, 2L, 1L, 2L, 2L, 2L, 1L, 2L, 1L, 2L, 2L, 2L, 1L, 2L, 1L, 
2L, 1L, 1L, 2L, 2L, 1L, 1L, 2L, 2L, 2L, 2L, 2L, 1L, 2L, 2L, 1L, 
2L, 2L, 1L, 2L, 2L, 2L, 1L, 1L, 2L, 1L, 2L, 2L, 2L, 2L, 2L, 2L, 
2L, 2L, 2L, 2L, 1L, 1L, 2L, 2L, 1L, 2L, 1L, 2L, 1L, 2L, 2L, 1L, 
1L, 2L, 2L, 2L, 2L, 1L, 1L, 2L, 2L, 2L, 1L, 2L, 2L, 1L, 1L, 1L, 
1L, 1L, 1L, 2L, 2L, 2L, 2L, 1L, 2L, 2L, 2L, 1L, 1L, 2L, 2L, 1L, 
2L, 1L, 1L, 1L, 2L, 2L, 1L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 
1L, 1L, 1L, 2L, 1L, 2L, 2L, 2L, 1L, 2L, 1L, 1L, 2L, 2L, 1L, 2L, 
2L, 1L, 1L, 2L, 1L, 1L, 1L, 1L, 2L, 2L, 1L, 1L, 2L, 1L, 1L, 2L, 
2L, 1L, 1L, 2L, 1L, 2L, 1L, 1L, 1L, 1L, 2L, 2L, 2L, 1L, 2L, 2L, 
1L, 2L, 2L, 2L, 1L, 2L, 2L, 2L, 1L, 1L, 1L, 2L, 2L, 2L, 2L, 2L, 
2L, 2L, 2L, 1L, 1L, 1L, 1L, 2L, 2L, 1L, 2L, 2L, 2L, 1L, 1L, 1L, 
1L, 2L, 2L, 2L, 2L, 1L, 1L, 1L, 2L, 2L, 2L, 1L, 1L, 2L, 1L, 2L, 
2L, 2L, 1L, 2L, 1L, 2L, 2L, 2L, 1L, 1L, 2L, 1L, 2L, 2L, 1L, 2L, 
2L, 1L, 2L, 1L, 2L, 2L, 2L, 2L, 1L, 2L, 2L, 1L, 2L, 2L, 1L, 1L, 
1L, 2L, 1L, 2L, 2L, 2L, 1L, 2L, 2L, 1L, 1L, 2L, 2L, 2L, 1L, 1L, 
2L, 2L, 1L, 1L, 1L, 2L, 2L, 1L, 2L, 2L, 1L, 2L, 2L, 1L, 2L, 1L, 
2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 1L, 1L, 2L, 2L, 2L, 2L, 2L, 2L, 
2L, 2L, 2L, 2L, 1L, 2L, 2L, 1L, 1L, 1L, 2L, 2L, 2L, 2L, 1L, 2L, 
2L, 2L, 1L, 2L, 1L, 1L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 1L, 
2L, 1L, 2L, 2L, 1L, 1L, 1L, 1L, 2L, 1L, 2L, 2L, 2L, 2L, 2L, 2L, 
1L, 2L, 2L, 1L, 2L, 1L, 2L, 1L, 2L, 2L, 1L, 2L, 2L, 1L, 2L, 2L, 
2L, 1L, 2L, 2L, 1L, 1L, 1L, 2L, 1L, 2L, 1L, 1L, 1L, 1L, 2L, 2L, 
2L, 1L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 1L, 2L, 1L, 2L, 1L, 1L, 2L, 
2L, 2L, 2L, 1L, 2L, 2L, 1L, 2L, 2L, 2L, 1L, 2L, 1L, 2L, 2L, 1L, 
1L, 1L, 2L, 1L, 1L, 2L, 2L, 2L, 1L, 2L, 2L, 2L, 2L, 2L, 1L, 2L, 
1L, 2L, 2L, 1L, 2L, 2L, 2L, 1L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 1L, 
1L, 1L, 2L, 1L, 2L, 2L, 1L, 2L, 1L, 1L, 2L, 2L, 2L, 2L, 2L, 2L, 
2L, 2L, 1L, 2L, 2L, 2L, 2L, 2L, 2L, 1L, 1L, 2L, 2L, 1L, 2L, 2L, 
1L, 1L, 2L, 1L, 2L, 2L, 2L, 2L, 1L, 2L, 1L, 2L, 1L, 1L, 2L, 2L, 
1L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 1L, 1L, 2L, 2L, 
2L, 2L, 2L, 2L, 1L, 1L, 2L, 1L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 
1L, 2L, 1L, 2L, 2L, 2L, 2L, 2L, 1L, 2L, 2L, 1L, 2L, 1L, 2L, 2L, 
2L, 1L, 2L, 1L, 2L, 1L, 2L, 2L, 2L, 2L, 2L, 1L, 1L, 2L, 2L, 1L, 
2L, 2L, 2L, 2L, 2L, 1L, 1L, 2L, 1L, 1L, 2L, 2L, 2L, 2L, 2L, 1L, 
2L, 2L, 2L, 2L, 2L, 1L, 2L, 2L, 2L, 2L, 1L, 2L, 2L, 1L, 2L, 2L, 
2L, 1L, 2L, 2L, 2L, 2L, 1L, 2L, 2L, 2L, 1L, 2L, 1L, 2L, 1L, 2L, 
2L, 2L, 2L, 1L, 2L, 1L, 2L, 2L, 1L, 2L, 1L, 1L, 1L, 2L, 2L, 2L, 
2L, 1L, 2L, 2L, 2L, 2L, 2L, 1L, 2L, 2L, 2L, 1L, 1L, 2L, 1L, 2L, 
1L, 2L, 2L, 2L, 2L, 2L, 1L, 2L, 1L, 2L, 2L, 2L, 1L, 2L, 2L, 1L, 
2L, 2L, 2L, 1L, 2L, 2L, 1L, 2L, 2L, 2L, 2L, 2L, 1L, 1L, 2L, 2L, 
2L, 2L, 1L, 2L, 2L, 2L, 2L, 2L, 2L, 1L, 2L, 2L, 2L, 2L, 2L, 2L, 
1L, 2L, 2L, 1L, 1L, 1L, 1L, 1L, 2L, 1L, 2L, 2L, 2L, 1L, 1L, 2L, 
1L, 1L, 2L, 2L, 2L, 2L, 1L, 2L, 2L, 1L, 1L, 2L, 2L, 2L, 1L, 1L, 
2L, 1L, 2L, 2L, 1L, 2L, 1L, 1L, 2L, 2L), .Label = c("female", 
"male"), class = "factor"), Age = c(22, 38, 26, 35, 35, NA, 54, 
2, 27, 14, 4, 58, 20, 39, 14, 55, 2, NA, 31, NA, 35, 34, 15, 
28, 8, 38, NA, 19, NA, NA, 40, NA, NA, 66, 28, 42, NA, 21, 18, 
14, 40, 27, NA, 3, 19, NA, NA, NA, NA, 18, 7, 21, 49, 29, 65, 
NA, 21, 28.5, 5, 11, 22, 38, 45, 4, NA, NA, 29, 19, 17, 26, 32, 
16, 21, 26, 32, 25, NA, NA, 0.83, 30, 22, 29, NA, 28, 17, 33, 
16, NA, 23, 24, 29, 20, 46, 26, 59, NA, 71, 23, 34, 34, 28, NA, 
21, 33, 37, 28, 21, NA, 38, NA, 47, 14.5, 22, 20, 17, 21, 70.5, 
29, 24, 2, 21, NA, 32.5, 32.5, 54, 12, NA, 24, NA, 45, 33, 20, 
47, 29, 25, 23, 19, 37, 16, 24, NA, 22, 24, 19, 18, 19, 27, 9, 
36.5, 42, 51, 22, 55.5, 40.5, NA, 51, 16, 30, NA, NA, 44, 40, 
26, 17, 1, 9, NA, 45, NA, 28, 61, 4, 1, 21, 56, 18, NA, 50, 30, 
36, NA, NA, 9, 1, 4, NA, NA, 45, 40, 36, 32, 19, 19, 3, 44, 58, 
NA, 42, NA, 24, 28, NA, 34, 45.5, 18, 2, 32, 26, 16, 40, 24, 
35, 22, 30, NA, 31, 27, 42, 32, 30, 16, 27, 51, NA, 38, 22, 19, 
20.5, 18, NA, 35, 29, 59, 5, 24, NA, 44, 8, 19, 33, NA, NA, 29, 
22, 30, 44, 25, 24, 37, 54, NA, 29, 62, 30, 41, 29, NA, 30, 35, 
50, NA, 3, 52, 40, NA, 36, 16, 25, 58, 35, NA, 25, 41, 37, NA, 
63, 45, NA, 7, 35, 65, 28, 16, 19, NA, 33, 30, 22, 42, 22, 26, 
19, 36, 24, 24, NA, 23.5, 2, NA, 50, NA, NA, 19, NA, NA, 0.92, 
NA, 17, 30, 30, 24, 18, 26, 28, 43, 26, 24, 54, 31, 40, 22, 27, 
30, 22, NA, 36, 61, 36, 31, 16, NA, 45.5, 38, 16, NA, NA, 29, 
41, 45, 45, 2, 24, 28, 25, 36, 24, 40, NA, 3, 42, 23, NA, 15, 
25, NA, 28, 22, 38, NA, NA, 40, 29, 45, 35, NA, 30, 60, NA, NA, 
24, 25, 18, 19, 22, 3, NA, 22, 27, 20, 19, 42, 1, 32, 35, NA, 
18, 1, 36, NA, 17, 36, 21, 28, 23, 24, 22, 31, 46, 23, 28, 39, 
26, 21, 28, 20, 34, 51, 3, 21, NA, NA, NA, 33, NA, 44, NA, 34, 
18, 30, 10, NA, 21, 29, 28, 18, NA, 28, 19, NA, 32, 28, NA, 42, 
17, 50, 14, 21, 24, 64, 31, 45, 20, 25, 28, NA, 4, 13, 34, 5, 
52, 36, NA, 30, 49, NA, 29, 65, NA, 50, NA, 48, 34, 47, 48, NA, 
38, NA, 56, NA, 0.75, NA, 38, 33, 23, 22, NA, 34, 29, 22, 2, 
9, NA, 50, 63, 25, NA, 35, 58, 30, 9, NA, 21, 55, 71, 21, NA, 
54, NA, 25, 24, 17, 21, NA, 37, 16, 18, 33, NA, 28, 26, 29, NA, 
36, 54, 24, 47, 34, NA, 36, 32, 30, 22, NA, 44, NA, 40.5, 50, 
NA, 39, 23, 2, NA, 17, NA, 30, 7, 45, 30, NA, 22, 36, 9, 11, 
32, 50, 64, 19, NA, 33, 8, 17, 27, NA, 22, 22, 62, 48, NA, 39, 
36, NA, 40, 28, NA, NA, 24, 19, 29, NA, 32, 62, 53, 36, NA, 16, 
19, 34, 39, NA, 32, 25, 39, 54, 36, NA, 18, 47, 60, 22, NA, 35, 
52, 47, NA, 37, 36, NA, 49, NA, 49, 24, NA, NA, 44, 35, 36, 30, 
27, 22, 40, 39, NA, NA, NA, 35, 24, 34, 26, 4, 26, 27, 42, 20, 
21, 21, 61, 57, 21, 26, NA, 80, 51, 32, NA, 9, 28, 32, 31, 41, 
NA, 20, 24, 2, NA, 0.75, 48, 19, 56, NA, 23, NA, 18, 21, NA, 
18, 24, NA, 32, 23, 58, 50, 40, 47, 36, 20, 32, 25, NA, 43, NA, 
40, 31, 70, 31, NA, 18, 24.5, 18, 43, 36, NA, 27, 20, 14, 60, 
25, 14, 19, 18, 15, 31, 4, NA, 25, 60, 52, 44, NA, 49, 42, 18, 
35, 18, 25, 26, 39, 45, 42, 22, NA, 24, NA, 48, 29, 52, 19, 38, 
27, NA, 33, 6, 17, 34, 50, 27, 20, 30, NA, 25, 25, 29, 11, NA, 
23, 23, 28.5, 48, 35, NA, NA, NA, 36, 21, 24, 31, 70, 16, 30, 
19, 31, 4, 6, 33, 23, 48, 0.67, 28, 18, 34, 33, NA, 41, 20, 36, 
16, 51, NA, 30.5, NA, 32, 24, 48, 57, NA, 54, 18, NA, 5, NA, 
43, 13, 17, 29, NA, 25, 25, 18, 8, 1, 46, NA, 16, NA, NA, 25, 
39, 49, 31, 30, 30, 34, 31, 11, 0.42, 27, 31, 39, 18, 39, 33, 
26, 39, 35, 6, 30.5, NA, 23, 31, 43, 10, 52, 27, 38, 27, 2, NA, 
NA, 1, NA, 62, 15, 0.83, NA, 23, 18, 39, 21, NA, 32, NA, 20, 
16, 30, 34.5, 17, 42, NA, 35, 28, NA, 4, 74, 9, 16, 44, 18, 45, 
51, 24, NA, 41, 21, 48, NA, 24, 42, 27, 31, NA, 4, 26, 47, 33, 
47, 28, 15, 20, 19, NA, 56, 25, 33, 22, 28, 25, 39, 27, 19, NA, 
26, 32), SibSp = c(1L, 1L, 0L, 1L, 0L, 0L, 0L, 3L, 0L, 1L, 1L, 
0L, 0L, 1L, 0L, 0L, 4L, 0L, 1L, 0L, 0L, 0L, 0L, 0L, 3L, 1L, 0L, 
3L, 0L, 0L, 0L, 1L, 0L, 0L, 1L, 1L, 0L, 0L, 2L, 1L, 1L, 1L, 0L, 
1L, 0L, 0L, 1L, 0L, 2L, 1L, 4L, 0L, 1L, 1L, 0L, 0L, 0L, 0L, 1L, 
5L, 0L, 0L, 1L, 3L, 0L, 1L, 0L, 0L, 4L, 2L, 0L, 5L, 0L, 1L, 0L, 
0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 3L, 1L, 0L, 3L, 0L, 0L, 
0L, 1L, 1L, 0L, 0L, 0L, 0L, 0L, 1L, 0L, 0L, 0L, 0L, 2L, 0L, 0L, 
0L, 0L, 1L, 0L, 1L, 0L, 1L, 0L, 0L, 0L, 1L, 0L, 4L, 2L, 0L, 1L, 
0L, 0L, 1L, 0L, 0L, 1L, 0L, 0L, 0L, 1L, 1L, 0L, 0L, 0L, 1L, 0L, 
0L, 0L, 0L, 1L, 0L, 0L, 1L, 0L, 2L, 0L, 0L, 0L, 1L, 0L, 0L, 0L, 
0L, 0L, 0L, 0L, 8L, 0L, 0L, 0L, 0L, 4L, 0L, 0L, 1L, 0L, 0L, 0L, 
4L, 1L, 0L, 0L, 1L, 3L, 0L, 0L, 0L, 8L, 0L, 4L, 2L, 0L, 0L, 1L, 
0L, 1L, 0L, 0L, 0L, 1L, 1L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 8L, 0L, 
0L, 0L, 0L, 1L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 1L, 1L, 0L, 1L, 0L, 
0L, 0L, 0L, 0L, 0L, 1L, 0L, 0L, 0L, 0L, 3L, 1L, 0L, 0L, 4L, 0L, 
0L, 1L, 0L, 0L, 0L, 1L, 1L, 0L, 0L, 0L, 2L, 0L, 0L, 1L, 1L, 0L, 
1L, 0L, 1L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 4L, 1L, 0L, 0L, 0L, 4L, 
1L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 1L, 0L, 0L, 4L, 1L, 0L, 0L, 0L, 
0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 1L, 0L, 0L, 0L, 0L, 0L, 1L, 0L, 
0L, 0L, 2L, 0L, 0L, 0L, 1L, 0L, 1L, 1L, 0L, 0L, 2L, 1L, 0L, 1L, 
0L, 1L, 0L, 0L, 1L, 0L, 0L, 0L, 1L, 8L, 0L, 0L, 0L, 1L, 0L, 2L, 
0L, 0L, 2L, 1L, 0L, 1L, 0L, 0L, 0L, 1L, 3L, 0L, 0L, 0L, 0L, 0L, 
1L, 1L, 0L, 0L, 0L, 1L, 1L, 0L, 0L, 0L, 0L, 0L, 0L, 1L, 1L, 0L, 
0L, 1L, 0L, 1L, 0L, 0L, 0L, 1L, 1L, 0L, 0L, 3L, 1L, 0L, 0L, 0L, 
0L, 0L, 0L, 0L, 1L, 0L, 0L, 5L, 0L, 0L, 0L, 1L, 0L, 2L, 1L, 0L, 
0L, 0L, 0L, 0L, 0L, 0L, 0L, 1L, 1L, 0L, 1L, 0L, 1L, 0L, 3L, 0L, 
0L, 1L, 0L, 0L, 0L, 1L, 0L, 0L, 0L, 0L, 0L, 0L, 1L, 1L, 0L, 1L, 
0L, 0L, 0L, 0L, 1L, 1L, 0L, 1L, 1L, 2L, 2L, 1L, 0L, 1L, 0L, 1L, 
0L, 0L, 0L, 0L, 0L, 2L, 0L, 1L, 1L, 0L, 1L, 0L, 0L, 0L, 1L, 0L, 
0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 2L, 0L, 0L, 1L, 0L, 0L, 
0L, 1L, 1L, 0L, 0L, 5L, 0L, 0L, 0L, 1L, 3L, 1L, 0L, 0L, 1L, 1L, 
0L, 0L, 0L, 0L, 0L, 1L, 0L, 1L, 0L, 0L, 0L, 0L, 0L, 0L, 1L, 0L, 
0L, 0L, 0L, 0L, 0L, 0L, 1L, 0L, 0L, 0L, 0L, 1L, 0L, 0L, 0L, 0L, 
0L, 0L, 0L, 0L, 0L, 0L, 2L, 1L, 0L, 1L, 0L, 0L, 0L, 0L, 0L, 0L, 
0L, 0L, 4L, 4L, 1L, 1L, 0L, 1L, 0L, 1L, 1L, 0L, 0L, 0L, 0L, 0L, 
0L, 1L, 0L, 1L, 1L, 0L, 0L, 0L, 0L, 0L, 2L, 0L, 0L, 0L, 0L, 0L, 
2L, 0L, 0L, 0L, 0L, 0L, 1L, 1L, 0L, 1L, 1L, 0L, 0L, 0L, 0L, 0L, 
1L, 0L, 0L, 0L, 1L, 0L, 0L, 1L, 1L, 0L, 0L, 0L, 1L, 2L, 0L, 0L, 
0L, 0L, 1L, 0L, 0L, 1L, 0L, 1L, 0L, 1L, 0L, 0L, 1L, 1L, 1L, 2L, 
0L, 1L, 1L, 1L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 3L, 
0L, 0L, 1L, 0L, 1L, 0L, 0L, 3L, 0L, 2L, 1L, 0L, 0L, 0L, 0L, 0L, 
0L, 0L, 0L, 0L, 2L, 0L, 1L, 0L, 0L, 2L, 0L, 0L, 0L, 1L, 2L, 0L, 
0L, 0L, 1L, 1L, 1L, 0L, 0L, 0L, 0L, 0L, 0L, 1L, 0L, 0L, 0L, 0L, 
5L, 1L, 1L, 4L, 0L, 0L, 0L, 1L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 1L, 
0L, 1L, 0L, 0L, 0L, 1L, 0L, 0L, 0L, 0L, 1L, 0L, 0L, 1L, 0L, 0L, 
0L, 0L, 0L, 0L, 0L, 0L, 1L, 0L, 0L, 1L, 0L, 3L, 0L, 1L, 1L, 0L, 
0L, 0L, 0L, 0L, 0L, 1L, 0L, 0L, 0L, 0L, 1L, 2L, 1L, 0L, 1L, 1L, 
0L, 1L, 0L, 1L, 0L, 0L, 0L, 1L, 1L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 
1L, 0L, 1L, 0L, 0L, 1L, 0L, 0L, 0L, 0L, 0L, 1L, 0L, 0L, 0L, 0L, 
0L, 0L, 1L, 0L, 1L, 0L, 0L, 0L, 4L, 1L, 0L, 0L, 0L, 8L, 0L, 0L, 
0L, 0L, 0L, 0L, 1L, 0L, 1L, 1L, 0L, 0L, 0L, 0L, 0L, 0L, 1L, 0L, 
0L, 0L, 4L, 0L, 0L, 0L, 1L, 0L, 3L, 1L, 0L, 0L, 0L, 4L, 0L, 0L, 
0L, 0L, 0L, 1L, 1L, 0L, 0L, 0L, 1L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 
0L, 0L, 0L, 8L, 0L, 0L, 1L, 4L, 0L, 1L, 0L, 1L, 0L, 1L, 0L, 0L, 
0L, 2L, 1L, 0L, 8L, 0L, 0L, 1L, 0L, 0L, 1L, 0L, 1L, 0L, 0L, 1L, 
0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 1L, 0L, 0L
), Parch = c(0L, 0L, 0L, 0L, 0L, 0L, 0L, 1L, 2L, 0L, 1L, 0L, 
0L, 5L, 0L, 0L, 1L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 1L, 5L, 0L, 2L, 
0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 2L, 
0L, 0L, 0L, 0L, 0L, 0L, 1L, 0L, 0L, 0L, 1L, 0L, 0L, 0L, 2L, 2L, 
0L, 0L, 0L, 2L, 0L, 1L, 0L, 0L, 2L, 0L, 0L, 2L, 0L, 0L, 0L, 0L, 
0L, 0L, 2L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 3L, 0L, 2L, 0L, 0L, 0L, 
0L, 2L, 0L, 0L, 0L, 1L, 1L, 0L, 0L, 0L, 1L, 0L, 0L, 0L, 0L, 0L, 
0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 1L, 2L, 0L, 0L, 0L, 0L, 
1L, 0L, 0L, 0L, 1L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 2L, 0L, 0L, 0L, 
2L, 0L, 0L, 0L, 0L, 1L, 0L, 2L, 2L, 0L, 0L, 0L, 0L, 2L, 0L, 1L, 
0L, 0L, 0L, 2L, 1L, 0L, 0L, 0L, 1L, 2L, 1L, 4L, 0L, 0L, 0L, 1L, 
1L, 0L, 0L, 1L, 1L, 0L, 0L, 0L, 2L, 0L, 2L, 1L, 2L, 0L, 0L, 0L, 
1L, 0L, 0L, 0L, 0L, 1L, 0L, 0L, 0L, 1L, 0L, 0L, 0L, 2L, 0L, 0L, 
0L, 1L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 
0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 1L, 0L, 0L, 0L, 2L, 0L, 0L, 
0L, 2L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 2L, 1L, 0L, 0L, 1L, 
0L, 0L, 2L, 2L, 0L, 0L, 0L, 1L, 0L, 2L, 1L, 0L, 0L, 0L, 1L, 0L, 
1L, 0L, 0L, 0L, 1L, 1L, 0L, 0L, 0L, 0L, 1L, 1L, 0L, 0L, 0L, 0L, 
0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 2L, 0L, 1L, 
0L, 0L, 0L, 0L, 0L, 2L, 0L, 0L, 0L, 0L, 0L, 2L, 1L, 0L, 1L, 0L, 
0L, 0L, 2L, 1L, 0L, 0L, 0L, 1L, 2L, 0L, 0L, 0L, 1L, 1L, 0L, 0L, 
1L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 1L, 2L, 0L, 0L, 0L, 0L, 0L, 0L, 
1L, 0L, 0L, 0L, 1L, 0L, 0L, 0L, 1L, 0L, 0L, 0L, 4L, 0L, 1L, 0L, 
0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 1L, 0L, 0L, 2L, 0L, 0L, 
0L, 2L, 0L, 0L, 0L, 0L, 2L, 0L, 0L, 0L, 2L, 0L, 0L, 0L, 2L, 0L, 
0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 1L, 0L, 1L, 0L, 0L, 
0L, 0L, 0L, 0L, 1L, 2L, 0L, 2L, 0L, 0L, 0L, 1L, 1L, 0L, 0L, 0L, 
0L, 0L, 0L, 0L, 0L, 0L, 0L, 2L, 2L, 3L, 4L, 0L, 1L, 0L, 0L, 0L, 
0L, 2L, 1L, 0L, 1L, 0L, 2L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 
0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 1L, 0L, 0L, 2L, 0L, 0L, 0L, 
0L, 0L, 0L, 1L, 2L, 0L, 0L, 0L, 0L, 1L, 0L, 0L, 0L, 1L, 0L, 0L, 
0L, 0L, 0L, 0L, 0L, 0L, 2L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 2L, 0L, 
0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 1L, 
0L, 0L, 0L, 0L, 0L, 1L, 1L, 0L, 1L, 2L, 0L, 2L, 0L, 0L, 0L, 2L, 
2L, 2L, 2L, 0L, 0L, 0L, 0L, 0L, 1L, 1L, 2L, 0L, 0L, 0L, 0L, 0L, 
0L, 0L, 1L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 4L, 0L, 0L, 0L, 0L, 
0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 1L, 1L, 0L, 0L, 0L, 2L, 0L, 1L, 
0L, 0L, 0L, 0L, 0L, 2L, 0L, 1L, 0L, 0L, 0L, 0L, 1L, 0L, 0L, 0L, 
0L, 0L, 0L, 0L, 2L, 0L, 5L, 0L, 0L, 0L, 0L, 2L, 1L, 0L, 1L, 0L, 
0L, 0L, 1L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 2L, 0L, 
0L, 1L, 5L, 0L, 0L, 0L, 2L, 0L, 1L, 0L, 0L, 0L, 0L, 0L, 0L, 1L, 
0L, 0L, 0L, 0L, 0L, 1L, 0L, 2L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 
0L, 0L, 1L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 6L, 1L, 0L, 0L, 0L, 2L, 
1L, 2L, 1L, 0L, 0L, 1L, 0L, 1L, 0L, 0L, 0L, 0L, 0L, 0L, 1L, 0L, 
0L, 0L, 1L, 0L, 0L, 0L, 0L, 0L, 0L, 1L, 0L, 0L, 0L, 0L, 0L, 0L, 
0L, 0L, 0L, 0L, 1L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 
0L, 0L, 0L, 0L, 3L, 0L, 0L, 0L, 0L, 0L, 2L, 0L, 0L, 1L, 1L, 0L, 
0L, 0L, 1L, 1L, 0L, 0L, 2L, 1L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 2L, 
0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 3L, 0L, 0L, 0L, 0L, 1L, 
0L, 0L, 0L, 2L, 0L, 0L, 0L, 1L, 2L, 0L, 0L, 0L, 2L, 0L, 0L, 0L, 
0L, 0L, 0L, 1L, 0L, 1L, 2L, 1L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 
0L, 2L, 0L, 0L, 0L, 1L, 0L, 2L, 1L, 0L, 0L, 1L, 1L, 0L, 0L, 2L, 
0L, 0L, 0L, 1L, 0L, 0L, 0L, 1L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 
0L, 0L, 2L, 0L, 1L, 0L, 2L, 0L, 1L, 1L, 0L, 1L, 1L, 0L, 3L, 0L, 
0L, 0L, 0L, 2L, 0L, 0L, 0L, 0L, 0L, 1L, 0L, 1L, 0L, 0L, 0L, 0L, 
0L, 0L, 0L, 1L, 1L, 0L, 0L, 0L, 0L, 5L, 0L, 0L, 2L, 0L, 0L), 
    Ticket = structure(c(524L, 597L, 670L, 50L, 473L, 276L, 86L, 
    396L, 345L, 133L, 617L, 39L, 536L, 334L, 414L, 154L, 481L, 
    152L, 302L, 185L, 140L, 153L, 279L, 43L, 396L, 330L, 180L, 
    96L, 284L, 363L, 599L, 587L, 289L, 550L, 601L, 44L, 203L, 
    515L, 303L, 187L, 508L, 54L, 392L, 635L, 283L, 618L, 463L, 
    77L, 191L, 378L, 250L, 523L, 588L, 240L, 30L, 95L, 556L, 
    216L, 562L, 567L, 198L, 33L, 458L, 338L, 602L, 190L, 554L, 
    624L, 249L, 271L, 558L, 567L, 622L, 205L, 81L, 348L, 356L, 
    474L, 159L, 433L, 305L, 312L, 281L, 26L, 638L, 248L, 677L, 
    653L, 96L, 298L, 299L, 342L, 679L, 549L, 429L, 476L, 608L, 
    613L, 118L, 151L, 385L, 362L, 418L, 506L, 246L, 355L, 297L, 
    255L, 389L, 470L, 3L, 194L, 275L, 492L, 177L, 667L, 461L, 
    54L, 586L, 334L, 622L, 541L, 133L, 220L, 418L, 187L, 464L, 
    542L, 197L, 319L, 381L, 641L, 538L, 112L, 553L, 631L, 56L, 
    50L, 505L, 593L, 204L, 333L, 669L, 439L, 119L, 559L, 406L, 
    677L, 114L, 148L, 623L, 37L, 517L, 539L, 574L, 596L, 419L, 
    655L, 259L, 569L, 471L, 560L, 324L, 266L, 250L, 424L, 29L, 
    338L, 577L, 81L, 6L, 481L, 345L, 660L, 91L, 413L, 489L, 594L, 
    170L, 575L, 569L, 630L, 330L, 115L, 272L, 35L, 460L, 12L, 
    436L, 387L, 123L, 228L, 407L, 114L, 605L, 587L, 453L, 495L, 
    462L, 161L, 307L, 569L, 243L, 178L, 531L, 317L, 248L, 217L, 
    446L, 18L, 643L, 571L, 527L, 164L, 444L, 417L, 671L, 143L, 
    64L, 680L, 654L, 106L, 103L, 376L, 94L, 616L, 673L, 526L, 
    126L, 489L, 458L, 323L, 129L, 330L, 555L, 678L, 199L, 557L, 
    232L, 637L, 194L, 445L, 675L, 659L, 214L, 93L, 326L, 167L, 
    55L, 145L, 423L, 317L, 32L, 537L, 459L, 186L, 590L, 1L, 609L, 
    116L, 483L, 330L, 2L, 17L, 479L, 547L, 250L, 335L, 589L, 
    614L, 48L, 575L, 163L, 595L, 466L, 72L, 327L, 136L, 481L, 
    552L, 291L, 341L, 311L, 534L, 25L, 379L, 309L, 354L, 135L, 
    465L, 92L, 65L, 636L, 377L, 375L, 607L, 213L, 38L, 99L, 586L, 
    514L, 442L, 575L, 110L, 528L, 38L, 84L, 612L, 576L, 585L, 
    60L, 603L, 168L, 383L, 572L, 344L, 151L, 233L, 454L, 82L, 
    525L, 366L, 125L, 159L, 569L, 614L, 300L, 230L, 424L, 8L, 
    442L, 21L, 589L, 303L, 606L, 371L, 37L, 82L, 511L, 40L, 114L, 
    96L, 160L, 150L, 113L, 158L, 258L, 484L, 563L, 263L, 504L, 
    31L, 215L, 378L, 183L, 315L, 29L, 132L, 280L, 286L, 338L, 
    634L, 212L, 642L, 460L, 545L, 5L, 176L, 79L, 582L, 59L, 245L, 
    274L, 614L, 396L, 601L, 546L, 28L, 184L, 325L, 611L, 188L, 
    666L, 44L, 372L, 622L, 567L, 221L, 448L, 626L, 34L, 402L, 
    247L, 417L, 617L, 411L, 415L, 227L, 147L, 141L, 664L, 293L, 
    493L, 669L, 268L, 231L, 322L, 238L, 256L, 489L, 368L, 488L, 
    93L, 136L, 656L, 296L, 225L, 169L, 226L, 308L, 393L, 535L, 
    260L, 332L, 459L, 522L, 101L, 171L, 438L, 645L, 4L, 477L, 
    628L, 658L, 73L, 34L, 677L, 238L, 96L, 548L, 572L, 306L, 
    329L, 117L, 500L, 290L, 163L, 46L, 195L, 41L, 562L, 498L, 
    23L, 85L, 529L, 380L, 74L, 87L, 573L, 469L, 97L, 430L, 7L, 
    122L, 540L, 640L, 136L, 45L, 422L, 195L, 273L, 264L, 562L, 
    629L, 510L, 3L, 253L, 316L, 412L, 252L, 567L, 137L, 532L, 
    490L, 65L, 489L, 94L, 62L, 518L, 563L, 499L, 650L, 42L, 604L, 
    519L, 177L, 455L, 565L, 38L, 403L, 262L, 434L, 277L, 491L, 
    1L, 612L, 181L, 11L, 543L, 81L, 480L, 649L, 578L, 600L, 357L, 
    457L, 561L, 470L, 111L, 382L, 68L, 391L, 174L, 8L, 218L, 
    447L, 674L, 584L, 251L, 236L, 181L, 182L, 211L, 197L, 261L, 
    572L, 22L, 615L, 427L, 76L, 681L, 334L, 334L, 234L, 615L, 
    502L, 234L, 632L, 424L, 559L, 84L, 149L, 285L, 172L, 336L, 
    53L, 58L, 611L, 2L, 301L, 472L, 390L, 104L, 652L, 646L, 520L, 
    353L, 396L, 208L, 416L, 625L, 61L, 579L, 78L, 521L, 421L, 
    144L, 73L, 210L, 662L, 134L, 84L, 227L, 69L, 292L, 2L, 130L, 
    75L, 80L, 516L, 657L, 455L, 533L, 435L, 627L, 308L, 156L, 
    575L, 193L, 585L, 143L, 361L, 47L, 431L, 10L, 397L, 386L, 
    51L, 635L, 589L, 334L, 639L, 445L, 468L, 432L, 107L, 332L, 
    537L, 115L, 254L, 189L, 57L, 188L, 401L, 496L, 456L, 105L, 
    72L, 370L, 287L, 219L, 346L, 71L, 14L, 338L, 131L, 665L, 
    557L, 250L, 477L, 410L, 582L, 338L, 81L, 195L, 588L, 374L, 
    70L, 621L, 568L, 367L, 118L, 513L, 278L, 440L, 622L, 369L, 
    436L, 242L, 417L, 606L, 173L, 497L, 358L, 661L, 622L, 124L, 
    257L, 530L, 100L, 241L, 570L, 551L, 146L, 139L, 398L, 295L, 
    494L, 567L, 609L, 282L, 588L, 501L, 567L, 241L, 635L, 250L, 
    373L, 404L, 142L, 90L, 394L, 81L, 201L, 49L, 157L, 426L, 
    420L, 84L, 347L, 611L, 580L, 212L, 450L, 399L, 171L, 109L, 
    581L, 38L, 190L, 583L, 20L, 100L, 507L, 165L, 349L, 611L, 
    294L, 441L, 320L, 156L, 409L, 66L, 162L, 52L, 267L, 253L, 
    452L, 128L, 668L, 142L, 217L, 138L, 229L, 120L, 496L, 677L, 
    609L, 350L, 365L, 83L, 92L, 603L, 478L, 663L, 681L, 552L, 
    166L, 36L, 288L, 235L, 487L, 313L, 352L, 107L, 167L, 405L, 
    239L, 425L, 1L, 421L, 647L, 192L, 34L, 328L, 72L, 19L, 437L, 
    470L, 512L, 314L, 408L, 620L, 202L, 237L, 331L, 482L, 433L, 
    451L, 142L, 209L, 90L, 27L, 676L, 644L, 475L, 244L, 481L, 
    549L, 593L, 67L, 140L, 569L, 598L, 351L, 224L, 88L, 384L, 
    207L, 308L, 165L, 557L, 34L, 175L, 339L, 321L, 13L, 337L, 
    155L, 52L, 340L, 520L, 223L, 334L, 428L, 16L, 672L, 619L, 
    544L, 338L, 68L, 270L, 98L, 487L, 250L, 449L, 81L, 619L, 
    443L, 33L, 189L, 238L, 200L, 343L, 108L, 610L, 269L, 486L, 
    81L, 63L, 648L, 620L, 48L, 206L, 265L, 564L, 569L, 360L, 
    156L, 85L, 334L, 318L, 204L, 592L, 145L, 485L, 454L, 24L, 
    195L, 179L, 400L, 222L, 89L, 569L, 121L, 127L, 633L, 591L, 
    310L, 345L, 388L, 55L, 503L, 304L, 576L, 196L, 505L, 359L, 
    364L, 60L, 116L, 395L, 509L, 566L, 651L, 481L, 102L, 15L, 
    676L, 9L, 467L), .Label = c("110152", "110413", "110465", 
    "110564", "110813", "111240", "111320", "111361", "111369", 
    "111426", "111427", "111428", "112050", "112052", "112053", 
    "112058", "112059", "112277", "112379", "113028", "113043", 
    "113050", "113051", "113055", "113056", "113059", "113501", 
    "113503", "113505", "113509", "113510", "113514", "113572", 
    "113760", "113767", "113773", "113776", "113781", "113783", 
    "113784", "113786", "113787", "113788", "113789", "113792", 
    "113794", "113796", "113798", "113800", "113803", "113804", 
    "113806", "113807", "11668", "11751", "11752", "11753", "11755", 
    "11765", "11767", "11769", "11771", "11774", "11813", "11967", 
    "12233", "12460", "12749", "13049", "13213", "13214", "13502", 
    "13507", "13509", "13567", "13568", "14311", "14312", "14313", 
    "14973", "1601", "16966", "16988", "17421", "17453", "17463", 
    "17464", "17465", "17466", "17474", "17764", "19877", "19928", 
    "19943", "19947", "19950", "19952", "19972", "19988", "19996", 
    "2003", "211536", "21440", "218629", "219533", "220367", 
    "220845", "2223", "223596", "226593", "226875", "228414", 
    "229236", "230080", "230136", "230433", "230434", "231919", 
    "231945", "233639", "233866", "234360", "234604", "234686", 
    "234818", "236171", "236852", "236853", "237442", "237565", 
    "237668", "237671", "237736", "237789", "237798", "239853", 
    "239854", "239855", "239856", "239865", "240929", "24160", 
    "243847", "243880", "244252", "244270", "244278", "244310", 
    "244358", "244361", "244367", "244373", "248698", "248706", 
    "248723", "248727", "248731", "248733", "248738", "248740", 
    "248747", "250643", "250644", "250646", "250647", "250648", 
    "250649", "250651", "250652", "250653", "250655", "2620", 
    "2623", "2624", "2625", "2626", "2627", "2628", "2629", "2631", 
    "26360", "2641", "2647", "2648", "2649", "2650", "2651", 
    "2653", "2659", "2661", "2662", "2663", "2664", "2665", "2666", 
    "2667", "2668", "2669", "26707", "2671", "2672", "2674", 
    "2677", "2678", "2680", "2683", "2685", "2686", "2687", "2689", 
    "2690", "2691", "2693", "2694", "2695", "2697", "2699", "2700", 
    "27042", "27267", "27849", "28134", "28206", "28213", "28220", 
    "28228", "28403", "28424", "28425", "28551", "28664", "28665", 
    "29011", "2908", "29103", "29104", "29105", "29106", "29108", 
    "2926", "29750", "29751", "3101264", "3101265", "3101267", 
    "3101276", "3101277", "3101278", "3101281", "3101295", "3101296", 
    "3101298", "31027", "31028", "312991", "312992", "312993", 
    "31418", "315037", "315082", "315084", "315086", "315088", 
    "315089", "315090", "315093", "315094", "315096", "315097", 
    "315098", "315151", "315153", "323592", "323951", "324669", 
    "330877", "330909", "330919", "330923", "330931", "330932", 
    "330935", "330958", "330959", "330979", "330980", "334912", 
    "335097", "335677", "33638", "336439", "3411", "341826", 
    "34218", "342826", "343095", "343120", "343275", "343276", 
    "345364", "345572", "345763", "345764", "345765", "345767", 
    "345769", "345770", "345773", "345774", "345777", "345778", 
    "345779", "345780", "345781", "345783", "3460", "347054", 
    "347060", "347061", "347062", "347063", "347064", "347067", 
    "347068", "347069", "347071", "347073", "347074", "347076", 
    "347077", "347078", "347080", "347081", "347082", "347083", 
    "347085", "347087", "347088", "347089", "3474", "347464", 
    "347466", "347468", "347470", "347742", "347743", "348121", 
    "348123", "348124", "349201", "349203", "349204", "349205", 
    "349206", "349207", "349208", "349209", "349210", "349212", 
    "349213", "349214", "349215", "349216", "349217", "349218", 
    "349219", "349221", "349222", "349223", "349224", "349225", 
    "349227", "349228", "349231", "349233", "349234", "349236", 
    "349237", "349239", "349240", "349241", "349242", "349243", 
    "349244", "349245", "349246", "349247", "349248", "349249", 
    "349251", "349252", "349253", "349254", "349256", "349257", 
    "349909", "349910", "349912", "350025", "350026", "350029", 
    "350034", "350035", "350036", "350042", "350043", "350046", 
    "350047", "350048", "350050", "350052", "350060", "350404", 
    "350406", "350407", "350417", "35273", "35281", "35851", 
    "35852", "358585", "36209", "362316", "363291", "363294", 
    "363592", "364498", "364499", "364500", "364506", "364511", 
    "364512", "364516", "364846", "364848", "364849", "364850", 
    "364851", "365222", "365226", "36568", "367226", "367228", 
    "367229", "367230", "367231", "367232", "367655", "368323", 
    "36864", "36865", "36866", "368703", "36928", "36947", "36963", 
    "36967", "36973", "370129", "370365", "370369", "370370", 
    "370371", "370372", "370373", "370375", "370376", "370377", 
    "371060", "371110", "371362", "372622", "373450", "374746", 
    "374887", "374910", "376564", "376566", "382649", "382651", 
    "382652", "383121", "384461", "386525", "392091", "392092", 
    "392096", "394140", "4133", "4134", "4135", "4136", "4137", 
    "4138", "4579", "54636", "5727", "65303", "65304", "65306", 
    "6563", "693", "695", "7267", "7534", "7540", "7545", "7546", 
    "7552", "7553", "7598", "8471", "8475", "9234", "A./5. 2152", 
    "A./5. 3235", "A.5. 11206", "A.5. 18509", "A/4 45380", "A/4 48871", 
    "A/4. 20589", "A/4. 34244", "A/4. 39886", "A/5 21171", "A/5 21172", 
    "A/5 21173", "A/5 21174", "A/5 2466", "A/5 2817", "A/5 3536", 
    "A/5 3540", "A/5 3594", "A/5 3902", "A/5. 10482", "A/5. 13032", 
    "A/5. 2151", "A/5. 3336", "A/5. 3337", "A/5. 851", "A/S 2816", 
    "A4. 54510", "C 17369", "C 4001", "C 7075", "C 7076", "C 7077", 
    "C.A. 17248", "C.A. 18723", "C.A. 2315", "C.A. 24579", "C.A. 24580", 
    "C.A. 2673", "C.A. 29178", "C.A. 29395", "C.A. 29566", "C.A. 31026", 
    "C.A. 31921", "C.A. 33111", "C.A. 33112", "C.A. 33595", "C.A. 34260", 
    "C.A. 34651", "C.A. 37671", "C.A. 5547", "C.A. 6212", "C.A./SOTON 34068", 
    "CA 2144", "CA. 2314", "CA. 2343", "F.C. 12750", "F.C.C. 13528", 
    "F.C.C. 13529", "F.C.C. 13531", "Fa 265302", "LINE", "P/PP 3381", 
    "PC 17318", "PC 17473", "PC 17474", "PC 17475", "PC 17476", 
    "PC 17477", "PC 17482", "PC 17483", "PC 17485", "PC 17558", 
    "PC 17569", "PC 17572", "PC 17582", "PC 17585", "PC 17590", 
    "PC 17592", "PC 17593", "PC 17595", "PC 17596", "PC 17597", 
    "PC 17599", "PC 17600", "PC 17601", "PC 17603", "PC 17604", 
    "PC 17605", "PC 17608", "PC 17609", "PC 17610", "PC 17611", 
    "PC 17612", "PC 17754", "PC 17755", "PC 17756", "PC 17757", 
    "PC 17758", "PC 17759", "PC 17760", "PC 17761", "PP 4348", 
    "PP 9549", "S.C./A.4. 23567", "S.C./PARIS 2079", "S.O./P.P. 3", 
    "S.O./P.P. 751", "S.O.C. 14879", "S.O.P. 1166", "S.P. 3464", 
    "S.W./PP 752", "SC 1748", "SC/AH 29037", "SC/AH 3085", "SC/AH Basle 541", 
    "SC/PARIS 2131", "SC/PARIS 2133", "SC/PARIS 2146", "SC/PARIS 2149", 
    "SC/PARIS 2167", "SC/Paris 2123", "SC/Paris 2163", "SCO/W 1585", 
    "SO/C 14885", "SOTON/O.Q. 3101305", "SOTON/O.Q. 3101306", 
    "SOTON/O.Q. 3101307", "SOTON/O.Q. 3101310", "SOTON/O.Q. 3101311", 
    "SOTON/O.Q. 3101312", "SOTON/O.Q. 392078", "SOTON/O.Q. 392087", 
    "SOTON/O2 3101272", "SOTON/O2 3101287", "SOTON/OQ 3101316", 
    "SOTON/OQ 3101317", "SOTON/OQ 392076", "SOTON/OQ 392082", 
    "SOTON/OQ 392086", "SOTON/OQ 392089", "SOTON/OQ 392090", 
    "STON/O 2. 3101269", "STON/O 2. 3101273", "STON/O 2. 3101274", 
    "STON/O 2. 3101275", "STON/O 2. 3101280", "STON/O 2. 3101285", 
    "STON/O 2. 3101286", "STON/O 2. 3101288", "STON/O 2. 3101289", 
    "STON/O 2. 3101292", "STON/O 2. 3101293", "STON/O 2. 3101294", 
    "STON/O2. 3101271", "STON/O2. 3101279", "STON/O2. 3101282", 
    "STON/O2. 3101283", "STON/O2. 3101290", "SW/PP 751", "W./C. 14258", 
    "W./C. 14263", "W./C. 6607", "W./C. 6608", "W./C. 6609", 
    "W.E.P. 5734", "W/C 14208", "WE/P 5735"), class = "factor"), 
    Fare = c(7.25, 71.2833, 7.925, 53.1, 8.05, 8.4583, 51.8625, 
    21.075, 11.1333, 30.0708, 16.7, 26.55, 8.05, 31.275, 7.8542, 
    16, 29.125, 13, 18, 7.225, 26, 13, 8.0292, 35.5, 21.075, 
    31.3875, 7.225, 263, 7.8792, 7.8958, 27.7208, 146.5208, 7.75, 
    10.5, 82.1708, 52, 7.2292, 8.05, 18, 11.2417, 9.475, 21, 
    7.8958, 41.5792, 7.8792, 8.05, 15.5, 7.75, 21.6792, 17.8, 
    39.6875, 7.8, 76.7292, 26, 61.9792, 35.5, 10.5, 7.2292, 27.75, 
    46.9, 7.2292, 80, 83.475, 27.9, 27.7208, 15.2458, 10.5, 8.1583, 
    7.925, 8.6625, 10.5, 46.9, 73.5, 14.4542, 56.4958, 7.65, 
    7.8958, 8.05, 29, 12.475, 9, 9.5, 7.7875, 47.1, 10.5, 15.85, 
    34.375, 8.05, 263, 8.05, 8.05, 7.8542, 61.175, 20.575, 7.25, 
    8.05, 34.6542, 63.3583, 23, 26, 7.8958, 7.8958, 77.2875, 
    8.6542, 7.925, 7.8958, 7.65, 7.775, 7.8958, 24.15, 52, 14.4542, 
    8.05, 9.825, 14.4583, 7.925, 7.75, 21, 247.5208, 31.275, 
    73.5, 8.05, 30.0708, 13, 77.2875, 11.2417, 7.75, 7.1417, 
    22.3583, 6.975, 7.8958, 7.05, 14.5, 26, 13, 15.0458, 26.2833, 
    53.1, 9.2167, 79.2, 15.2458, 7.75, 15.85, 6.75, 11.5, 36.75, 
    7.7958, 34.375, 26, 13, 12.525, 66.6, 8.05, 14.5, 7.3125, 
    61.3792, 7.7333, 8.05, 8.6625, 69.55, 16.1, 15.75, 7.775, 
    8.6625, 39.6875, 20.525, 55, 27.9, 25.925, 56.4958, 33.5, 
    29.125, 11.1333, 7.925, 30.6958, 7.8542, 25.4667, 28.7125, 
    13, 0, 69.55, 15.05, 31.3875, 39, 22.025, 50, 15.5, 26.55, 
    15.5, 7.8958, 13, 13, 7.8542, 26, 27.7208, 146.5208, 7.75, 
    8.4042, 7.75, 13, 9.5, 69.55, 6.4958, 7.225, 8.05, 10.4625, 
    15.85, 18.7875, 7.75, 31, 7.05, 21, 7.25, 13, 7.75, 113.275, 
    7.925, 27, 76.2917, 10.5, 8.05, 13, 8.05, 7.8958, 90, 9.35, 
    10.5, 7.25, 13, 25.4667, 83.475, 7.775, 13.5, 31.3875, 10.5, 
    7.55, 26, 26.25, 10.5, 12.275, 14.4542, 15.5, 10.5, 7.125, 
    7.225, 90, 7.775, 14.5, 52.5542, 26, 7.25, 10.4625, 26.55, 
    16.1, 20.2125, 15.2458, 79.2, 86.5, 512.3292, 26, 7.75, 31.3875, 
    79.65, 0, 7.75, 10.5, 39.6875, 7.775, 153.4625, 135.6333, 
    31, 0, 19.5, 29.7, 7.75, 77.9583, 7.75, 0, 29.125, 20.25, 
    7.75, 7.8542, 9.5, 8.05, 26, 8.6625, 9.5, 7.8958, 13, 7.75, 
    78.85, 91.0792, 12.875, 8.85, 7.8958, 27.7208, 7.2292, 151.55, 
    30.5, 247.5208, 7.75, 23.25, 0, 12.35, 8.05, 151.55, 110.8833, 
    108.9, 24, 56.9292, 83.1583, 262.375, 26, 7.8958, 26.25, 
    7.8542, 26, 14, 164.8667, 134.5, 7.25, 7.8958, 12.35, 29, 
    69.55, 135.6333, 6.2375, 13, 20.525, 57.9792, 23.25, 28.5, 
    153.4625, 18, 133.65, 7.8958, 66.6, 134.5, 8.05, 35.5, 26, 
    263, 13, 13, 13, 13, 13, 16.1, 15.9, 8.6625, 9.225, 35, 7.2292, 
    17.8, 7.225, 9.5, 55, 13, 7.8792, 7.8792, 27.9, 27.7208, 
    14.4542, 7.05, 15.5, 7.25, 75.25, 7.2292, 7.75, 69.3, 55.4417, 
    6.4958, 8.05, 135.6333, 21.075, 82.1708, 7.25, 211.5, 4.0125, 
    7.775, 227.525, 15.7417, 7.925, 52, 7.8958, 73.5, 46.9, 13, 
    7.7292, 12, 120, 7.7958, 7.925, 113.275, 16.7, 7.7958, 7.8542, 
    26, 10.5, 12.65, 7.925, 8.05, 9.825, 15.85, 8.6625, 21, 7.75, 
    18.75, 7.775, 25.4667, 7.8958, 6.8583, 90, 0, 7.925, 8.05, 
    32.5, 13, 13, 24.15, 7.8958, 7.7333, 7.875, 14.4, 20.2125, 
    7.25, 26, 26, 7.75, 8.05, 26.55, 16.1, 26, 7.125, 55.9, 120, 
    34.375, 18.75, 263, 10.5, 26.25, 9.5, 7.775, 13, 8.1125, 
    81.8583, 19.5, 26.55, 19.2583, 30.5, 27.75, 19.9667, 27.75, 
    89.1042, 8.05, 7.8958, 26.55, 51.8625, 10.5, 7.75, 26.55, 
    8.05, 38.5, 13, 8.05, 7.05, 0, 26.55, 7.725, 19.2583, 7.25, 
    8.6625, 27.75, 13.7917, 9.8375, 52, 21, 7.0458, 7.5208, 12.2875, 
    46.9, 0, 8.05, 9.5875, 91.0792, 25.4667, 90, 29.7, 8.05, 
    15.9, 19.9667, 7.25, 30.5, 49.5042, 8.05, 14.4583, 78.2667, 
    15.1, 151.55, 7.7958, 8.6625, 7.75, 7.6292, 9.5875, 86.5, 
    108.9, 26, 26.55, 22.525, 56.4958, 7.75, 8.05, 26.2875, 59.4, 
    7.4958, 34.0208, 10.5, 24.15, 26, 7.8958, 93.5, 7.8958, 7.225, 
    57.9792, 7.2292, 7.75, 10.5, 221.7792, 7.925, 11.5, 26, 7.2292, 
    7.2292, 22.3583, 8.6625, 26.25, 26.55, 106.425, 14.5, 49.5, 
    71, 31.275, 31.275, 26, 106.425, 26, 26, 13.8625, 20.525, 
    36.75, 110.8833, 26, 7.8292, 7.225, 7.775, 26.55, 39.6, 227.525, 
    79.65, 17.4, 7.75, 7.8958, 13.5, 8.05, 8.05, 24.15, 7.8958, 
    21.075, 7.2292, 7.8542, 10.5, 51.4792, 26.3875, 7.75, 8.05, 
    14.5, 13, 55.9, 14.4583, 7.925, 30, 110.8833, 26, 40.125, 
    8.7125, 79.65, 15, 79.2, 8.05, 8.05, 7.125, 78.2667, 7.25, 
    7.75, 26, 24.15, 33, 0, 7.225, 56.9292, 27, 7.8958, 42.4, 
    8.05, 26.55, 15.55, 7.8958, 30.5, 41.5792, 153.4625, 31.275, 
    7.05, 15.5, 7.75, 8.05, 65, 14.4, 16.1, 39, 10.5, 14.4542, 
    52.5542, 15.7417, 7.8542, 16.1, 32.3208, 12.35, 77.9583, 
    7.8958, 7.7333, 30, 7.0542, 30.5, 0, 27.9, 13, 7.925, 26.25, 
    39.6875, 16.1, 7.8542, 69.3, 27.9, 56.4958, 19.2583, 76.7292, 
    7.8958, 35.5, 7.55, 7.55, 7.8958, 23, 8.4333, 7.8292, 6.75, 
    73.5, 7.8958, 15.5, 13, 113.275, 133.65, 7.225, 25.5875, 
    7.4958, 7.925, 73.5, 13, 7.775, 8.05, 52, 39, 52, 10.5, 13, 
    0, 7.775, 8.05, 9.8417, 46.9, 512.3292, 8.1375, 76.7292, 
    9.225, 46.9, 39, 41.5792, 39.6875, 10.1708, 7.7958, 211.3375, 
    57, 13.4167, 56.4958, 7.225, 26.55, 13.5, 8.05, 7.7333, 110.8833, 
    7.65, 227.525, 26.2875, 14.4542, 7.7417, 7.8542, 26, 13.5, 
    26.2875, 151.55, 15.2458, 49.5042, 26.55, 52, 9.4833, 13, 
    7.65, 227.525, 10.5, 15.5, 7.775, 33, 7.0542, 13, 13, 53.1, 
    8.6625, 21, 7.7375, 26, 7.925, 211.3375, 18.7875, 0, 13, 
    13, 16.1, 34.375, 512.3292, 7.8958, 7.8958, 30, 78.85, 262.375, 
    16.1, 7.925, 71, 20.25, 13, 53.1, 7.75, 23, 12.475, 9.5, 
    7.8958, 65, 14.5, 7.7958, 11.5, 8.05, 86.5, 14.5, 7.125, 
    7.2292, 120, 7.775, 77.9583, 39.6, 7.75, 24.15, 8.3625, 9.5, 
    7.8542, 10.5, 7.225, 23, 7.75, 7.75, 12.475, 7.7375, 211.3375, 
    7.2292, 57, 30, 23.45, 7.05, 7.25, 7.4958, 29.125, 20.575, 
    79.2, 7.75, 26, 69.55, 30.6958, 7.8958, 13, 25.9292, 8.6833, 
    7.2292, 24.15, 13, 26.25, 120, 8.5167, 6.975, 7.775, 0, 7.775, 
    13, 53.1, 7.8875, 24.15, 10.5, 31.275, 8.05, 0, 7.925, 37.0042, 
    6.45, 27.9, 93.5, 8.6625, 0, 12.475, 39.6875, 6.95, 56.4958, 
    37.0042, 7.75, 80, 14.4542, 18.75, 7.2292, 7.8542, 8.3, 83.1583, 
    8.6625, 8.05, 56.4958, 29.7, 7.925, 10.5, 31, 6.4375, 8.6625, 
    7.55, 69.55, 7.8958, 33, 89.1042, 31.275, 7.775, 15.2458, 
    39.4, 26, 9.35, 164.8667, 26.55, 19.2583, 7.2292, 14.1083, 
    11.5, 25.9292, 69.55, 13, 13, 13.8583, 50.4958, 9.5, 11.1333, 
    7.8958, 52.5542, 5, 9, 24, 7.225, 9.8458, 7.8958, 7.8958, 
    83.1583, 26, 7.8958, 10.5167, 10.5, 7.05, 29.125, 13, 30, 
    23.45, 30, 7.75), Cabin = structure(c(1L, 83L, 1L, 57L, 1L, 
    1L, 131L, 1L, 1L, 1L, 147L, 51L, 1L, 1L, 1L, 1L, 1L, 1L, 
    1L, 1L, 1L, 113L, 1L, 15L, 1L, 1L, 1L, 65L, 1L, 1L, 1L, 43L, 
    1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 
    1L, 1L, 1L, 1L, 1L, 103L, 1L, 25L, 73L, 1L, 1L, 1L, 1L, 1L, 
    23L, 82L, 1L, 1L, 1L, 144L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 
    142L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 65L, 
    1L, 1L, 1L, 124L, 1L, 1L, 1L, 14L, 93L, 1L, 1L, 1L, 1L, 100L, 
    1L, 1L, 1L, 1L, 1L, 1L, 1L, 54L, 1L, 1L, 1L, 1L, 1L, 1L, 
    1L, 38L, 1L, 1L, 1L, 1L, 118L, 100L, 1L, 1L, 1L, 140L, 1L, 
    1L, 1L, 1L, 1L, 1L, 1L, 109L, 57L, 1L, 47L, 1L, 1L, 1L, 1L, 
    1L, 1L, 1L, 1L, 143L, 1L, 1L, 63L, 1L, 1L, 1L, 1L, 1L, 1L, 
    1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 125L, 1L, 1L, 1L, 20L, 1L, 
    1L, 1L, 16L, 1L, 1L, 71L, 1L, 1L, 1L, 1L, 1L, 146L, 1L, 11L, 
    1L, 1L, 1L, 1L, 1L, 1L, 1L, 143L, 30L, 45L, 1L, 1L, 1L, 1L, 
    1L, 1L, 1L, 1L, 1L, 147L, 1L, 1L, 1L, 10L, 1L, 1L, 1L, 1L, 
    1L, 105L, 1L, 1L, 95L, 1L, 1L, 1L, 1L, 1L, 89L, 1L, 1L, 1L, 
    1L, 1L, 82L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 
    1L, 1L, 1L, 80L, 1L, 1L, 104L, 1L, 1L, 147L, 85L, 1L, 1L, 
    1L, 1L, 42L, 1L, 1L, 1L, 1L, 136L, 48L, 1L, 1L, 1L, 1L, 59L, 
    91L, 1L, 1L, 1L, 56L, 1L, 115L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 
    1L, 5L, 1L, 1L, 1L, 1L, 1L, 1L, 33L, 92L, 1L, 1L, 1L, 1L, 
    64L, 53L, 38L, 1L, 1L, 1L, 118L, 1L, 64L, 1L, 76L, 1L, 127L, 
    74L, 37L, 1L, 1L, 1L, 1L, 1L, 1L, 78L, 126L, 1L, 1L, 1L, 
    1L, 1L, 67L, 1L, 92L, 1L, 19L, 1L, 58L, 87L, 1L, 1L, 1L, 
    63L, 129L, 1L, 148L, 143L, 65L, 1L, 1L, 1L, 144L, 1L, 1L, 
    1L, 1L, 1L, 61L, 1L, 1L, 1L, 1L, 125L, 1L, 1L, 1L, 1L, 1L, 
    1L, 1L, 1L, 1L, 106L, 1L, 1L, 26L, 133L, 1L, 1L, 1L, 1L, 
    1L, 1L, 81L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 
    1L, 49L, 1L, 1L, 105L, 147L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 
    1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 80L, 1L, 1L, 1L, 
    1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 117L, 
    73L, 1L, 1L, 1L, 130L, 49L, 1L, 1L, 65L, 1L, 1L, 1L, 1L, 
    1L, 1L, 12L, 1L, 1L, 1L, 52L, 1L, 1L, 55L, 88L, 1L, 1L, 128L, 
    99L, 1L, 1L, 119L, 1L, 135L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 
    1L, 1L, 1L, 92L, 1L, 3L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 
    33L, 1L, 89L, 27L, 1L, 1L, 1L, 1L, 66L, 1L, 1L, 1L, 98L, 
    1L, 64L, 1L, 1L, 1L, 1L, 1L, 44L, 76L, 1L, 1L, 1L, 1L, 1L, 
    1L, 123L, 1L, 1L, 108L, 144L, 1L, 1L, 1L, 41L, 1L, 1L, 19L, 
    1L, 1L, 1L, 90L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 28L, 1L, 
    1L, 29L, 22L, 1L, 1L, 1L, 84L, 1L, 1L, 1L, 1L, 1L, 79L, 1L, 
    1L, 1L, 1L, 1L, 4L, 1L, 136L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 
    1L, 1L, 1L, 1L, 1L, 50L, 123L, 1L, 1L, 1L, 1L, 130L, 1L, 
    1L, 1L, 77L, 1L, 2L, 1L, 137L, 1L, 31L, 1L, 1L, 1L, 98L, 
    1L, 1L, 1L, 1L, 1L, 1L, 1L, 6L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 
    1L, 1L, 59L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 146L, 1L, 1L, 
    97L, 1L, 1L, 1L, 112L, 1L, 116L, 1L, 1L, 7L, 1L, 35L, 1L, 
    1L, 1L, 1L, 1L, 1L, 1L, 1L, 26L, 1L, 1L, 1L, 103L, 1L, 9L, 
    1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 110L, 1L, 1L, 
    134L, 1L, 1L, 1L, 1L, 1L, 1L, 60L, 1L, 40L, 1L, 1L, 1L, 1L, 
    1L, 1L, 1L, 36L, 1L, 111L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 34L, 
    21L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 77L, 141L, 75L, 122L, 1L, 
    1L, 1L, 1L, 1L, 122L, 1L, 1L, 86L, 58L, 60L, 1L, 1L, 142L, 
    68L, 118L, 1L, 1L, 1L, 1L, 1L, 1L, 139L, 1L, 1L, 1L, 1L, 
    1L, 34L, 1L, 1L, 1L, 1L, 1L, 1L, 17L, 1L, 1L, 107L, 69L, 
    37L, 1L, 1L, 22L, 1L, 1L, 102L, 1L, 1L, 120L, 1L, 1L, 1L, 
    1L, 1L, 1L, 1L, 42L, 1L, 1L, 1L, 49L, 1L, 94L, 1L, 1L, 1L, 
    1L, 1L, 1L, 138L, 1L, 1L, 1L, 145L, 1L, 1L, 24L, 1L, 21L, 
    114L, 1L, 1L, 1L, 1L, 1L, 1L, 46L, 1L, 1L, 1L, 1L, 1L, 1L, 
    96L, 1L, 1L, 1L, 1L, 1L, 49L, 1L, 1L, 1L, 13L, 1L, 1L, 139L, 
    1L, 1L, 1L, 1L, 1L, 18L, 1L, 1L, 1L, 1L, 39L, 1L, 1L, 120L, 
    1L, 1L, 1L, 1L, 1L, 23L, 1L, 1L, 1L, 1L, 1L, 132L, 1L, 1L, 
    1L, 70L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 88L, 1L, 1L, 
    1L, 101L, 1L, 1L, 1L, 121L, 1L, 1L, 1L, 1L, 96L, 1L, 1L, 
    1L, 1L, 8L, 1L, 1L, 1L, 104L, 36L, 1L, 1L, 1L, 1L, 1L, 1L, 
    72L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 32L, 1L, 62L, 1L), .Label = c("", 
    "A10", "A14", "A16", "A19", "A20", "A23", "A24", "A26", "A31", 
    "A32", "A34", "A36", "A5", "A6", "A7", "B101", "B102", "B18", 
    "B19", "B20", "B22", "B28", "B3", "B30", "B35", "B37", "B38", 
    "B39", "B4", "B41", "B42", "B49", "B5", "B50", "B51 B53 B55", 
    "B57 B59 B63 B66", "B58 B60", "B69", "B71", "B73", "B77", 
    "B78", "B79", "B80", "B82 B84", "B86", "B94", "B96 B98", 
    "C101", "C103", "C104", "C106", "C110", "C111", "C118", "C123", 
    "C124", "C125", "C126", "C128", "C148", "C2", "C22 C26", 
    "C23 C25 C27", "C30", "C32", "C45", "C46", "C47", "C49", 
    "C50", "C52", "C54", "C62 C64", "C65", "C68", "C7", "C70", 
    "C78", "C82", "C83", "C85", "C86", "C87", "C90", "C91", "C92", 
    "C93", "C95", "C99", "D", "D10 D12", "D11", "D15", "D17", 
    "D19", "D20", "D21", "D26", "D28", "D30", "D33", "D35", "D36", 
    "D37", "D45", "D46", "D47", "D48", "D49", "D50", "D56", "D6", 
    "D7", "D9", "E10", "E101", "E12", "E121", "E17", "E24", "E25", 
    "E31", "E33", "E34", "E36", "E38", "E40", "E44", "E46", "E49", 
    "E50", "E58", "E63", "E67", "E68", "E77", "E8", "F E69", 
    "F G63", "F G73", "F2", "F33", "F38", "F4", "G6", "T"), class = "factor"), 
    Embarked = structure(c(4L, 2L, 4L, 4L, 4L, 3L, 4L, 4L, 4L, 
    2L, 4L, 4L, 4L, 4L, 4L, 4L, 3L, 4L, 4L, 2L, 4L, 4L, 3L, 4L, 
    4L, 4L, 2L, 4L, 3L, 4L, 2L, 2L, 3L, 4L, 2L, 4L, 2L, 4L, 4L, 
    2L, 4L, 4L, 2L, 2L, 3L, 4L, 3L, 3L, 2L, 4L, 4L, 4L, 2L, 4L, 
    2L, 4L, 4L, 2L, 4L, 4L, 2L, 1L, 4L, 4L, 2L, 2L, 4L, 4L, 4L, 
    4L, 4L, 4L, 4L, 2L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 3L, 4L, 
    4L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 2L, 2L, 4L, 
    4L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 3L, 4L, 2L, 4L, 4L, 
    2L, 4L, 3L, 4L, 2L, 4L, 4L, 4L, 2L, 4L, 4L, 2L, 3L, 4L, 2L, 
    4L, 2L, 4L, 4L, 4L, 4L, 2L, 4L, 4L, 4L, 2L, 2L, 4L, 4L, 3L, 
    4L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 2L, 3L, 4L, 4L, 
    4L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 3L, 4L, 4L, 
    2L, 4L, 4L, 2L, 4L, 4L, 4L, 2L, 4L, 4L, 4L, 4L, 3L, 4L, 3L, 
    4L, 4L, 4L, 4L, 4L, 2L, 2L, 3L, 4L, 3L, 4L, 4L, 4L, 4L, 2L, 
    4L, 4L, 4L, 2L, 3L, 2L, 4L, 4L, 4L, 4L, 3L, 2L, 4L, 4L, 2L, 
    4L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 
    4L, 4L, 4L, 4L, 4L, 4L, 2L, 3L, 4L, 4L, 2L, 3L, 4L, 4L, 4L, 
    4L, 4L, 4L, 4L, 4L, 4L, 2L, 2L, 4L, 2L, 4L, 3L, 4L, 4L, 4L, 
    3L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 2L, 3L, 4L, 4L, 4L, 3L, 
    4L, 3L, 4L, 4L, 4L, 4L, 2L, 4L, 4L, 4L, 3L, 4L, 2L, 2L, 4L, 
    4L, 2L, 2L, 4L, 4L, 2L, 3L, 3L, 4L, 3L, 4L, 4L, 2L, 2L, 2L, 
    2L, 2L, 2L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 2L, 4L, 4L, 3L, 4L, 
    4L, 2L, 4L, 4L, 4L, 2L, 3L, 4L, 4L, 4L, 4L, 4L, 4L, 2L, 4L, 
    4L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 2L, 4L, 
    2L, 4L, 4L, 4L, 3L, 3L, 4L, 2L, 2L, 4L, 3L, 4L, 2L, 2L, 3L, 
    2L, 2L, 4L, 4L, 2L, 4L, 2L, 4L, 2L, 2L, 4L, 2L, 2L, 4L, 4L, 
    4L, 4L, 4L, 4L, 3L, 2L, 4L, 4L, 4L, 2L, 4L, 4L, 4L, 4L, 4L, 
    4L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 3L, 3L, 4L, 
    4L, 4L, 4L, 4L, 4L, 4L, 2L, 3L, 4L, 4L, 4L, 4L, 4L, 4L, 3L, 
    4L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 
    4L, 4L, 4L, 4L, 2L, 4L, 4L, 4L, 2L, 2L, 4L, 2L, 4L, 4L, 4L, 
    3L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 3L, 2L, 4L, 4L, 4L, 2L, 
    4L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 2L, 4L, 4L, 2L, 4L, 
    4L, 4L, 4L, 4L, 2L, 4L, 2L, 2L, 4L, 4L, 4L, 4L, 3L, 3L, 4L, 
    4L, 2L, 4L, 4L, 4L, 4L, 3L, 4L, 4L, 2L, 4L, 4L, 4L, 3L, 4L, 
    4L, 4L, 4L, 2L, 2L, 2L, 3L, 4L, 4L, 4L, 4L, 4L, 2L, 2L, 2L, 
    4L, 4L, 4L, 2L, 4L, 2L, 4L, 4L, 4L, 4L, 2L, 4L, 4L, 2L, 4L, 
    4L, 2L, 4L, 3L, 2L, 4L, 4L, 2L, 2L, 4L, 4L, 3L, 4L, 4L, 4L, 
    4L, 4L, 4L, 4L, 2L, 4L, 4L, 4L, 4L, 3L, 4L, 4L, 4L, 4L, 2L, 
    4L, 4L, 2L, 4L, 2L, 2L, 4L, 4L, 2L, 4L, 4L, 4L, 2L, 4L, 3L, 
    4L, 4L, 4L, 4L, 2L, 2L, 4L, 4L, 4L, 4L, 2L, 4L, 4L, 4L, 2L, 
    4L, 4L, 4L, 3L, 3L, 4L, 4L, 4L, 4L, 4L, 4L, 2L, 4L, 2L, 4L, 
    4L, 4L, 3L, 4L, 4L, 3L, 4L, 4L, 2L, 4L, 4L, 4L, 4L, 4L, 4L, 
    4L, 4L, 2L, 4L, 4L, 2L, 2L, 4L, 2L, 4L, 4L, 4L, 4L, 4L, 3L, 
    3L, 4L, 4L, 3L, 4L, 2L, 4L, 2L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 
    4L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 2L, 3L, 2L, 4L, 4L, 
    4L, 2L, 4L, 4L, 4L, 4L, 4L, 2L, 4L, 2L, 4L, 4L, 4L, 3L, 2L, 
    4L, 2L, 4L, 2L, 3L, 4L, 4L, 4L, 4L, 4L, 2L, 2L, 4L, 4L, 4L, 
    4L, 4L, 2L, 4L, 3L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 3L, 4L, 
    4L, 4L, 2L, 4L, 4L, 4L, 4L, 4L, 2L, 4L, 4L, 4L, 4L, 2L, 4L, 
    4L, 4L, 4L, 4L, 4L, 3L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 
    4L, 4L, 4L, 2L, 4L, 4L, 4L, 2L, 3L, 3L, 4L, 4L, 4L, 4L, 2L, 
    4L, 4L, 3L, 4L, 3L, 4L, 2L, 4L, 4L, 4L, 4L, 4L, 4L, 3L, 4L, 
    2L, 3L, 4L, 4L, 2L, 4L, 4L, 4L, 4L, 2L, 4L, 4L, 4L, 4L, 2L, 
    4L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 2L, 4L, 
    4L, 4L, 4L, 4L, 4L, 4L, 3L, 4L, 2L, 3L, 1L, 2L, 4L, 2L, 4L, 
    4L, 2L, 4L, 4L, 4L, 2L, 4L, 4L, 2L, 2L, 4L, 4L, 4L, 2L, 4L, 
    2L, 4L, 4L, 2L, 4L, 4L, 4L, 4L, 4L, 2L, 2L, 4L, 4L, 4L, 4L, 
    4L, 4L, 2L, 4L, 4L, 4L, 4L, 4L, 4L, 4L, 2L, 2L, 4L, 4L, 4L, 
    2L, 4L, 4L, 4L, 4L, 4L, 3L, 4L, 4L, 4L, 2L, 3L), .Label = c("", 
    "C", "Q", "S"), class = "factor")), .Names = c("PassengerId", 
"Survived", "Pclass", "Name", "Sex", "Age", "SibSp", "Parch", 
"Ticket", "Fare", "Cabin", "Embarked"), class = "data.frame", row.names = c(NA, 
-891L))
```

## Titanic

```{r}

# Call dim on titanic
dim(titanic)

# Generate histogram of Age column
hist(titanic$Age)

```

## Exploratory queries

```{r}

# Print out total value of fares
str(titanic)
sum(titanic$Fare)

# Print out proportion of passengers that survived
mean(titanic$Survived)

```

## Infer gender from names(1)

```{r}

# Extract the name column from titanic
pass_names <- titanic$Name

# Create the logical vectror is_man
is_man <- grepl(", Mr\\.", pass_names)

# Count the number of men
sum(is_man)

# Count number of men based on gender
sum(titanic$Sex == "male")
```

## Infer gender from names(2)

```{r}

# Extract the name column from titanic
pass_names <- titanic$Name

# Create titles
titles <- gsub("^.*, (.*?)\\..*$", "\\1", pass_names)

# Call unique() on titles
unique(titles)

```

## Infer gender from names(3)

```{r}

pass_names <- titanic$Name
titles_male <- paste(",", c("Mr\\.", "Master", "Don", "Rev", "Dr\\.", "Major", "Sir", "Col", "Capt", "Jonkheer"))

# Finish the vapply() command
hits <- vapply(titles_male,
               FUN = grepl,
               FUN.VALUE = logical(length(pass_names)),
               pass_names)

# Calculate the sum() of hits
sum(hits)

# Count number of men based on gender
sum(titanic$Sex == "male")

```


### Why is it off by one?

  - They don't really address this in the class
  - But one function I use a lot is the `table` function
  - It easily shows whats off with the title assumptions and the gender
    - There is a female doctor on the boat but they are counting Dr as a male title
    - Definitely need to be careful with things like this. Other titles could go eaither way. 
  
```{r}
table(titles, titanic$Sex)
```

## Reformat passenger names

  - this is some great regex
  
```{r}

# pass_names is available in your workspace

convert_name <- function(name) {
  # women: take name from inside parentheses
  if (grepl("\\(.*?\\)", name)) {
    gsub("^.*?\\((.*?)\\)$", "\\1", name)
  # men: take name before comma and after title
  } else {
    # Finish the gsub() function
    gsub("^(.*?),\\s[a-zA-Z\\.]*?\\s(.*?)$", "\\2 \\1", name)
  }
}

# Call convert_name on name
clean_pass_names <- vapply(pass_names, FUN = convert_name,
                           FUN.VALUE = character(1), USE.NAMES = FALSE)

# Print out clean_pass_names
head(clean_pass_names, 20)

```

## Add birth dates

  - I'm adding the dob data manually
  
```{r, echo=F, results='hide'}


dob1 <- c("1890-01-03", "1874-02-28", "1886-01-17", "1877-03-04", "1876-09-28", 
NA, "1857-06-14", "1909-08-11", "1884-10-17", "1898-01-05", "1908-04-11", 
"1853-04-27", "1891-06-09", "1873-01-08", "1897-04-15", "1856-09-04", 
"1909-09-30", NA, "1880-09-03", NA, "1876-09-08", "1877-06-19", 
"1897-03-28", "1883-09-28", "1903-08-12", "1874-04-02", NA, "1892-12-29", 
NA, NA, "1871-12-06", NA, NA, "1845-09-04", "1883-12-26", "1869-10-30", 
NA, "1890-07-06", "1893-10-08", "1898-01-29", "1871-04-17", "1884-06-22", 
NA, "1908-12-19", "1892-08-28", NA, NA, NA, NA, "1893-08-24", 
"1905-02-26", "1891-03-11", "1862-11-26", "1882-06-02", "1847-01-25", 
NA, "1890-09-18", "1884-03-19", "1907-01-22", "1900-07-18", "1889-08-10", 
"1873-05-06", "1866-04-30", "1907-05-05", NA, NA, "1882-11-29", 
"1893-04-02", "1895-02-10", "1885-08-07", "1879-07-01", "1896-01-08", 
"1890-10-14", "1886-01-06", "1879-05-14", "1886-09-20", NA, NA, 
"1910-07-10", "1882-01-28", "1889-07-18", "1883-04-09", NA, "1884-01-17", 
"1895-04-07", "1878-07-03", "1896-03-27", NA, "1888-06-08", "1887-11-20", 
"1883-03-26", "1891-12-22", "1865-10-16", "1885-08-29", "1852-08-15", 
NA, "1840-05-02", "1888-11-26", "1878-03-31", "1877-12-10", "1883-10-18", 
NA, "1890-06-08", "1879-02-26", "1874-10-19", "1883-04-18", "1890-04-22", 
NA, "1873-06-06", NA, "1864-08-04", "1898-02-10", "1889-10-15", 
"1892-02-03", "1894-05-07", "1891-03-19", "1842-02-03", "1882-05-13", 
"1887-11-20", "1909-12-31", "1890-10-19", NA, "1880-01-03", "1880-01-14", 
"1857-05-19", "1899-09-07", NA, "1887-07-31", NA, "1866-11-26", 
"1879-04-05", "1891-11-27", "1864-10-22", "1883-03-10", "1886-12-03", 
"1889-02-24", "1892-07-15", "1874-07-02", "1895-11-24", "1887-09-02", 
NA, "1890-01-15", "1887-08-31", "1893-01-30", "1894-03-11", "1893-04-09", 
"1884-11-15", "1902-04-18", "1875-08-10", "1869-11-05", "1860-10-19", 
"1889-05-17", "1856-12-20", "1871-12-30", NA, "1861-02-01", "1896-03-25", 
"1881-06-22", NA, NA, "1867-07-25", "1872-01-23", "1885-04-27", 
"1894-12-16", "1911-02-05", "1902-12-13", NA, "1867-03-25", NA, 
"1884-02-29", "1850-12-25", "1907-06-21", "1910-05-20", "1891-02-03", 
"1855-10-22", "1893-12-24", NA, "1861-07-02", "1881-11-30", "1875-06-07", 
NA, NA, "1903-03-31", "1911-04-11", "1907-10-13", NA, NA, "1867-03-16", 
"1871-08-29", "1876-02-11", "1879-11-05", "1893-01-30", "1892-08-21", 
"1908-06-30", "1868-01-27", "1853-12-20", NA, "1869-09-19", NA, 
"1888-02-18", "1883-11-21", NA, "1878-01-10", "1866-11-18", "1894-02-19", 
"1910-02-06", "1879-09-01", "1885-10-11", "1896-04-12", "1871-06-17", 
"1888-02-14", "1876-09-06", "1890-04-03", "1881-06-22", NA, "1881-01-13", 
"1884-08-05", "1869-08-31", "1880-01-14", "1881-12-28", "1896-03-16", 
"1885-01-30", "1861-02-11", NA, "1873-06-15", "1889-09-11", "1892-06-25", 
"1891-07-18", "1894-04-04", NA, "1877-02-05", "1882-08-19", "1853-03-23", 
"1906-06-13", "1887-06-11", NA, "1868-01-15", "1904-03-04", "1892-08-04", 
"1878-08-06", NA, NA, "1883-04-09", "1889-04-17", "1881-07-24", 
"1868-02-01", "1886-07-12", "1887-09-29", "1874-11-15", "1858-03-06", 
NA, "1882-05-30", "1849-10-19", "1881-12-13", "1871-01-19", "1882-05-18", 
NA, "1881-05-09", "1876-11-03", "1861-12-08", NA, "1908-08-01", 
"1859-12-25", "1871-08-31", NA, "1875-07-19", "1896-03-24", "1886-09-19", 
"1853-08-12", "1876-10-27", NA, "1886-11-29", "1870-06-16", "1874-05-24", 
NA, "1848-05-24", "1866-05-18", NA, "1904-10-27", "1877-03-08", 
"1846-12-30", "1884-03-11", "1895-09-21", "1892-10-02", NA, "1878-11-19", 
"1881-08-04", "1889-11-09", "1869-10-06", "1889-12-25", "1886-01-08", 
"1892-10-20", "1876-04-09", "1888-03-15", "1887-10-04", NA, "1888-08-31", 
"1910-03-23", NA, "1862-03-25", NA, NA, "1893-03-11", NA, NA, 
"1911-01-26", NA, "1894-11-29", "1881-12-17", "1882-03-23", "1887-10-27", 
"1893-08-15", "1886-01-28", "1883-05-30", "1869-03-31", "1885-04-26", 
"1887-07-19", "1858-02-28", "1881-02-17", "1871-04-21", "1889-06-21", 
"1884-05-27", "1881-10-06", "1890-03-27", NA, "1876-02-11", "1851-02-05", 
"1876-03-15", "1880-12-26", "1895-05-15", NA, "1866-05-28", "1874-01-29", 
"1896-04-13", NA, NA, "1882-12-23", "1870-06-07", "1866-08-09", 
"1866-06-10", "1910-04-13", "1888-01-13", "1883-09-05", "1886-11-02", 
"1875-04-25", "1887-08-16", "1872-02-23", NA, "1908-05-31", "1870-01-28", 
"1888-05-20", NA, "1896-04-21", "1886-07-24", NA, "1883-09-12", 
"1889-05-09", "1874-01-30", NA, NA, "1871-10-17", "1882-10-21", 
"1866-12-28", "1877-02-12", NA, "1881-05-07", "1851-08-16", NA, 
NA, "1888-02-02", "1886-07-27", "1894-02-06", "1892-07-02", "1889-11-03", 
"1908-04-20", NA, "1889-05-04", "1884-11-19", "1891-12-30", "1892-12-09", 
"1869-11-11", "1910-12-18", "1879-06-30", "1876-07-10", NA, "1893-09-28", 
"1910-08-17", "1875-09-08", NA, "1894-09-19", "1876-03-29", "1890-07-05", 
"1883-08-26", "1889-03-03", "1887-10-24", "1889-11-14", "1880-07-18", 
"1866-02-26", "1888-08-07", "1883-12-17")

dob2 <- c("November 26, 1872", "September 09, 1885", "September 09, 1890", 
"April 01, 1884", "August 08, 1891", "April 20, 1877", "June 12, 1860", 
"July 07, 1908", "September 04, 1890", NA, NA, NA, "February 18, 1879", 
NA, "January 09, 1868", NA, "December 17, 1877", "May 24, 1893", 
"May 06, 1881", "March 31, 1902", NA, "January 06, 1891", "October 11, 1882", 
"March 02, 1884", "November 12, 1893", NA, "December 25, 1883", 
"March 13, 1893", NA, "February 15, 1880", "October 20, 1883", 
NA, "November 29, 1869", "January 30, 1895", "March 25, 1862", 
"May 01, 1897", "January 15, 1891", "December 21, 1887", "November 28, 1847", 
"April 12, 1881", "December 03, 1866", "December 11, 1891", "January 05, 1887", 
"November 15, 1883", NA, "November 11, 1907", "April 27, 1898", 
"January 27, 1878", "June 20, 1906", "June 15, 1859", "October 07, 1875", 
NA, "July 12, 1881", "August 18, 1862", NA, "January 01, 1883", 
"February 26, 1847", NA, "July 16, 1861", NA, "January 30, 1864", 
"November 09, 1877", "October 25, 1864", "January 27, 1864", 
NA, "March 19, 1874", NA, "November 28, 1855", NA, "December 06, 1910", 
NA, "September 20, 1873", "September 19, 1878", "April 05, 1889", 
"August 05, 1889", NA, "September 28, 1877", "December 28, 1882", 
"February 05, 1890", "August 04, 1909", "December 18, 1902", 
NA, "January 26, 1862", "February 22, 1849", "November 26, 1886", 
NA, "June 22, 1876", "March 11, 1854", "April 26, 1881", "January 27, 1903", 
NA, "February 20, 1891", "February 01, 1857", "June 15, 1840", 
"August 12, 1890", NA, "June 26, 1857", NA, "August 12, 1886", 
"September 01, 1887", "October 16, 1894", "May 30, 1890", NA, 
"May 12, 1874", "October 05, 1895", "July 20, 1893", "July 08, 1878", 
NA, "November 19, 1883", "June 12, 1885", "February 21, 1883", 
NA, "July 10, 1875", "January 30, 1858", "May 13, 1887", "October 29, 1864", 
"January 01, 1878", NA, "April 29, 1875", "April 24, 1879", "January 10, 1882", 
"January 21, 1890", NA, "October 29, 1867", NA, "September 04, 1871", 
"October 03, 1861", NA, "May 29, 1872", "February 15, 1889", 
"July 21, 1909", NA, "June 25, 1894", NA, "July 15, 1881", "September 04, 1904", 
"February 16, 1867", "April 07, 1882", NA, "August 04, 1889", 
"July 22, 1875", "November 28, 1902", "November 28, 1900", "May 05, 1879", 
"January 12, 1862", "August 06, 1847", "April 30, 1892", NA, 
"November 08, 1878", "April 21, 1903", "February 28, 1895", "October 24, 1884", 
NA, "May 24, 1889", "February 03, 1890", "February 04, 1850", 
"April 19, 1863", NA, "December 20, 1872", "September 23, 1875", 
NA, "March 10, 1872", "September 06, 1883", NA, NA, "April 16, 1887", 
"April 01, 1893", "April 02, 1883", NA, "January 14, 1880", "September 08, 1849", 
"September 14, 1858", "February 10, 1876", NA, "October 16, 1895", 
"December 25, 1892", "March 23, 1878", "July 17, 1872", NA, "October 26, 1879", 
"May 18, 1886", "October 25, 1872", "March 18, 1858", "July 03, 1875", 
NA, "February 23, 1894", "November 07, 1864", "September 07, 1851", 
"November 21, 1889", NA, "August 30, 1876", "February 06, 1860", 
"July 30, 1864", NA, "August 09, 1874", "September 14, 1875", 
NA, "January 23, 1863", NA, "December 04, 1862", "July 19, 1887", 
NA, NA, "August 07, 1867", "January 08, 1877", "January 23, 1876", 
"November 19, 1881", "May 27, 1884", "July 12, 1889", "July 20, 1871", 
"July 28, 1872", NA, NA, NA, "October 26, 1876", "September 07, 1887", 
"December 06, 1877", "June 06, 1885", "February 26, 1908", "September 29, 1885", 
"March 09, 1885", "January 24, 1870", "February 07, 1892", "April 13, 1891", 
"November 16, 1890", "October 02, 1850", "August 13, 1854", "December 28, 1890", 
"January 16, 1886", NA, "January 23, 1832", "February 27, 1861", 
"June 30, 1879", NA, "January 19, 1903", "February 17, 1884", 
"March 21, 1880", "December 17, 1880", "February 22, 1871", NA, 
"May 12, 1891", "December 03, 1887", "September 06, 1909", NA, 
"January 12, 1911", "January 10, 1864", "August 28, 1892", "August 17, 1855", 
NA, "December 16, 1888", NA, "August 10, 1893", "December 20, 1890", 
NA, "July 02, 1893", "July 14, 1887", NA, "May 30, 1879", "January 26, 1889", 
"January 17, 1854", "June 25, 1861", "March 08, 1872", "November 18, 1864", 
"September 24, 1875", "August 02, 1891", "August 03, 1879", "November 10, 1886", 
NA, "March 20, 1869", NA, "February 05, 1872", "September 20, 1880", 
"August 22, 1841", "October 20, 1880", NA, "December 09, 1893", 
"December 21, 1887", "February 09, 1894", "May 28, 1868", "March 15, 1876", 
NA, "July 16, 1884", "September 10, 1891", "July 19, 1897", "August 10, 1851", 
"July 12, 1886", "October 28, 1897", "January 02, 1893", "October 28, 1893", 
"April 21, 1896", "November 28, 1880", "November 21, 1907", NA, 
"January 18, 1887", "February 29, 1852", "September 24, 1859", 
"October 17, 1867", NA, "March 06, 1863", "August 30, 1869", 
"July 18, 1893", "June 15, 1876", "May 16, 1893", "December 19, 1886", 
"August 23, 1885", "May 08, 1872", "May 09, 1866", "November 07, 1869", 
"December 19, 1889", NA, "June 24, 1887", NA, "July 21, 1863", 
"March 25, 1883", "October 22, 1859", "January 10, 1893", "July 04, 1873", 
"December 04, 1884", NA, "November 29, 1878", "October 29, 1905", 
"July 11, 1894", "July 15, 1877", "October 02, 1861", "February 13, 1885", 
"July 09, 1891", "January 22, 1882", NA, "October 22, 1886", 
"August 14, 1886", "January 24, 1883", "July 10, 1900", NA, "August 15, 1888", 
"September 16, 1888", "December 26, 1883", "December 10, 1863", 
"January 03, 1877", NA, NA, NA, "June 14, 1875", "May 03, 1890", 
"November 10, 1887", "May 26, 1880", "April 20, 1841", "January 24, 1896", 
"July 21, 1881", "December 19, 1892", "July 05, 1880", "April 05, 1908", 
"December 18, 1905", "January 25, 1879", "December 26, 1888", 
"April 18, 1863", "October 15, 1910", "January 07, 1884", "August 13, 1893", 
"February 08, 1878", "March 03, 1879", NA, "February 21, 1871", 
"September 27, 1891", "October 04, 1875", "October 20, 1895", 
"June 13, 1860", NA, "March 03, 1882", NA, "June 27, 1879", "December 03, 1887", 
"May 01, 1863", "August 28, 1854", NA, "November 21, 1857", "March 28, 1894", 
NA, "July 02, 1906", NA, "June 19, 1868", "February 25, 1899", 
"October 01, 1894", "July 06, 1882", NA, "March 11, 1887", "April 04, 1887", 
"March 30, 1894", "November 01, 1903", "April 09, 1911", "August 20, 1865", 
NA, "August 30, 1895", NA, NA, "September 15, 1886", "July 27, 1872", 
"May 09, 1862", "May 14, 1880", "September 19, 1881", "April 10, 1882", 
"May 26, 1877", "July 04, 1880", "July 04, 1900", "December 07, 1910", 
"October 17, 1884", "January 05, 1881", "August 08, 1872", "February 17, 1894", 
"October 16, 1872", "December 20, 1878", "March 25, 1886", "April 20, 1872", 
"October 22, 1876", "February 09, 1906", "August 12, 1881", NA, 
"December 20, 1888", "November 26, 1880", "December 08, 1868", 
"March 30, 1902", "July 09, 1859", "August 13, 1884", "February 14, 1874", 
"June 06, 1884", "March 17, 1910", NA, NA, "May 16, 1910", NA, 
"December 21, 1849", "April 03, 1897", "December 03, 1910", NA, 
"September 15, 1888", "February 02, 1894", "July 19, 1872", "December 10, 1890", 
NA, "November 14, 1879", NA, "December 06, 1891", "April 24, 1895", 
"May 08, 1881", "April 08, 1878", "March 23, 1895", "June 25, 1869", 
NA, "December 23, 1876", "August 13, 1883", NA, "December 18, 1907", 
"May 11, 1837", "April 08, 1903", "May 31, 1895", "August 31, 1867", 
"February 20, 1894", "February 11, 1867", "December 15, 1860", 
"December 28, 1887", NA, "March 03, 1871", "September 30, 1890", 
"December 27, 1863", NA, "September 24, 1887", "March 13, 1870", 
"May 28, 1884", "May 05, 1880", NA, "December 29, 1907", "February 26, 1886", 
"April 16, 1864", "January 16, 1879", "January 25, 1865", "October 12, 1883", 
"November 25, 1896", "August 23, 1891", "October 01, 1892", NA, 
"August 02, 1855", "May 11, 1886", "January 03, 1879", "June 04, 1889", 
"January 25, 1884", "February 21, 1887", "August 18, 1872", "August 24, 1884", 
"March 23, 1893", NA, "December 03, 1885", "August 16, 1879")
```

  - the `strptime` function documentation has all the codes for matching dates
  
```{r}

# Have a look at head() of dob1 and dob2
head(dob1)
head(dob2)

# Convert dob1 to dob1d, convert dob2 to dob2d
dob1d <- as.Date(dob1, '%Y-%m-%d')

dob2d <- as.Date(dob2, '%b %d, %Y')

# Combine dob1d and dob2d into single vector: birth_dates
birth_dates <- c(dob1d,dob2d)

```

## Average age

  - I have not used subset in a while. Just use `dplyr::filter` now
```{r}

# titanic, dob1 and dob2 are preloaded
dob1d <- as.Date(dob1)
dob2d <- as.Date(dob2, format = "%B %d, %Y")
birth_dates <- c(dob1d, dob2d)
disaster_date <- as.Date("1912-04-15")

# Add birth_dates to titanic (column Birth)
titanic$Birth <- birth_dates
str(titanic)

# Create subset: survivors
survivors <- subset(titanic, Survived == TRUE)
str(survivors)

# Calculate average age of survivors
mean(disaster_date - survivors$Birth, na.rm=T)

```

## Age in perspective

```{r}

## How old was the youngest survivor when Apollo 11 landed on the moon on July 20, 1969?

min(as.Date('1969-07-20') - survivors$Birth, na.rm=T)

```

# Conclusion

  - This is great practice for the skillsets
  - I do think a lot of data analsyis is made simpler with the use of dataframes
  - The dplyr, tidyr, and other libraries replace a lot of the disparate R functions around data crunching into a clean and consistent grammar
    - There are courses on these topics later in the Data Science Track
  - A lot of this is core data crunching skills or even just programming skills that are needed in any language really so this is a good class for anyone getting into programing
    - But I would caution that there are much smoother ways to work with data in R. Its really what makes R so great. It has the best libraries I think. 
    - I'm looking forward to covering more of the library functions in docs like this soon. 
    