---
title: "Correlation and Regression"
author: "William Surles"
date: "2017-09-04"
output: 
 html_document:
  self_contained: yes
  theme: flatly
  highlight: tango
  toc: true
  toc_float: true
  toc_depth: 3
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(eval=T, echo=T, cache=T, message=F, warning=F)
```

# Introduction

  - Course notes from the [Correlation and Regression](https://www.datacamp.com/courses/correlation-and-regression) course on DataCamp
  
## Whats Covered

  - Visualizing two variables
  - Correlation
  - Simple linear regression
  - Interpretting regression models
  - Model Fit
  
## Additional Resources
  
  - [Equation samples in Rmarkdown](http://www.statpower.net/Content/310/R%20Stuff/SampleMarkdown.html)
  

## Libraries and Data

```{r, cache=F} 

library(dplyr)
library(tidyr)
library(ggplot2)
library(openintro)
library(broom)

source('create_datasets.R')

```


&nbsp; &nbsp;

***  
# Visualizing two variables
***  


## Visualizing bivariate relationships

  - Both variables are numerical
  - Response variable
    - y, dependent
  - Explanatory variable
    - x, independent, predictor
    - Something you think is related to the response
  - Plot with a scatter plot
    - Or discretize the explanotory variable with `cut()` function and use a box plot
  
    
### -- Scatterplots

```{r}

# Scatterplot of weight vs. weeks
glimpse(ncbirths)

ggplot(ncbirths, aes(x = weeks, y = weight)) + 
  geom_point()

```

### -- Boxplots as discretized/conditioned scatterplots

```{r}

# Boxplot of weight vs. weeks
ggplot(data = ncbirths, 
       aes(x = cut(weeks, breaks = 5), y = weight)) + 
  geom_boxplot()

```

## Characterizing bivariate relationships

  - Form (linear, quadratic, non-linear, etc)
  - Direction (positive or negative)
  - Strength (how much scatter/noise)
  - Outliers
  
### -- Creating scatterplots

  - The `mammals` dataset contains information about 39 different species of mammals, including their body weight, brain weight, gestation time, and a few other variables.
  - The `mlbBat10` dataset contains batting statistics for 1,199 Major League Baseball players during the 2010 season.
  - The `bdims` dataset contains body girth and skeletal diameter measurements for 507 physically active individuals.
  - The `smoking` dataset contains information on the smoking habits of 1,691 citizens of the United Kingdom.

```{r}

# Mammals scatterplot
glimpse(mammals)
ggplot(mammals, aes(x = BodyWt, y = BrainWt )) + 
  geom_point()

# Baseball player scatterplot
glimpse(mlbBat10)
ggplot(mlbBat10, aes(x = OBP, y = SLG)) + 
  geom_point()

# Body dimensions scatterplot
glimpse(bdims)
ggplot(bdims, aes(x = hgt, y = wgt, color = factor(sex))) + 
  geom_point()

# Smoking scatterplot
glimpse(smoking)
ggplot(smoking, aes(x = age, y = amtWeekdays)) + 
  geom_point()

```

### -- Transformations

```{r}

# Scatterplot with coord_trans()
str(mammals)
ggplot(data = mammals, aes(x = BodyWt, y = BrainWt)) +
  geom_point() + 
  coord_trans(x = "log10", y = "log10")

# Scatterplot with scale_x_log10() and scale_y_log10()
ggplot(data = mammals, aes(x = BodyWt, y = BrainWt)) +
  geom_point() +
  scale_x_log10() + 
  scale_y_log10()


```

## Outliers

### -- Identifying outliers

  - Be mindul of the count of datapoints when working with rate statistics
    - Its common to filter on count so that the rates have the chance to approach their long-run frequencies.
```{r}

# Scatterplot of SLG vs. OBP
str(mlbBat10)
mlbBat10 %>%
  filter(AB >= 200) %>%
  ggplot(aes(x = OBP, y = SLG)) +
  geom_point()

# Identify the outlying player
mlbBat10 %>%
  filter(
    AB >= 200, 
    OBP < 0.2
    )

```


&nbsp; &nbsp;

***  
# Correlation
***

## Quantifying the strength of bivariate relationships

  - Correlation coefficient between -1 and 1
    - Sign -> direction
    - Magnitude -> strength
  - Becareful. If the variables are related in a non-linear fashion they may have week correlation but there can be a great model fit. e.g. quadratic
  - The Pearson product-momemt correlation is the most common measure of correlation
    - This is usually what is meant when someone says 'correlation'

$$ r(x,y) = \frac{Cov(x,y)}{\sqrt{SXX \dot{} SYY}} $$

  - Another definition with just x and y and their means

$$ r(x,y) = \frac{\sum_{i=1}^n(x_i - \bar{x})(y_i - \bar{y})}{\sum_{i=1}^n(x_i - \bar{x})^2 \dot{} \sum_{i=1}^n(y_i - \bar{y})^2 } $$

  - denominator is the sums of the squared deviations
  - numerator is the sum of terms involving both x and y
    - each term in the sum is the area of the rectangle with corners at each datapoint $x_i$ and $y_i$ and the means $\bar{x}$ and $\bar{y}$
    

### -- Understanding correlation scale

  - Any correlation above 1 or below -1 is invalid
  
### -- Understanding correlation sign

  - A correlation of -.7 is stronger than a correlation of .6. It is just in the negative direction. 

### -- Computing correlation

```{r}

# Compute correlation
ncbirths %>%
  summarize(N = n(), r = cor(mage, weight))

# Compute correlation for all non-missing pairs
ncbirths %>%
  summarize(N = n(), r = cor(weeks, weight, use = "pairwise.complete.obs"))

```

## The Anscombe dataset

  - 4 data sets with similar statistics but very different patterns
  - This is an example of why we need to be careful when just looking at summary statistics
  
```{r}

ggplot(Anscombe, aes(x,y)) + 
  geom_point() + 
  facet_wrap(~ set)

```

### -- Exploring Anscombe

```{r}

## Anscombe dataset is loaded

# Compute properties of Anscombe
Anscombe %>%
  group_by(set) %>%
  summarize(
    N = n(), 
    mean(x), 
    sd(x), 
    mean(y), 
    sd(y), 
    cor(x, y)
    )

```

### -- Perception of correlation (2)

My guesses:
  
  - All baseball players
    - .6
  - All players with atleast 200 AB
    - .75
  - Body dimensions
    - male - .64
    - female - .72
  - mammals with and without log
    - .9

Actual:

```{r}

str(mlbBat10)
# Correlation for all baseball players
mlbBat10 %>%
  summarize(N = n(), r = cor(OBP, SLG))

# Correlation for all players with at least 200 ABs
mlbBat10 %>%
  filter(AB >= 200) %>%
  summarize(N = n(), r = cor(OBP, SLG))

# Correlation of body dimensions
str(bdims)
bdims %>%
  group_by(sex) %>%
  summarize(N = n(), r = cor(hgt, wgt))

# Correlation among mammals, with and without log
str(mammals)
mammals %>%
  summarize(N = n(), 
            r = cor(BodyWt, BrainWt), 
            r_log = cor(log(BodyWt), log(BrainWt)))

```

## Interpretation of correlation

  - Be careful not to inflate a correlation with a causation. 
  - Use careful language to suggest there is an association, but not to say one thing causes another if you don't know for sure that it does. 

### -- Correlation and causation

In the San Francisco Bay Area from 1960-1967, the correlation between the birthweight of 1,236 babies and the length of their gestational period was 0.408. Which of the following conclusions *is not* a valid statistical interpretation of these results.

  - We observed that babies with longer gestational periods tended to be heavier at birth.
  - It may be that a longer gestational period contributes to a heavier birthweight among babies, but a randomized, controlled experiment is needed to confirm this observation.
  - *Staying in the womb longer causes babies to be heavier when they are born.*
  - These data suggest that babies with longer gestational periods tend to be heavier at birth, but there are many potential confounding factors that were not taken into account.
  
## Spurious correlation

  - There can be correlations that fit but are just linked by time. 
  - Also maps can create spurious correltations where the confounding factor is just population

### -- Spurious correlation in random data

```{r}

# Create faceted scatterplot
str(noise)
ggplot(noise, aes(x, y)) + 
  geom_point() + 
  facet_wrap(~ z)

# Compute correlations for each dataset
noise_summary <- noise %>%
  group_by(z) %>%
  summarize(N = n(), spurious_cor = cor(x, y))

# Isolate sets with correlations above 0.2 in absolute strength
noise_summary %>%
  filter(abs(spurious_cor) > 0.2)

```


&nbsp; &nbsp;

***
# Simple linear regression
***

## Visualization of linear models

### -- The "best fit" line

```{r}

# Scatterplot with regression line
str(bdims)
ggplot(data = bdims, aes(x = hgt, y = wgt)) + 
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE)

```

### -- Uniqueness of least squares regression line

```{r, echo=F}

add_line <- function (my_slope) 
{
    bdims_summary <- bdims %>% summarize(N = n(), r = cor(hgt, 
        wgt), mean_hgt = mean(hgt), mean_wgt = mean(wgt), sd_hgt = sd(hgt), 
        sd_wgt = sd(wgt)) %>% mutate(true_slope = r * sd_wgt/sd_hgt, 
        true_intercept = mean_wgt - true_slope * mean_hgt)
    p <- ggplot(data = bdims, aes(x = hgt, y = wgt)) + geom_point() + 
        geom_point(data = bdims_summary, aes(x = mean_hgt, y = mean_wgt), 
            color = "red", size = 3)
    my_data <- bdims_summary %>% mutate(my_slope = my_slope, 
        my_intercept = mean_wgt - my_slope * mean_hgt)
    p + geom_abline(data = my_data, aes(intercept = my_intercept, 
        slope = my_slope), color = "dodgerblue")
}
```

```{r}

# Estimate optimal value of my_slope
ggplot(data = bdims, aes(x = hgt, y = wgt)) + 
  geom_point()

add_line(my_slope = 1) 
add_line(my_slope = 2)
add_line(my_slope = .5)


```

## Understanding the linear model

General statistic model

$$ response = f(explanatory) + noise $$
General linear model

$$ response = intercept + (slope * explanatory) + noise $$

Regression Model

$$ Y = \beta_0 +  \beta_1 \dot{} X + \epsilon $$
$$ \epsilon \sim N(0,\sigma_\epsilon) $$
  
  - $\beta_0$ is the Y-intercept 
  - $\beta_1$ is the slope
  - $\epsilon$ is the noise term
    - The distribution of the noise is normal with mean 0 and a fixed standard deviation
    
Fitted Values
$$ \hat{Y} = \hat{\beta_0} + \hat{\beta_1} \dot{} X$$
  
  - The difference between $\hat{Y}$ and $Y$ is that $Y$ is the actual observed values of the respose, while $\hat{Y}$ is the expected value of the response based on the model 
  
Residuals
$$ e = Y - \hat{Y} $$
 
  - the residuals are the realization of the noise term
  - $\epsilon$ is an unknown true quantity while $e$ is a known estimate of the quantity
    - $e$ is based on our model which is the estimate. We don't know the true fit. 

Fitting procedure

  - The best fit model is the one that minimiazes the sum of the squarred residuals. 
  - Find $\hat{\beta_0}$,$\hat{\beta_1}$ that minimize $\sum{i=1}^n e_i^2$
    
Key Concepts

  - Y-hat is expected value given corresponding X
  - Beta-hats are estimates of true, unknown betas
  - Residuals (e's) are estimates of true, unknown epsilons
  - "Error" may be misleading term - beter to use "noise"

### -- Fitting a linear model "by hand"

```{r}

# Print bdims_summary
bdims_summary <- bdims %>% 
  summarize(
    N = n(), 
    r = cor(hgt, wgt), 
    mean_hgt = mean(hgt), 
    mean_wgt = mean(wgt), 
    sd_hgt = sd(hgt), 
    sd_wgt = sd(wgt)
    )

# Add slope and intercept
bdims_summary %>%
  mutate(
    slope = r * sd_wgt/sd_hgt, 
    intercept = mean_wgt - (slope * mean_hgt)
    )

```

## Regression vs. regression to the mean

  - E.G sons are not likely to be as tall as their fathers or as short as their fathers
    - The mean height of all sons is the same as the mean height as all fathers
    - They are just not as tall or as short. They have regressed toward the mean

### -- Regression to the mean

```{r}

# Height of children vs. height of father
glimpse(Galton_men)
ggplot(data = Galton_men, aes(x = father, y = height)) +
  geom_point() + 
  geom_abline(slope = 1, intercept = 0) + 
  geom_smooth(method = "lm", se = FALSE)

# Height of children vs. height of mother
glimpse(Galton_women)
ggplot(data = Galton_women, aes(x = mother, y = height)) +
  geom_point() + 
  geom_abline(slope = 1, intercept = 0) + 
  geom_smooth(method = "lm", se = FALSE)

```

### -- "Regression" in the parlance of our time

In an opinion piece about nepotism published in The New York Times in 2015, economist Seth Stephens-Davidowitz wrote that:

> "Regression to the mean is so powerful that once-in-a-generation talent basically never sires once-in-a-generation talent. It explains why Michael Jordan’s sons were middling college basketball players and Jakob Dylan wrote two good songs. It is why there are no American parent-child pairs among Hall of Fame players in any major professional sports league."

The author is arguing that ...
  
  - Because of regression to the mean, an outstanding basketball player is likely to have sons that are good at basketball, but not as good as him.

&nbsp; &nbsp;

***
# Interpretting regression models
***

## Interpretation of regression coefficients

### -- Fitting simple linear models

```{r}

# Linear model for weight as a function of height
lm(wgt ~ hgt, data = bdims)

# Linear model for SLG as a function of OBP
lm(SLG ~ OBP, data = mlbBat10)

# Log-linear model for body weight as a function of brain weight
lm(log(BodyWt) ~ log(BrainWt), data = mammals)

```

## Your linear model object

### -- The lm summary output

```{r}

mod <- lm(wgt ~ hgt, data = bdims)

# Show the coefficients
coef(mod)

# Show the full output
summary(mod)

```

### -- Fitted values and residuals

```{r}

mod
# Mean of weights equal to mean of fitted values?
mean(fitted.values(mod)) == mean(bdims$wgt)

# Mean of the residuals
mean(residuals(mod))

```

### -- Tidying your linear model

```{r}

# Load broom

# Create bdims_tidy
bdims_tidy <- augment(mod)

# Glimpse the resulting data frame
glimpse(bdims_tidy)
head(bdims_tidy)
```

## Using your linear model

Looking at what books are most overpriced based on the model. 
  
```{r}

mod <- lm(uclaNew ~ amazNew, data = textbooks)

augment(mod) %>%
  arrange(desc(.resid)) %>%
  head()

textbooks %>%
  filter(uclaNew == 197)

```

Make predictions
  
```{r}

new_data <- data.frame(amazNew = 8.49)
predict(mod, newdata = new_data)

```

Visualize new observations
  
```{r, eval=F}

isrs <- broom::augment(mod, newdata = new_data)

ggplot(textbooks, aes(amazNew, uclaNew)) +
  geom_point() + 
  geom_smooth(method = "lm") + 
  geom_point(data = isrs, aes(y = .fitted), size = 3, color = "red")
  
```


### -- Making predictions

```{r}

# Print ben
ben <- data.frame(wgt = 74.8, hgt = 182.8)
ben

mod <- lm(wgt ~ hgt, data = bdims)
 
# Predict the weight of ben
predict(mod, newdata = ben)

```

### -- Adding a regression line to a plot manually

```{r}

coefs <- coef(mod)

# Add the line to the scatterplot
ggplot(data = bdims, aes(x = hgt, y = wgt)) + 
  geom_point() + 
  geom_abline(data = as.data.frame(as.list(coefs)), 
              aes(intercept = X.Intercept., slope = hgt),  
              color = "dodgerblue")

```


&nbsp; &nbsp;

***
# Model Fit
***

## Assessing model fit

SSE

  - penalizes large residuals dispoportionally
  - it can be calcualted easily after using broom to tidy the model
  - its a little hard to interpret because the values have been squarred

```{r, eval=F}

mod_possum <- lm(totalL ~ tailL, data = possum)
mod_possum %>%
  augment() %>%
  summarize(
    SSE = sum(.resid^2),
    SSE_also = (n() -1) * var(.resid)
  )

```

RMSE
  
  - We use degrees of freedom not the number of values
    - Degrees of freedom is not discussed here but its an important concept to learn about. 
  - RMSE is in the same units ad the response. Thats very useful and in intuative

$$RMSE = \sqrt{\frac{\sum_ie_i^2}{d.f}} = \sqrt{\frac{SSE}{n-2}}$$ 

### -- Standard error of residuals

```{r}

# View summary of model
summary(mod)

# Compute the mean of the residuals
mean(residuals(mod))

# Compute RMSE
sqrt(sum(residuals(mod)^2) / df.residual(mod))

```

## Comparing model fits

NULL(average) model
  
  - Where the average value of all observations is used as the predictor value
  
$$ \hat{y} = \bar{y} $$

```{r}

## the null model
lm(totalL ~ 1, data = possum) %>%
  augment(possum) %>%
  summarize(SST = sum(.resid^2))

## Using tail as an explanatory variable
lm(totalL ~ tailL, data = possum) %>%
  augment(possum) %>%
  summarize(SST = sum(.resid^2))
  

```
  
Coeeficient of Determination

$$ R^2 = 1 - \frac{SSE}{SST} = 1 - \frac{Var(e)}{Var(y)} $$
  
  - `e` is the vector of residuals and `y` is the response variable.
  - SSE for the null model is called the SST (Total sum of squares)
  - By building a regression model we hope to explain some of the variability in the response variable.
  - The portion of the variability that is not explained by our model is the SSE
  - The coefficient of determination (R^2) is the porportion of the variability in the response variable that is explained by our model.
    - This is the most commonly cited measure of the quality of the fit of a regression model

Connection to correlation

For simple linear regresssion...

$$ r_{x,y}^2 = R^2 $$
  
  - The value of R^2 is just the square of the correlation coefficient
  - use the `summary()` function to see the R^2 value
  - Again, this shows what percent of the variablity in the response variable is explained by our model
  -  A high R^2 value does not mean you have a good model. It could be overfit, or it could violate the conditions for inference (which will be discussed later)
  
### -- Assessing simple linear model fit

```{r}

# View model summary
str(bdims_tidy)
summary(mod)

var(bdims_tidy$.resid)
var(bdims_tidy$wgt)

# Compute R-squared
bdims_tidy %>%
  summarize(
    var_y = var(wgt), 
    var_e = var(.resid)
    ) %>%
  mutate(R_squared = 1 - (var_e/var_y))

```

### -- Linear vs. average

```{r}

# Compute SSE for null model
mod_null <- lm(wgt ~ 1, data = bdims) %>%
  augment()
head(mod_null)

mod_null %>%
  summarize(SSE = var(.resid))

# Compute SSE for regression model
mod_hgt <- mod %>%
  augment()

head(mod_hgt)

mod_hgt %>%
  summarize(SSE = var(.resid))

```

## Unusual points

  - **Leverage** is a function of the distance between the value of the explanatory variable and the mean of the explanatory variable
    - in other words points that are close to the horizontal center of the scatterplot have low leverage and points far from the horizontal center have high leverage. 
    - the y coordinate does not matter at all. 
    - look at `.hat` variable from `augment()` function

$$  h_i = \frac{1}{n} + \frac{(x_i - \bar{x})^2}{\sum_{i=1}^n(x_i - \bar{x})^2}$$

  - these points may or may not have a high **influence** on the regression line. If the points are already close to the regression line then they will not have much influence. 
    - the combination of leverage and the residual determine the influence
    - the cooks distance shows the influence. 
    - `.cooksd` in augment
    
### -- Leverage

```{r}

# Rank points of high leverage
mod <- lm(formula = SLG ~ OBP, data = filter(mlbBat10, AB >= 10))

mod %>%
  augment %>%
  arrange(desc(.hat)) %>%
  head()

```

### -- Influence

```{r}

# Rank influential points
mod %>%
  augment() %>%
  arrange(desc(.cooksd)) %>%
  head()

```

## Dealing with unusual points

  - All you can do is remove the outlier
    - This is a decision you can make but its important to do this cautiously
    - There needs to be really good reason. Not just that it messes up your ideal model
  - There are actually more sophisticated ways to determine if a point can be removed or not, but its not covered in the class. 

### -- Removing outliers

```{r}

# Create nontrivial_players
nontrivial_players <- mlbBat10 %>%
  filter(AB >= 10, OBP < 0.5)

# Fit model to new data
mod_cleaner <- lm(SLG ~ OBP, data = nontrivial_players)

# View model summary
summary(mod_cleaner)

# Visualize new model
ggplot(data = nontrivial_players, aes(x = OBP, y = SLG)) +
  geom_point() + 
  geom_smooth(method = "lm")

```

### -- High leverage points

```{r}

# Rank high leverage points
mod %>%
  augment() %>%
  arrange(desc(.hat), .cooksd) %>%
  head()

```

## Conclusion

  - Pretty good class. Definitely hits the basics of modeling (linear models) in R. 
  - There is a lot more that needs to be understood to really know I am taking the right steps to fit a correct model to a real world dataset. 