---
title: "Introduction to R"
author: "William Surles"
date: "2017-03-28"
output: 
 html_document:
  self_contained: yes
  theme: flatly
  highlight: tango
---

***
> This document was updated on **`r Sys.Date()`**.

***

```{r setup, include=FALSE}
library(shiny)
library(codetools)
knitr::opts_chunk$set(eval=T, echo=T, cache=T, message=F, warning=F)
```

# Introduction

  - This all comes from [Introduction to R](https://campus.datacamp.com/courses/free-introduction-to-r) course on DataCamp
    - I highly recommend the [DataCamp](https://www.datacamp.com/home) courses
    - They step you through the work and include videos and much more description than I include here
    - This doc is no replacement for working through the classes on your own!
  - I am copying my work from the class to this document
  - This will make it much easier for me to find and refernce what I have done/learned  
  - And I may insert a few of my own thoughts and comments about things along the way

## Whats Covered

  - Intro
    - Arithmetic
    - Variable assignment
    - Basic data types
  - Vectors
    - Creating, naming, aritmetic, selection
  - Matrices
    - Creating, naming, aritmetic, selection
  - Factors
    - Creating, levles, aritmetic, ordering
  - Data frames
    - Creating, looking, selection, sorting
  - Lists
    - Creating, naming, selection
    
  
&nbsp; &nbsp;

***  

# Intro to Basics

### How it Works

  - This section has some pretty juicy R charts here.   
  - This was hott stuff back in 1996 
    - Many of these styles of charts are still usefull now for data exploration
    - Thought I would not use base R graphics to make them
    - I'd start with ggplot2 or ggvis
    - Still its fun to see these charts. Don't wory about the code to make them for now. 

```{r results='hide'}

# The hashtag is used to add comments

# Show some demo graphs generated with R
demo("graphics")

```

### Arthmetic with R

Yeah, it does all the usual stuff.

```{r}

# An addition
5 + 5

# A substraction
5 - 5

# A multiplication
3 * 5

# A division
(5 + 5)/2

# Exponentiation
2^5

# Modulo
17%%4

```

### Variable assignment

```{r}

# Assign the value 42 to x
x <- 42

# Print out the value of the variable x
x

```

### Variable assignment 2

```{r}

# Assign the value 5 to the variable called my_apples
my_apples <- 5
my_apples

```

### Variable assignment 3

```{r}
# Assign a value to the variables called my_apples and my_oranges
my_apples <- 5
my_oranges <- 6

# Add these two variables together and print the result
my_fruit <- my_apples + my_oranges
my_fruit

```

### Apples and oranges

  - This would fail because you can't add character vectors.
  - Rstudio won't even knit the document. 
  - It will throw an error in the code.
  - I have `eval=F` so this will run and just paste the error below

```{r, eval=F}

# Assign a value to the variable called my_apples
my_apples <- 'apples'
my_oranges <- 'oranges'

# Add a character
my_fruit <- my_apples + my_oranges

```

`Error in my_apples + my_oranges : non-numeric argument to binary operator`

### Basic data types in R

```{r}

# What's the answer to the universe
my_numeric <- 42

# The quotation marks indicate that the variable is of type character
my_character <- "forty-two"

my_logical <- FALSE


```

### Whats's that data type?

```{r}

# Declare variables of different types
my_numeric <- 42
class(my_numeric)

my_character <- "forty-two"
class(my_character)

my_logical <- FALSE
class(my_logical)


```

&nbsp; &nbsp;

***  

# Vectors

  - Note: I almost **always** use dataframes
  - But I guess its still good to know the other data types 
  - They all have their uses

### Create a vector

```{r}

Vegas <- as.character("Here we go!")
Vegas

```

### Create a vector (2)

```{r}

numeric_vector <- c(1, 10, 49)
character_vector <- c("a", "b", "c")
boolean_vector <- c(T,F,T)

# Print the vectors
numeric_vector
character_vector
boolean_vector

```

### Create a vector (3)

```{r}

# Poker winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)
poker_vector

# Roulette winnings form Monday to Friday
roulette_vector <- c(-24, -50, 100, -350, 10)
roulette_vector

```

### Naming a vector

```{r}

## Name the vectors
names(poker_vector) <- c("Mon","Tues","Wed","Thur","Fri")
poker_vector

names(roulette_vector) <- c("Mon","Tues","Wed","Thur","Fri")
roulette_vector

```

### Naming a vector (2)

```{r}

# Create the variable days_vector
days_vector <- c("Mon","Tues","Wed","Thur","Fri")

# Assign the names of the day to the roulette and poker_vectors
names(poker_vector) <- days_vector
poker_vector

names(roulette_vector) <- days_vector
roulette_vector

```

### Calculating total winnings

```{r}
## First, just an example
A_vector <- c(1, 2, 3)
B_vector <- c(4, 5, 6)

# Take the sum of A_vector and B_vector
total_vector <- A_vector + B_vector
total_vector

```

### Calculating total winnings (2)

```{r}

# Name poker and roulette
days <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(roulette_vector) <- days
names(poker_vector) <- days

total_daily <- poker_vector + roulette_vector
total_daily

```

### Calculating total winnings (3)

```{r}

total_poker <- sum(poker_vector)
total_roulette <- sum(roulette_vector)

total_week <- sum(roulette_vector) + sum(poker_vector)
total_week

```

### Comparing total winnings

```{r}

# Check if you realized higher total gains in poker then in roulette
answer <- total_poker > total_roulette
answer

```

### Vector selection: the good times

```{r}

# Define new variable based on a selection
poker_wednesday <- poker_vector["Wednesday"]
poker_wednesday

```

### Vector selection: the good times (2)

```{r}

# Define new variable based on a selection
poker_midweek <- poker_vector[c("Tuesday","Wednesday","Thursday")]       
poker_midweek 

```

### Vector selection: the good times (3)

```{r}

# Define new variable based on a selection
roulette_selection_vector <- roulette_vector[2:5]
roulette_selection_vector 

```

### Vector selection: the good times (4)

```{r}

average_midweek_gain <- mean(poker_vector[c("Monday","Tuesday","Wednesday")])
average_midweek_gain

```

### Selection by comparison - Step 1

```{r}

# What days of the week did you make money on poker
selection_vector <- poker_vector > 0
selection_vector

```

### Selection by comparison - Step 2

```{r}

# What days of the week did you make money on poker
selection_vector <- poker_vector > 0

# Select from poker_vector these days
poker_winning_days <- poker_vector[selection_vector]
poker_winning_days

```

### Advanced selection

```{r}

# Show me
roulette_winning_days <- roulette_vector[roulette_vector > 0]
roulette_winning_days

```

&nbsp; &nbsp;

***

# Matrices

### What's a matrix?

```{r}

# Construction of a matrix with 3 rows containing the numbers 1 up to 9
matrix(1:9, byrow=T, nrow=3)

```

### Analyzing matrices, you shall

```{r}

# Box office Star Wars: In Millions!
# The first element: US, the second element: Non-US 
new_hope = c(460.998007, 314.4)
empire_strikes = c(290.475067, 247.900000)
return_jedi = c(309.306177,165.8)

# Add your code below to construct the matrix
star_wars_matrix = matrix(c(new_hope, empire_strikes, return_jedi), byrow=T, nrow=3)
star_wars_matrix

```

### Naming a matrix

```{r}

# Add your code here such that rows and columns of star_wars_matrix have a
# name!
rownames(star_wars_matrix) = c('A New Hope','The Empire Strikes Back','Return of the Jedi')
colnames(star_wars_matrix) = c('US','non-US')
star_wars_matrix

```

### Calculating the worldwide box office

```{r}

# Box office Star Wars: In Millions (!) 
# Construct matrix: 
box_office_all = c(461, 314.4,290.5, 247.9,309.3,165.8)
movie_names = c("A New Hope","The Empire Strikes Back","Return of the Jedi")
col_titles = c("US","non-US")

star_wars_matrix = matrix(box_office_all, nrow=3, byrow=TRUE, dimnames=list(movie_names,col_titles))
star_wars_matrix

worldwide_vector = rowSums(star_wars_matrix)
worldwide_vector
```

### Adding a column for the worldwide box office

```{r}

# Bind the new variable worldwide_vector as a column to star_wars_matrix
all_wars_matrix = cbind(star_wars_matrix, worldwide_vector)
all_wars_matrix

```

### Adding a row
  - In the course they had the star_wars_matrix2 pre loaded
  - Here I just copy the original
  - this works to show what rbind does
  
```{r}

# Matrix containing first trilogy box office
star_wars_matrix  

# Create a Matrix containing second trilogy box office
box_office_all = c(474.5, 552.5, 310.7, 338.7, 380.3, 468.5)
movie_names = c("The Phantom Menace","Attack of the Clones","Revenge of the Sixth")
col_titles = c("US","non-US")

star_wars_matrix2 = matrix(box_office_all, nrow=3, byrow=TRUE, dimnames=list(movie_names,col_titles))
star_wars_matrix2

# Combine both Star Wars trilogies in one matrix
all_wars_matrix = rbind(star_wars_matrix, star_wars_matrix2)
all_wars_matrix

```

### The total box office revenue for the entire saga

```{r}

# Print box office Star Wars: In Millions (!) for 2 trilogies
all_wars_matrix

total_revenue_vector = colSums(all_wars_matrix)
total_revenue_vector

```

### Selection of matix elements

```{r}

mean_non_us_all  =  mean(star_wars_matrix[,2])
mean_non_us_all

mean_non_us_some = mean(star_wars_matrix[1:2,2])
mean_non_us_some

```

### A little arithmetic with matrices

```{r}

visitors = star_wars_matrix/5
visitors

```

### A little arithmetic with matrices (2)

```{r}

ticket_prices_matrix = matrix(c(5,5,6,6,7,7), nrow=3, byrow=TRUE, dimnames=list(movie_names,col_titles)) 
ticket_prices_matrix

visitors = star_wars_matrix/ticket_prices_matrix
visitors

average_us_visitor = mean(visitors[,1])
average_us_visitor

average_non_us_visitor = mean(visitors[,2])
average_non_us_visitor

```

&nbsp; &nbsp;

***

# Factors

### What's a factor and why would you use it?

R uses factors for categorical variables!

### What's a factor and why would you use it? (2)

```{r}

gender_vector = c("Male","Female","Female","Male","Male")

factor_gender_vector = factor(gender_vector)
factor_gender_vector

```

### What's a factor and why would you use it? (3)

```{r}

animals_vector = c("Elephant", "Giraffe", "Donkey", "Horse")
factor_animals_vector = factor(animals_vector)
factor_animals_vector

temperature_vector = c("High", "Low", "High", "Low", "Medium")
factor_temperature_vector = factor(temperature_vector, order = TRUE, levels = c("Low", 
    "Medium", "High"), labels = c("L","M","H"))
factor_temperature_vector

```

### Factor levels

```{r}

survey_vector = c("M", "F", "F", "M", "M")
factor_survey_vector = factor(survey_vector)
factor_survey_vector

levels(factor_survey_vector) = c("Female","Male")
factor_survey_vector

```

### Summarizing a factor

```{r}

# Type your code here for survey_vector
summary(survey_vector)

# Type your code here for factor_survey_vector
summary(factor_survey_vector)

```

### Battle of the sexes

```{r}

# Male
factor_survey_vector[1] 

# Female
factor_survey_vector[2] 

# Battle of the sexes: Male 'larger' than female?
factor_survey_vector[1] > factor_survey_vector[2] 

```

### Ordered Factors

```{r}

# Create 'speed_vector'
speed_vector <- c('Fast','Slow','Slow','Fast','Ultra-fast')

factor_speed_vector <- factor(speed_vector, ordered = T, levels = c('Slow','Fast','Ultra-fast'))
factor_speed_vector

# R prints automagically in the right order
summary(factor_speed_vector) 

```


### Comparing ordered factors

```{r}

# Is data analyst 2 faster than data analyst 5?
compare_them <- speed_vector[2] > speed_vector[5]
compare_them

```

&nbsp; &nbsp;

***

# Data frames

### What's a data frame?

```{r}

mtcars  # Built-in R dataset stored in a data frame

```

### Quick, hae a look at your data set

```{r}

# Have a quick look at your data
head(mtcars)

```

### Have a look at the structure

```{r}

# Investigate the structure of the mtcars dataset to get started!
str(mtcars)

```

### Create a data frame

```{r}

planets = c("Mercury", "Venus", "Earth", "Mars", "Jupiter", "Saturn", "Uranus", 
    "Neptune")

type = c("Terrestrial planet", "Terrestrial planet", "Terrestrial planet", "Terrestrial planet", 
    "Gas giant", "Gas giant", "Gas giant", "Gas giant")

diameter = c(0.382, 0.949, 1, 0.532, 11.209, 9.449, 4.007, 3.883)

rotation = c(58.64, -243.02, 1, 1.03, 0.41, 0.43, -0.72, 0.67)

rings = c(FALSE, FALSE, FALSE, FALSE, TRUE, TRUE, TRUE, TRUE)

# Create the data frame:
planets_df = data.frame(planets, type, diameter, rotation, rings)
planets_df

```

### Create a data frame (2)

```{r}

# Check the structure of planets.df
str(planets_df)

```

### Selection of data frame elements

```{r}

# The planets_df data frame from the previous exercise is pre-loaded
closest_planets_df <- planets_df[1:3,]
closest_planets_df

furthest_planets_df <- planets_df[6:8,]
furthest_planets_df

```

### Selection of data frame elements (2)

```{r}

# The planets_df data frame from the previous exercise is pre-loaded: 
furthest_planets_diameter = planets_df$diameter[3:8]

```

### Only planets with rings

```{r}

# Create the rings_vector
rings_vector = planets_df$rings

```

### Only planets with rings (2)

```{r}

# Select the information on planets with rings:
planets_with_rings_df =  planets_df[planets_df$rings,]

```

### Only planets with rings but shorter

```{r}

# Planets smaller than earth:
small_planets_df  = subset(planets_df, diameter < 1)

```

### Sorting

```{r}

# Just play around with the order function in the console to see how it
# works
a <- c(100,9,101)
order(a)

```

### Sorting your data frame

```{r}

# What is the correct ordering based on the planets_df$diameter variable?
positions =  order(planets_df$diameter, decreasing = T)
positions

# Create new "ordered" data frame:
largest_first_df = planets_df[positions,]
largest_first_df

```

&nbsp; &nbsp;

***

# Lists

### Lists, why would you need them?

They are useful sometimes.

### Lists, why would you need them? (2)

  - A list in R is similar to your to-do list at work or school: 
    - the different items on that list most likely differ in length, characteristic, type of activity that has to do be done, ...
  - A list in R allows you to gather a variety of objects under one name (that is, the name of the list) in an ordered way. 
    - These objects can be matrices, vectors, data frames, even other lists, etc. 
    - It is not even required that these objects are related to each other in any way.
  - You could say that a list is some kind super data type: 
    - you can store practically any piece of information in it!

### Creating a list

```{r}

# Vector with numerics from 1 up to 10
my_vector <- 1:10 
my_vector

# Matrix with numerics from 1 up to 9
my_matrix <- matrix(1:9, ncol = 3)
my_matrix

# First 10 elements of the built-in data frame 'mtcars'
my_df <- mtcars[1:10,]
my_df

# Construct list with these different elements:
my_list <- list(my_vector, my_matrix, my_df)
my_list

```

### Creating a named list

```{r}

# Construct 'my_list' with these different elements:
my_list <- list(vec = my_vector, mat = my_matrix, df = my_df)

# Print 'my_list' to the console
my_list

```

### Creating a named list (2)

```{r}

# The vectors 'actors' and 'reviews' are pre-loaded in the workspace
actors <- c("Jack Nicholson", "Shelley Duvall", "Danny Lloyd", "Scatman Crothers", "Barry Nelson")

reviews <- data.frame(
  scores = c(4.5, 4.0, 5.0),
  sources = c("IMDb1", "IMDb2", "IMDb3"),
  comments = c("Best Horror Film I Have Ever Seen",
               "A truly brilliant and scary film from Stanley Kubrick",
               "A masterpiece of psychological horror")
)

# Create the list 'shining_list'
shining_list <- list(
    moviename = "The Shining",
    actors = actors,
    reviews = reviews)

shining_list
```

### Selecting elements from a list

```{r}

# Define 'last_actor'
last_actor <- shining_list$actors[length(shining_list$actors)]
last_actor

# Define 'second_review'
second_review <- shining_list$reviews[2,]
second_review

```

### Adding more movie information to the list

```{r}

# We forgot something; add the year to shining_list
shining_list_full <- c(shining_list, year = 1980)

# Have a look at shining_list_full
str(shining_list_full)

```
