---
title: "dplyr"
author: "William Surles"
date: "2017-03-29"
output: 
 html_document:
  self_contained: yes
  theme: flatly
  highlight: tango
---

***
> This document was updated on **`r Sys.Date()`**.

***

```{r setup, include=FALSE}
library(shiny)
library(codetools)
knitr::opts_chunk$set(eval=T, echo=T, cache=T, message=F, warning=F)
```

# Introduction

  - This all comes from the [Data Manipulation in R with dplyr](https://www.datacamp.com/courses/dplyr-data-manipulation-r-tutorial) course on DataCamp
    - I highly recommend the [DataCamp](https://www.datacamp.com/home) courses
    - Go *do* the [course](https://www.datacamp.com/courses/dplyr-data-manipulation-r-tutorial) for yourself to get more details and videos on these topics
  - I am copying my work from the class to this document and adding comments

## Whats Covered

  - Conditionals and Control Flow
    - Relational Operators. `==` `<` `>` 
    - Logical Operators. `and` `or` `not`
    - Conditional Statements. `if` `else`
  - Loops
    - While loops
    - For loops
    - Breaks
  - Functions
    - Documentation
    - Required vs optional inputs
    - Scoping
    - Writing functions
    - Loading packages
  - Apply statements
    - lapply, sapply, vapply
  - Utilities
    - Useful base functions
    - Regular Expressions and grep/sub
    - Times and Dates
    
  
  
&nbsp; &nbsp;

***  

# Introduction to dplyr and tbls
# Select and mutate
# Filter and arrange
# Summarise and the pipe operator
# Group_by and working with databases
# 


```{r}
```




