---
title: "Data Manipulation in R with dplyr"
author: "William Surles"
date: "2017-07-19"
output: 
 html_document:
  self_contained: yes
  theme: flatly
  highlight: tango
  toc: true
  toc_float: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(eval=T, echo=T, cache=T, message=F, warning=F)


# library(shiny)
# library(codetools)

```

# Introduction

  - This all comes from the [Data Manipulation in R with dplyr](https://www.datacamp.com/courses/dplyr-data-manipulation-r-tutorial) course on DataCamp
    - I highly recommend the [DataCamp](https://www.datacamp.com/home) courses
    - Go *do* the [course](https://www.datacamp.com/courses/dplyr-data-manipulation-r-tutorial) for yourself to get more details and videos on these topics
  - I am copying my work from the class to this document and adding comments for my own future reference

## Whats Covered

  - Introduction to dplyr and tbls
  - Select and mutate
  - Filter and arrange
  - Summarise and the pipe operator
  - Group_by and working with databases
    
## Aditional Resources

  - [dplyr vignette](http://dplyr.tidyverse.org/)
  - [Data Wrangling Cheat Sheet (dplyr and tidyr)](https://www.rstudio.com/wp-content/uploads/2015/02/data-wrangling-cheatsheet.pdf)
  
&nbsp; &nbsp;

***  
# Introduction to dplyr and tbls
***  

## Section 1 - Introduction to dplyr

  - dplyr is a grammar of data manipulation
  - it provides a consistent set of vebs that help you solve the most common data manipulation challenges
  - `mutate`, `select`, `filter`, `summarise`, `arrange`, and the joins
  - [dplyr vignette](http://dplyr.tidyverse.org/)

## Load the dplyr and hflights package

```{r}

# Load the dplyr package
library(dplyr)

# Load the hflights package
library(hflights)

# Call both head() and summary() on hflights
head(hflights)
summary(hflights)

```

## Explore the dataset

```{r}

str(hflights)

```

## Section 2 - tbl, a special type of data.frame

  - tibble print adapts to the size of your window
  - glimpse gives you a more complete view of the tibble
  - if you don't like it go back to data.frame and use `str` and `head`

## Convert data.frame to tibble

```{r}

# Convert the hflights data.frame into a hflights tbl
hflights <- tbl_df(hflights)

# Display the hflights tbl
hflights
glimpse(hflights)

# Create the object carriers, containing only the UniqueCarrier variable of hflights
carriers <- hflights$UniqueCarrier
str(carriers)

```

## Changing labels of hflight, part 1 of 2

```{r}

# Both the dplyr and hflights packages are loaded into workspace
lut <- c("AA" = "American", "AS" = "Alaska", "B6" = "JetBlue", "CO" = "Continental", 
         "DL" = "Delta", "OO" = "SkyWest", "UA" = "United", "US" = "US_Airways", 
         "WN" = "Southwest", "EV" = "Atlantic_Southeast", "F9" = "Frontier", 
         "FL" = "AirTran", "MQ" = "American_Eagle", "XE" = "ExpressJet", "YV" = "Mesa")

# Use lut to translate the UniqueCarrier column of hflights
hflights$UniqueCarrier <- lut[hflights$UniqueCarrier]

# Inspect the resulting raw values of your variables
glimpse(hflights)

```

## Changing labels of hflights, part 2 or 2

```{r}
## The "E" in my dataset is "" for some reason so I have changed it here
table(hflights$CancellationCode)

hflights <- hflights %>%
  mutate(
    CancellationCode = ifelse(CancellationCode == "", "E", CancellationCode)
  )

table(hflights$CancellationCode)

# Build the lookup table: lut
lut <- c("A" = "carrier",
         "B" = "weather",
         "C" = "FFA",
         "D" = "security",
         "E" = "not cancelled")

# Use the lookup table to create a vector of code labels. Assign the vector to the CancellationCode column of hflights
hflights$Code <- lut[hflights$CancellationCode]

# Inspect the resulting raw values of your variables
glimpse(hflights)

```

&nbsp; &nbsp;

***  
# Select and mutate
***  

## Section 3 - The five verbs and select in more detail

  - 5 verbs
    - `select()` which returns a subset of the columns,
    - `filter()` that is able to return a subset of the rows,
    - `arrange()` that reorders the rows according to single or multiple variables,
    - `mutate()` used to add columns from existing data,
    - `summarise()` which reduces each group to a single row by calculating aggregate measures.
  - verb focus
    - `select` and `mutate` manipulate variables
    - `filter` and `arrange` manipulate observations
    - `summarize` manipulates `groups` of observations

## Choosing is not losing! The select verb

```{r}

# Print out a tbl with the four columns of hflights related to delay
select(hflights, ActualElapsedTime, AirTime, ArrDelay, DepDelay)

# Print out hflights, nothing has changed!
hflights

# Print out the columns Origin up to Cancelled of hflights
select(hflights, Origin:Cancelled)

# Answer to last question: be concise!
select(hflights, 1:4, 12:21)

```

## Helper functions for variable selection

```{r}

# Print out a tbl containing just ArrDelay and DepDelay
select(hflights, ends_with(c('Delay')))

# Print out a tbl as described in the second instruction, using both helper functions and variable names
select(hflights, UniqueCarrier, ends_with(c('Num')), starts_with(c('Cancel')))

# Print out a tbl as described in the third instruction, using only helper functions.
select(hflights, contains(c('Time')), contains(c('Delay')))

```

## Comparison to base R

```{r}

ex1r <- hflights[c("TaxiIn","TaxiOut","Distance")]
ex1d <- select(hflights, contains(c('Taxi')), Distance)

ex2r <- hflights[c("Year","Month","DayOfWeek","DepTime","ArrTime")]
ex2d <- select(hflights, Year:ArrTime, -3)

ex3r <- hflights[c("TailNum","TaxiIn","TaxiOut")]
ex3d <- select(hflights, TailNum, contains(c('Taxi')))

```

## Section 4 - The second of five verbs: mutate

## Mutating is creating

```{r}

# Add the new variable ActualGroundTime to a copy of hflights and save the result as g1.
g1 <- mutate(hflights, ActualGroundTime = ActualElapsedTime - AirTime)

# Add the new variable GroundTime to a g1. Save the result as g2.
g2 <- mutate(g1, GroundTime = TaxiIn + TaxiOut)

# Add the new variable AverageSpeed to g2. Save the result as g3.
g3 <- mutate(g2, AverageSpeed = Distance / AirTime * 60)

# Print out g3
glimpse(g3)

```

## Add multiple variables using mutate

```{r}
# hflights and dplyr are ready, are you?

# Add a second variable loss_percent to the dataset: m1
m1 <- mutate(hflights, 
        loss = ArrDelay - DepDelay,
        loss_percent = (ArrDelay - DepDelay)/DepDelay * 100
        )
glimpse(m1)

# Copy and adapt the previous command to reduce redendancy: m2
m2 <- mutate(hflights, 
        loss = ArrDelay - DepDelay,
        loss_percent = loss/DepDelay * 100
        )
glimpse(m2)

# Add the three variables as described in the third instruction: m3
m3 <- mutate(hflights, 
        TotalTaxi = TaxiIn + TaxiOut,
        ActualGroundTime = ActualElapsedTime - AirTime,
        Diff = TotalTaxi - ActualGroundTime
        )
glimpse(m3)
```

&nbsp; &nbsp;

***  
# Filter and arrange
***  

## Section 5 - The third of five verbs: filter

## Logical operators

```{r}

# All flights that traveled 3000 miles or more
filter(hflights, Distance >= 3000) %>% glimpse()

# All flights flown by one of JetBlue, Southwest, or Delta
filter(hflights, UniqueCarrier %in% c('JetBlue', 'Southwest', 'Delta')) %>% glimpse()

# All flights where taxiing took longer than flying
filter(hflights, (TaxiIn + TaxiOut) > AirTime) %>% glimpse()

```

## Combining tests using boolean operators

```{r}

# All flights that departed before 5am or arrived after 10pm
filter(hflights, DepTime < 500 | ArrTime > 2200) %>% glimpse()

# All flights that departed late but arrived ahead of schedule
filter(hflights, DepDelay > 0 & ArrDelay < 0) %>% glimpse()

# All cancelled weekend flights
filter(hflights, Cancelled == 1 & DayOfWeek %in% c(6,7)) %>% glimpse()

# All flights that were cancelled after being delayed
filter(hflights, DepDelay > 0 & Cancelled == 1) %>% glimpse()

```

## Blend together what you've learned!

```{r}

# Select the flights that had JFK as their destination: c1
c1 <- filter(hflights, Dest == 'JFK')

# Combine the Year, Month and DayofMonth variables to create a Date column: c2
c2 <- mutate(c1, Date = paste(Year,Month,DayofMonth, sep="-"))

# Print out a selection of columns of c2
select(c2, Date, DepTime, ArrTime, TailNum)

```

## Recap on select, mutate and filter

  - How many weekend flights flew a distance of more than 1000 miles but had a total taxiing time below 15 minutes?
  
```{r}

hflights %>%
  filter(
    Distance > 1000,
    DayOfWeek > 5,
    TaxiIn + TaxiOut < 15
    ) %>%
  glimpse()
    
```

  - In this dataset it is 1,739 flights
  - In the class the answer was 155
  - I think they just have the data filtered to one city (Houston)

## Section 6 - Almost there: the arrange verb

## Arranging your data

```{r}

# Definition of dtc
dtc <- filter(hflights, Cancelled == 1, !is.na(DepDelay))

# Arrange dtc by departure delays
arrange(dtc, DepDelay) %>% glimpse()

# Arrange dtc so that cancellation reasons are grouped
arrange(dtc, CancellationCode) %>% glimpse()

# Arrange dtc according to carrier and departure delays
arrange(dtc, UniqueCarrier, DepDelay) %>% glimpse()

```

## Reverse the order of arranging

```{r}

# Arrange according to carrier and decreasing departure delays
arrange(hflights, UniqueCarrier, desc(DepDelay)) %>% glimpse()

# Arrange flights by total delay (normal order).
arrange(hflights, (DepDelay + ArrDelay)) %>% glimpse()

# Keep flights leaving to DFW before 8am and arrange according to decreasing AirTime 
hflights %>% filter(Dest == 'DFW', DepTime < 800) %>% arrange(desc(AirTime)) %>% glimpse()

```

&nbsp; &nbsp;

***  
# Summarise and the pipe operator
***  

## Section 7 - Last but not least: summarise

## The syntax of summarize

```{r}

glimpse(hflights)

# Print out a summary with variables min_dist and max_dist
summarize(hflights, min_dist = min(Distance), max_dist = max(Distance))

# Print out a summary with variable max_div
hflights %>% filter(Diverted == 1) %>% summarize(max_div = max(Distance))

```

## Aggregate functions

Aggregate functions defined in R:

  - `min(x)` - minimum value of vector x
  - `max(x)` - maximum value of vector x
  - `mean(x)` - mean value of vector x
  - `median(x)` - median value of vector x
  - `quantile(x, p)` - pth quantile of vector x
  - `sd(x)` - standard deviation of vector x
  - `var(x)` - variance of vector x
  - `IQR(x)` - Inter Quartile Range (IQR) of vector x
  - `diff(range(x))` - total range of vector x
  
```{r}

# Remove rows that have NA ArrDelay: temp1
temp1 <- filter(hflights, !is.na(ArrDelay))

# Generate summary about ArrDelay column of temp1
summarize(temp1, 
  earliest = min(ArrDelay),
  average = mean(ArrDelay),
  latest = max(ArrDelay),
  sd = sd(ArrDelay)
  )

# Keep rows that have no NA TaxiIn and no NA TaxiOut: temp2
temp2 <- filter(hflights, !is.na(TaxiIn) & !is.na(TaxiOut))

# Print the maximum taxiing difference of temp2 with summarise()
summarise(temp2, max_taxi_diff = max(abs(TaxiIn - TaxiOut)))

```

## dplyr aggregate functions

dplyr has some of its own aggregate functions:

  - `first(x)` - The first element of vector x
  - `last(x)` - The last element of vector x
  - `nth(x, n)` - The nth element of vector x
  - `n()` - The number of rows in the data.frame or group of observations that summarise() describes
  - `n_distinct(x)` - The number of unique values in vector x
  
```{r}

# Generate summarizing statistics for hflights
summarise(hflights, 
  n_obs = n(),
  n_carrier = n_distinct(UniqueCarrier),
  n_dest = n_distinct(Dest),
  dest100 = nth(Dest, 100)
  )

# Filter hflights to keep all American Airline flights: aa
aa <- filter(hflights, UniqueCarrier == 'American')

# Generate summarizing statistics for aa 
summarise(aa, 
  n_flights = n(),
  n_canc = sum(Cancelled),
  p_canc = n_canc/n_flights * 100,
  avg_delay = mean(ArrDelay, na.rm=T)
  )

```

## Section 8 - Chaining your functions: the pipe operator

  - the pipe operator `%>%` is probably my favorite thing in R
  - it takes the output from the last function and passes it into the next function as the first argument
    - But you can also use a `.` to pass it into the next function in any location!
  - This lets you make code easily readable from left to right and top to bottom. 
    - I can't say enough about how much better this makes it to read code and see what someone is doing
    
## Overview of syntax

```{r}

# Write the 'piped' version of the English sentences.
hflights %>%
  mutate(diff = TaxiOut - TaxiIn) %>%
  filter(!is.na(diff)) %>%
  summarize(avg = mean(diff))

```

## Drive of fly? Part 1 of 2

```{r}

# Build data frame with 4 columns of hflights and 2 self-defined columns: d
d <- hflights %>%
  select(Dest, UniqueCarrier, Distance, ActualElapsedTime) %>%
  mutate(
    RealTime = ActualElapsedTime + 100,
    mph = Distance/RealTime*60
    )

# Filter and summarise d according to the instructions
d %>%
  filter(
    !is.na(mph),
    mph < 70
    ) %>%
  summarize(
    n_less = n(),
    n_dest = n_distinct(Dest),
    min_dist = min(Distance),
    max_dist = max(Distance)
    )
  
```

## Drive or fly? Part 2 of 2

```{r}

# Solve the exercise using a combination of dplyr verbs and %>%
hflights %>%
  mutate(RealTime = ActualElapsedTime + 100, mph = Distance / RealTime * 60) %>%
  filter(mph < 105 | Cancelled == 1 | Diverted == 1) %>%
  summarise(n_non = n(), 
            p_non = n_non / nrow(hflights) * 100, 
            n_dest = n_distinct(Dest), 
            min_dist = min (Distance), 
            max_dist = max(Distance))

```

## Advanced piping exercise

```{r}

hflights %>%
  filter(
    !is.na(ArrTime),
    !is.na(DepTime),
    ArrTime < DepTime
    ) %>%
  summarise(n = n())

```


&nbsp; &nbsp;

***  
# Group_by and working with databases
***

## Section 9 - get group-wise insights: group_by

  - Combining `group_by` with `summarize` is very powerful
    - You can also combine it with `mutate` and `arrange` to create powerful window functions
    
## Unite and conquer using group_by

```{r}

# Make an ordered per-carrier summary of hflights
hflights %>%
 group_by(UniqueCarrier) %>%
 summarise(
   n_flights = n(),
   n_canc = sum(Cancelled),
   p_canc = n_canc/n_flights * 100,
   avg_delay = mean(ArrDelay, na.rm=T)
   ) %>%
 arrange(avg_delay, p_canc)

# Make an ordered per-day summary of hflights
hflights %>%
  group_by(DayOfWeek) %>%
  summarize(avg_taxi = mean(TaxiIn + TaxiOut, na.rm=T)) %>%
  arrange(desc(avg_taxi))

```

## Combine group_by with mutate

```{r}

# Solution to first instruction
hflights %>%
  filter(!is.na(ArrDelay)) %>%
  group_by(UniqueCarrier) %>%
  summarize(p_delay = sum(ArrDelay > 0)/n()) %>%
  mutate(rank = rank(p_delay)) %>%
  arrange(rank)

# Solution to second instruction
hflights %>%
  filter(
    !is.na(ArrDelay),
    ArrDelay > 0
    ) %>%
  group_by(UniqueCarrier) %>%
  summarize(avg = mean(ArrDelay)) %>%
  mutate(rank = rank(avg)) %>%
  arrange(rank)

```

## Advanced group_by exercises

```{r}

# Which plane (by tail number) flew out of Houston the most times? How many times? adv1
adv1 <- hflights %>%
  group_by(TailNum) %>%
  summarize(n = n()) %>%
  filter(n == max(n))
adv1

# How many airplanes only flew to one destination from Houston? adv2
adv2 <- hflights %>%
  group_by(TailNum) %>%
  summarize(n_dest = n_distinct(Dest)) %>%
  filter(n_dest == 1) %>%
  summarize(nplanes = n())
adv2

# Find the most visited destination for each carrier: adv3
adv3 <- hflights %>%
  group_by(UniqueCarrier, Dest) %>%
  summarize(n = n()) %>%
  group_by(UniqueCarrier) %>%
  mutate(rank = rank(desc(n))) %>%
  filter(rank == 1) %>%
  arrange(UniqueCarrier, rank)
adv3

# Find the carrier that travels to each destination the most: adv4
adv4 <- hflights %>%
  group_by(Dest, UniqueCarrier) %>%
  summarize(n = n()) %>%
  group_by(Dest) %>%
  mutate(rank = rank(desc(n))) %>%
  filter(rank == 1)
adv4  
  
```

## Section 10 - dplyr and databases

  - dplyr can connect to a database
    - You can manipulate the data in the dabase (query essentially) and then only pull back the result into R
    - This lets you work with much larger datasets stored in a relational database than you could on your local machine or having to augment R with hadoop
  - I used this functionality a lot at work.
    - Getting comfortable with this also help when using spark and sparklyr. 
    - It really nice to have one consistent way to manipulate data where ever its stored: 
      - locally in the workspace, 
      - a relational database
      - in HDFS
      - any file system accessed with spark

## dplyr deals with different types
  
```{r}
library(data.table)

# Convert hflights to a data.table
class(hflights)
hflights2 <- as.data.table(hflights)
class(hflights2)

# Use summarise to calculate n_carrier
s2 <- hflights2 %>%
  summarize(n_carrier = n_distinct(UniqueCarrier))
s2
```

## dplyr and mySQL databases

```{r}

library(RMySQL)
library(dbplyr)

# Set up a connection to the mysql database
my_db <- src_mysql(dbname = "dplyr", 
                   host = "courses.csrrinzqubik.us-east-1.rds.amazonaws.com", 
                   port = 3306, 
                   user = "student",
                   password = "datacamp")

# Reference a table within that source: nycflights
nycflights <- tbl(my_db, "dplyr")

# glimpse at nycflights
glimpse(nycflights)

# Ordered, grouped summary of nycflights
nycflights %>%
   group_by(carrier) %>%
   summarise(n_flights = n(), avg_delay = mean(arr_delay)) %>%
   arrange(avg_delay)
 
```

## Talk with Hadley Wickham

  - Two goals
    - Make it easier to think about data manipulation. What are the fundamental verbs
    - Compute efficiently with the data. It uses C++. It can generate SQL for you and send to database
  - plyr was about using split, apply, combine
    - dplyr focuses on just data frames, but thats what most people use anyways
  - Learn about tidy data
  - Get a dataset you are motivated by and start playing with it
  - Get familiar with window functions
    - There are a wide class a problems that can be solved by window functions
    