---
title: "Writing Functions in R"
author: "William Surles"
date: "2017-07-17"
output: 
 html_document:
  self_contained: yes
  theme: flatly
  highlight: tango
  toc: true
  toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval=T, echo=T, cache=T, message=F, warning=F)

source('create_datasets.R')

```

***
# Introduction
***

Course notes for [Writing Functions in R](https://www.datacamp.com/courses/writing-functions-in-r)

## Whats Covered

  - A quick refresher
  - When and how you should write a function
  - Functional programming
  - Advanced inputs and outputs
  - Robust functions

## Additional Resources

  - [Nonstandard Evaluation Vignette](http://rpubs.com/hadley/157957)
  - [purr vignette](http://purrr.tidyverse.org/)
  
&nbsp; &nbsp;

***  
# A quick refresher
***  

## Writing a function in R

  - The parts of a function
    - Argument
    - Body
    - Environment
  - Return value is the last evaluated expression or the first evaluated `return()` expression
  - Functions can be treated like usual R objects
  
## Writing a function

```{r}

# Define ratio() function
ratio <- function(x, y) {
  x / y
}

# Call ratio() with arguments 3 and 4
ratio(3,4)

```

## Arguments

```{r}

# Rewrite the call to follow best practices
mean(c(1:9, NA), trim = 0.1, na.rm = TRUE)

```

## Function output

```{r}

f <- function(x) {
  if (TRUE) {
    return(x + 1)
  }
  x
}

f(2)
```

## Environments

  - When you call a function, a new environment is made for the function to do its work
  - The new environment is populated with the argument values
  - Objects are looked for first in this environment
  - If they are not found they are looked for in the environmnet that the function was created in
  
## Testing your understanding of scoping (1)

```{r}

y <- 10
f <- function(x) {
  x + y
}
f(10)

```

## Testing your understanding of scoping (2)

```{r}

y <- 10
f <- function(x) {
  y <- 5
  x + y
}
f(10)

```

## Testing your understanding of scoping (3)

```{r}

f <- function(x) {
  y <- 5
  x + y
}
f(5)

```

## Data structures

  - Two types of vectors
    - Atomic vectors of six types: logical, integer, double, character, complex, raw
    - Lists (recursive vectors), because lists can contain other lists
  - Vectors have two key properties
    - type and length
  - Missing values
    - `NULL` often used to indicate the absence of a vector
    - `NA` used to indicate the absence of a value in a vector, aka a missing value
    - Missing values are contaigous. i.e. calling `sum` on a vector with NA will result in NA.
  - Lists
    - Useful because they con contain heterogeneous objects
    - Complicated return objects are often lists, i.e. from lm()
    - Created with list()
    - Subset with `[`, `[[`, or `$`
      - `[` extracts a sublist
      - `[[` and `$` extract elements, remove a level of hierarchy

## Subsetting lists

```{r}

# 2nd element in tricky_list
typeof(tricky_list[[2]])

# Element called x in tricky_list
typeof(tricky_list[['x']])

# 2nd element inside the element called x in tricky_list
typeof(tricky_list[['x']][[2]])

```

## Exploring lists

```{r}

# Guess where the regression model is stored
names(tricky_list)

# Use names() and str() on the model element
names(tricky_list$model)
str(tricky_list$model)

# Subset the coefficients element
tricky_list$model$coefficients

# Subset the wt element
tricky_list$model$coefficients[['wt']]

```

## for loops

  - We have covered for loops in the intermediate R class
  - Her we cover:
    - A saver way to generate the sequence with `seq_along()`
    - Saving output instead of printing it
    
## A safter way to create the sequence

```{r}

# Replace the 1:ncol(df) sequence
for (i in seq_along(df1)) {
  print(median(df1[[i]]))
}

# Change the value of df
df <- data.frame()

# Repeat for loop to verify there is no error
for (i in seq_along(df)) {
  print(median(df[[i]]))
}

```

## Keeping output

```{r}

# Create new double vector: output
output <- vector("double", ncol(df1))

# Alter the loop
for (i in seq_along(df1)) {
  # Change code to store result in output
  output[i] <- median(df1[[i]])
}

# Print output
output

```

&nbsp; &nbsp;

***  
# When and how you should write a funciton
***  

## Why should you write a function?

  - When
    - When you have copied and pasted a piece of code twice
  - Why
    - Reduces mistakes from copying and pasting
    - Makes updating code easier
  
## Start with a snippet of code

```{r}

# Define example vector x
x <- 1:10

# Rewrite this snippet to refer to x
(x - min(x, na.rm = TRUE)) /
  (max(x, na.rm = TRUE) - min(x, na.rm = TRUE))

```

## Rewrite for clarity

```{r}

# Define example vector x
x <- 1:10

# Define rng
rng <- range(x, na.rm = T)

# Rewrite this snippet to refer to the elements of rng
(x - rng[1]) /
  (rng[2] - rng[1])

```

## Finally turn it into a function!

```{r}

# Define example vector x
x <- 1:10 

# Use the function template to create the rescale01 function
rescale01 <- function(x) {
  rng <- range(x, na.rm = TRUE) 
  (x - rng[1]) / (rng[2] - rng[1])
}

# Test your function, call rescale01 using the vector x as the argument
rescale01(x)

```

## How should you write a function

  - Start with a simple problem
  - Get a working snippet of code
  - Rewrite to use temporarty variables
  - Rewrite for clarity
  - Finanlly, turn into a function (wrap in curly braces and give it a name)

## Start with a simple problem

```{r}

# Define example vectors x and y
x <- c( 1, 2, NA, 3, NA)
y <- c(NA, 3, NA, 3,  4)

# Count how many elements are missing in both x and y
sum(is.na(x) & is.na(y))

```

## Rewrite snippet as function

```{r}

# Define example vectors x and y
x <- c( 1, 2, NA, 3, NA)
y <- c(NA, 3, NA, 3,  4)

# Turn this snippet into a function: both_na()

both_na <- function(x,y) {
  sum(is.na(x) & is.na(y))
}

```

## Put our function to use

```{r}

# Define x, y1 and y2
x <-  c(NA, NA, NA)
y1 <- c( 1, NA, NA)
y2 <- c( 1, NA, NA, NA)

# Call both_na on x, y1
both_na(x, y1)

# Call both_na on x, y2
both_na(x, y2)

```

## How can you write a good funciton?

  - Functions are for computers and humans
    - it should be correct AND understandable
  - Naming principles
    - Pick a consistent style for long names (lowercase with underscores)
    - Do not override existing variables or functions (T, `c`, `mean` )
  - Function names 
    - should generally be verbs and 
    - should be descriptive
    - e.g. `impute_mising`, `collapse_years`
  - Argument names 
    - should generally be nouns
    - use very common short names when appropriate
    - x, y, z : vectors
    - df : data frame
    - i, j : numeric indicies of rows and columns respectively
    - n, p : length or rows and columns respectively
  - Argument Order
    - Data arguments first
    - Detail arguments next
    - Detail arguments should have sensible defaults
    - Use an intuitive argumnet order
  - Make it clear what the function returns
  - Use good coding style in the body
    - Adopt an R syle guide

## Good function names

  - What should this function be called?
  
```{r}

f2 <- function(x) {
  if (length(x) <= 1) return(NULL)
  x[-length(x)]
}

```

  - `remove_last`
  
## Argument names

  - `c` is already they name of a function 
  - `nums` is non conventional
  
```{r}

# Rewrite mean_ci to take arguments named level and x, rather then c and nums,

mean_ci <- function(level, x) {
  se <- sd(x) / sqrt(length(x))
  alpha <- 1 - level
  mean(x) + se * qnorm(c(alpha / 2, 1 - alpha / 2))
}

```
 
## Argument order

  - Data arguments go first
  - Then detail arguments

```{r}

# Alter the arguments to mean_ci
mean_ci <- function(x, level = .95) {
  se <- sd(x) / sqrt(length(x))
  alpha <- 1 - level
  mean(x) + se * qnorm(c(alpha / 2, 1 - alpha / 2))
}

```
  
## Return statements

```{r}

# Alter the mean_ci function
# Edit the mean_ci function using an if statement to check for the case when x is empty and if so, to produce the same warning as the code above then immediately return() c(-Inf, Inf).

mean_ci <- function(x, level = 0.95) {
  
  if (length(x) == 0) {
    warning("x was empty", call. = FALSE)
    return(c(-Inf, Inf))
  }
  
  se <- sd(x) / sqrt(length(x))
  alpha <- 1 - level
  mean(x) + se * qnorm(c(alpha / 2, 1 - alpha / 2))
}

mean_ci(numeric(0))

```

## What does this function do?

```{r}

f <- function(x, y) {
  x[is.na(x)] <- y
  cat(sum(is.na(x)), y, "\n")
  x
}

# Define a numeric vector x with the values 1, 2, NA, 4 and 5
x <- c(1, 2, NA, 4, 5)

# Call f() with the arguments x = x and y = 3
f(x = x, y = 3)

# Call f() with the arguments x = x and y = 10
f(x = x, y = 10)

```
  - It replaces `NA`s in `x` with the value given as `y`
  
## Let's make it clear from its name

```{r}

# Rename the function f() to replace_missings()
replace_missings <- function(x, replacement) {
  # Change the name of the y argument to replacement
  x[is.na(x)] <- replacement
  cat(sum(is.na(x)), replacement, "\n")
  x
}

# Rewrite the call on df$z to match our new names
df <- structure(list(z = c(0.473582631011786, -0.371809943354702, -0.88562980458499, 
-0.768223158006804, 0.886983968322944, 0.238444716245814, -1.13439205742083, 
NA, NA, -1.27240097583594)), .Names = "z", row.names = c(NA, 
-10L), class = "data.frame")

df$z <- replace_missings(df$z, 0)
df$z
```

## Make the body more understandable

```{r}

replace_missings <- function(x, replacement) {
  # Define is_miss
  is_miss <- is.na(x)
  
  # Rewrite rest of function to refer to is_miss
  x[is_miss] <- replacement
  cat(sum(is_miss), replacement, "\n")
  x
}

```

## Much better! But a few more tweaks...

```{r}

replace_missings <- function(x, replacement) {
  is_miss <- is.na(x)
  x[is_miss] <- replacement
  
  # Rewrite to use message()
  message(sum(is_miss), " missings replaced by the value ", replacement)
  x
}

# Check your new function by running on df$z
replace_missings(df$z, 0)

```

&nbsp; &nbsp;

***  
# Functional Programming
***  

## Why functional programming?

  - For loops
    - Emphasises the objects and pattern of implementation
    - Hides actions
    - Its like reading pages in a cookbook
  - Functional programming
    - Gives equal weights to verbs and nouns
    - Abstracts away the details of implementation
    
## Using a for loop to remove duplication

```{r}

# Initialize output vector
output <- vector("double", ncol(df1))  

# Fill in the body of the for loop
for (i in seq_along(df1)) {            

  output[i] <- median(df1[[i]])

}

# View the result
output

```

## Turning the for loop into a function

```{r}

# Turn this code into col_median()
output <- vector("double", ncol(df))  

col_median <- function(df) {

  for (i in seq_along(df)) {            
    output[[i]] <- median(df[[i]])      
  }
  output
}

col_median(df1)

```

## What about column means?

```{r}

# Create col_mean() function to find column means
col_mean <- function(df) {
  output <- numeric(length(df))
  for (i in seq_along(df)) {
    output[[i]] <- mean(df[[i]])
  }
  output
}

col_mean(df1)

```

## What about column standard devitions?

```{r}

# Define col_sd() function
col_sd <- function(df) {
  output <- numeric(length(df))
  for (i in seq_along(df)) {
    output[[i]] <- sd(df[[i]])
  }
  output
}

col_sd(df1)

```

## Uh oh...time to write a function again

```{r}

# Add a second argument called power
f <- function(x, power) {
    # Edit the body to return absolute deviations raised to power
    abs(x - mean(x)) ^ power
}

```

## Using a function as an argument

```{r}

col_summary <- function(df, fun) {
  output <- vector("numeric", length(df))
  for (i in seq_along(df)) {
    output[[i]] <- fun(df[[i]])
  }
  output
}

# Find the column medians using col_median() and col_summary()
col_median(df1)
col_summary(df1, fun = median)


# Find the column means using col_mean() and col_summary()
col_mean(df1)
col_summary(df1, fun = mean)

# Find the column IQRs using col_summary()
col_summary(df1, fun = IQR)

```

## Introducing purrr

  - purrr is a functional programming toolset for r
    - [purr vignette](http://purrr.tidyverse.org/)
  - It has a bunch of functions for mapping functions to data
  - The map functions all work like this
    - Loop over a vector `.x`
    - Apply the function `.f` to each element
    - Return the results
  - There is one map function for each type
    - `map` returns a list
    - `map_dbl` returns a vector of doubles
    - `map_lgl` returns a vector of logicals
    - `map_int` same for integers
    - `map_chr` same for characters
  - It can handle different types of inputs
    - For data frames it will iterate over the columns
    - For lists it will iterate over the elements
    - For vectos it will iterate over the elements
  - Advantages
    - Handy shortcuts for specifying `.f`
    - More consistent than sapply, lapply functions
    
  
## The map functions

```{r}

# Load the purrr package
library(purrr)

head(df)

# Use map_dbl() to find column means
map_dbl(df, mean)

# Use map_dbl() to column medians
map_dbl(df, median)

# Use map_dbl() to find column standard deviations
map_dbl(df, sd)

```

## The ... argumnet to the map functions

```{r}

head(planes)

# Find the mean of each column
map_dbl(planes, mean)

# Find the mean of each column, excluding missing values
map_dbl(planes, mean, na.rm = T)

# Find the 5th percentile of each column, excluding missing values
map_dbl(planes, quantile, probs = .05, na.rm = T)

```

## Picking the right map function

```{r}

# Find the columns that are numeric
map_lgl(df3, is.numeric)

# Find the type of each column
map_chr(df3, typeof)

# Find a summary of each column
map(df3, summary)

```

## Shortcuts

### Function shortcuts

```{r}

## An existing function
map(df1, summary)

## An existing funciton you defined
map(df1, rescale01)

## An anonymous function defined on the fly
map(df1, function(x) sum(is.na(x)))

## An anonymous function defined using a formula shortcut
map(df1, ~ sum(is.na(.)))

```

### Shortscuts when `.f` is `[[`

```{r}

list_of_results <- list(
  list(a = 1, b = "A"),
  list(a = 2, b = "C"),
  list(a = 3, b = "D")
)

map_dbl(list_of_results, function(x) x[["a"]])

map_dbl(list_of_results, "a")

map_dbl(list_of_results, 1)

```

## Solve a simple problem first

```{r}

# Examine the structure of cyl
str(cyl)

# Extract the first element into four_cyls
four_cyls <- cyl[[1]]

# Fit a linear regression of mpg on wt using four_cyls
lm(four_cyls$mpg ~ four_cyls$wt)

```

## Using an anonymous function

```{r}

# Rewrite to call an anonymous function
map(cyl, function(df) lm(mpg ~ wt, data = df))

```


## Using a formula

```{r}

# Rewrite to use the formula shortcut instead
map(cyl, ~ lm(mpg ~ wt, data = .))

```

## Using a string

```{r}

# Save the result from the previous exercise to the variable models
models <- map(cyl, ~ lm(mpg ~ wt, data = .))

# Use map and coef to get the coefficients for each model: coefs
coefs <- map(models, coef)

# Use string shortcut to extract the wt coefficient 
map(coefs, "wt")

```

## using a numeric vector

```{r}

coefs <- map(models, coef)

# use map_dbl with the numeric shortcut to pull out the second element
map_dbl(coefs, 2)

```

## Putting it toegether with pipes

```{r}

# Define models (don't change)
models <- mtcars %>% 
  split(mtcars$cyl) %>%
  map(~ lm(mpg ~ wt, data = .))

# Rewrite to be a single command using pipes 
summaries <- models %>%
map(summary) %>%
map_dbl("r.squared")

summaries
```


&nbsp; &nbsp;

***  
# Advanced inputs and outputs
***  

## Dealing with failure

  - functions to help with failure cases
    - `safely()` captures the successful result or the error, always returns a list
    - `possible()` always succeeds, you give it a default value to return when there is an error
    - `quietly()` captures printed output, messages, and warnings instead of capturing errors

## Creating a safe function

```{r}

# Create safe_readLines() by passing readLines() to safely()
safe_readLines <- safely(readLines)

# Call safe_readLines() on "http://example.org"
safe_readLines("http://example.org")

# Call safe_readLines() on "http://asdfasdasdkfjlda"
safe_readLines("http://asdfasdasdkfjlda")

```

## using map safely

```{r}

urls <- structure(list(
  example = "http://example.org", 
  rproj = "http://www.r-project.org", 
  asdf = "http://asdfasdasdkfjlda"), 
  .Names = c("example", "rproj", "asdf"))

# Define safe_readLines()
safe_readLines <- safely(readLines)

# Use the safe_readLines() function with map(): html
html <- map(urls, safe_readLines)

# Call str() on html
str(html)

# Extract the result from one of the successful elements
html[[1]][['result']]

# Extract the error from the element that was unsuccessful
html[[3]][['error']]

```

## Working with safe output

```{r}

# Define save_readLines() and html
safe_readLines <- safely(readLines)
html <- map(urls, safe_readLines)

# Examine the structure of transpose(html)
str(transpose(html))

# Extract the results: res
res <- transpose(html)[['result']]
str(res)

# Extract the errors: errs
errs <- transpose(html)[['error']]
str(errs)
```

## Working with errors and results

```{r}

# Initialize some objects
safe_readLines <- safely(readLines)
html <- map(urls, safe_readLines)
res <- transpose(html)[["result"]]
errs <- transpose(html)[["error"]]

# Create a logical vector is_ok
is_ok <- map_lgl(errs, is_null)

# Extract the successful results
## ha, I just used map to shorten my results with head. : ) 
map(res[is_ok],head)

# Extract the input from the unsuccessful results
urls[!is_ok]

```

## Maps over multiple arguments

  - `map2()` - iterate over two arguments
  - `pmap()` - iterate over many arguments
  - `invoke_map()` - iterate over funtions and arguments
  - Like `map()`, each has a whole family of functions:
    - `map2_dpl`, `map2_lgl`, `pmap_dbl`, etc

## Getting started

```{r}

# Create a list n containing the values: 5, 10, and 20
n <- list(5,10,20)

# Call map() on n with rnorm() to simulate three samples
map(n, rnorm)

```

## Mapping over two arguments

```{r}

# Initialize n
n <- list(5, 10, 20)

# Create a list mu containing the values: 1, 5, and 10
mu <- list(1, 5, 10)

# Edit to call map2() on n and mu with rnorm() to simulate three samples
map2(n, mu, rnorm)

```

## Mapping over more than two arguments

```{r}

# Initialize n and mu
n <- list(5, 10, 20)
mu <- list(1, 5, 10)

# Create a sd list with the values: 0.1, 1 and 0.1
sd <- list(0.1, 1, 0.1)

# Edit this call to pmap() to iterate over the sd list as well
pmap(list(n, mu, sd), rnorm)

```

## Argument matching

```{r}

# Name the elements of the argument list
pmap(list(mean = mu, n = n, sd = sd), rnorm)

```

## mapping over functions and their arguments

```{r}

# Define list of functions
f <- list("rnorm", "runif", "rexp")

# Parameter list for rnorm()
rnorm_params <- list(mean = 10)

# Add a min element with value 0 and max element with value 5
runif_params <- list(min = 0, max = 5)

# Add a rate element with value 5
rexp_params <- list(rate = 5)

# Define params for each function
params <- list(
  rnorm_params,
  runif_params,
  rexp_params
)

# Call invoke_map() on f supplying params as the second argument
invoke_map(f, params, n = 5)

```

## Maps with side effects

  - Side effects
    - Describe things that happen beyond the results of a function
    - Examples include: printing output, plotting, and saving files to disk
  - `walk()` works just like `map()`, but is designed for functions called for their side effects
  
## Walk

```{r}

# Define list of functions
f <- list(Normal = "rnorm", Uniform = "runif", Exp = "rexp")

# Define params
params <- list(
  Normal = list(mean = 10),
  Uniform = list(min = 0, max = 5),
  Exp = list(rate = 5)
)

# Assign the simulated samples to sims
sims <- invoke_map(f, params, n = 50)

# Use walk() to make a histogram of each element in sims
walk(sims, hist)

```

## Walking over two or more arguments

```{r}

# Replace "Sturges" with reasonable breaks for each sample
breaks_list <- list(
  Normal = seq(6, 16, 0.5),
  Uniform = seq(0, 5, 0.25),
  Exp = seq(0, 1.5, 0.1)
)

# Use walk2() to make histograms with the right breaks
walk2(sims, breaks_list, hist)

```

## Putting together writing functions and walk

```{r}

# Turn this snippet into find_breaks()

find_breaks <- function(x) {
  rng <- range(x, na.rm = TRUE)
  seq(rng[1], rng[2], length.out = 30)
}


# Call find_breaks() on sims[[1]]
find_breaks(sims[[1]])

```

## Nice breaks for all

```{r}

# Use map() to iterate find_breaks() over sims: nice_breaks
nice_breaks <- map(sims, find_breaks)

# Use nice_breaks as the second argument to walk2()
walk2(sims, nice_breaks, hist)

```

## Walking with many argumnets: pwalk

```{r}

# Increase sample size to 1000
sims <- invoke_map(f, params, n = 1000)

# Compute nice_breaks (don't change this)
nice_breaks <- map(sims, find_breaks)

# Create a vector nice_titles
nice_titles <- c(
  "Normal(10, 1)",
  "Uniform(0, 5)",
  "Exp(5)"
  )

# Use pwalk() instead of walk2()
pwalk(list(x = sims, breaks = nice_breaks, main = nice_titles), hist, xlab = "")

```

## Walking with pipes

```{r}

# Pipe this along to map(), using summary() as .f
sims %>%
  walk(hist) %>%
  map(summary)

```


&nbsp; &nbsp;

***  
# Robust functions
***  

## Robust functions

  - The outputs to these change based on the input
    - `df[, vars]`
    - `subset(df, x == y)`
    - `data.frame(x = "a")`
  - The aim at interactive analysis which is helpful
    - But in programming you want strict outputs
  - Three main problems
    - Type-unstable functions
    - Non-standard evaluations
    - Hidden arguments
  - Throwing clear errors is important
    
## An error is better than a surprise

  - This will give an error rather then weird results like we saw before
  
```{r, error = T}

# Define troublesome x and y
x <- c(NA, NA, NA)
y <- c( 1, NA, NA, NA)

both_na <- function(x, y) {
  # Add stopifnot() to check length of x and y
  stopifnot(length(x) == length(y))
  
  sum(is.na(x) & is.na(y))
}

# Call both_na() on x and y
both_na(x, y)

```

## An informative error is even better

  - The error message should tell the uses what is needed
  
```{r, error = T}

# Define troublesome x and y
x <- c(NA, NA, NA)
y <- c( 1, NA, NA, NA)

both_na <- function(x, y) {
  # Replace condition with logical
  if (length(x) != length(y)) {
    # Replace "Error" with better message
    stop("x and y must have the same length", call. = FALSE)
  }  
  
  sum(is.na(x) & is.na(y))
}

# Call both_na() 
both_na(x, y)

```

## unstable types

  - Type-inconsistent: the type of the return object depends on the input
  - Surprises occur when you've used a type-inconsistent functio inside your own function
  - `[` is a common source of surprises
    - use `drop = FALSE`: `df[x, , drop = FALSE]`
    - Subset the data frame like a list: df[x]
  - sapply is another type unstable function
  - Avoid these when writting your functions and use type consistent functions instead or use tests to ensure type

## sapply is another common culprite

```{r}

df <- data.frame(
  a = 1L,
  b = 1.5,
  y = Sys.time(),
  z = ordered(1)
)

A <- sapply(df[1:4], class) 
A

B <- sapply(df[3:4], class)
B
```

## Using purrr solves the problem

```{r}

# sapply calls
A <- sapply(df[1:4], class) 
B <- sapply(df[3:4], class)
C <- sapply(df[1:2], class) 

# Demonstrate type inconsistency
str(A)
str(B)
str(C)

# Use map() to define X, Y and Z
X <- map(df[1:4], class) 
Y <- map(df[3:4], class)
Z <- map(df[1:2], class) 

# Use str() to check type consistency
str(X)
str(Y)
str(Z)

```

## A type consistent solution

```{r}

col_classes <- function(df) {
  # Assign list output to class_list
  class_list <- map(df, class)
  
  # Use map_chr() to extract first element in class_list
  map_chr(class_list, 1)
  
}

# Check that our new function is type consistent
df %>% col_classes() %>% str()
df[3:4] %>% col_classes() %>% str()
df[1:2] %>% col_classes() %>% str()

```

## Or fail early if something goes wrong

```{r, error = T}


## or make our own error message
col_classes <- function(df) {
  class_list <- map(df, class)
  
  # Add a check that no element of class_list has length > 1
  if (any(map_dbl(class_list, length) > 1)) {
    stop("Some columns have more than one class", call. = FALSE)
  }
  
  # Use flatten_chr() to return a character vector
  flatten_chr(class_list)
}

# Check that our new function is type consistent
df %>% col_classes() %>% str()
df[3:4] %>% col_classes() %>% str()
df[1:2] %>% col_classes() %>% str()


```

## Non-standard evaluation

  - NSE function don't use the normal lookup roles
  - Things like subsetting, dply filter, ggplot are examples
  - What to do
    - Using NSE functions inside your own functions can cause surprises
    - Avoind using NSE functions inside your functions
    - Or learn the suprising cases and protect against them
    - I'll probably choose the later, because using dplyr inside functions is something I do a lot
    - But I am usually working on a constrained dataset and problem, not writting packages for other people

## Programming with NSE functions

```{r}

big_x <- function(df, threshold) {
  dplyr::filter(df, x > threshold)
}

# Use big_x() to find rows in diamonds_sub where x > 7
big_x(diamonds_sub, 7)

```

## When things go wrong

```{r}

# Remove the x column from diamonds
diamonds_sub$x <- NULL

# Create variable x with value 1
x <- 1

# Use big_x() to find rows in diamonds_sub where x > 7
big_x(diamonds_sub, 7)

# Create a threshold column with value 100
diamonds_sub$threshold <- 100

# Use big_x() to find rows in diamonds_sub where x > 7
big_x(diamonds_sub, 7)

```

## What to do?

```{r, error = T}

big_x <- function(df, threshold) {
  # Write a check for x not being in df
  if (!'x' %in% names(df)) {
    stop("df must contain variable called x", call. = FALSE)
  }
  
  # Write a check for threshold being in df
  if ('threshold' %in% names(df)) {
    stop("df must not contain variable called threshold", call. = FALSE)
  }
  
  dplyr::filter(df, x > threshold)
}

big_x(diamonds_sub, 7)

```

## Hidden arguments

  - Pure functions
    - Their output only depends on their inputs
    - They don't affect the outside workd except through their return value
  - Hidden arguments are function inputs taht may be different for different users or sessions
    - Common example: argument defaults taht depend on global options
  - The return value of a function shouild **never** depend on a global option
    - Side effects may be controlled by global options

## A hidden dependence

```{r}

# Read in the swimming_pools.csv to pools
url_csv <- "http://s3.amazonaws.com/assets.datacamp.com/production/course_1478/datasets/swimming_pools.csv"
swimming_pools <- read.csv(url_csv)

# Examine the structure of pools
str(pools)

# Change the global stringsAsFactor option to FALSE
options(stringsAsFactors = F)
getOption("stringsAsFactors")

# Read in the swimming_pools.csv to pools2
pools2 <- read.csv(url_csv)

# Examine the structure of pools2
str(pools2)

```

## Legitimate use of options

```{r}

# Fit a regression model
fit <- lm(mpg ~ wt, data = mtcars)

# Look at the summary of the model
summary(fit)

# Set the global digits option to 2
options(digits = 2)

# Take another look at the summary
summary(fit)

```




