---
title: "Data Visualization with ggplot2 (Part 1) - Chapter 4 and 5"
author: "William Surles"
date: "2017-07-26"
output: 
 html_document:
  self_contained: yes
  theme: flatly
  highlight: tango
  toc: true
  toc_float: true
  toc_depth: 3
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(eval=T, echo=T, cache=T, message=F, warning=F)

source("create_datasets.R")

library(dplyr)
library(ggplot2)
```

# My Introduction

  - Course notes from the [Data Vizualization with ggplot2 (Part 1)](https://www.datacamp.com/courses/data-visualization-with-ggplot2-1) course on DataCamp
  
## Whats Covered

I had to split this up becasue of the size of the document.

  - Doc A 
    - Introduction 
    - Data
    - Aesthetics
  - Doc B 
    - Geometries
    - qplot and wrap-up
    
## Aditional Resources
 
&nbsp; &nbsp;

***  
# Geometries
***  

## Scatter Plots

  - 37 geometries. whaa!
    - `abline`, `area`, `bar`, `bin2d`, `blank`, `boxplot`
    - `contour`, `crossbar`, `density`, `density2d`, `dotplot`
    - `errorbar`, `errorbarh`, `freqpoly`, `hex`, `histogram`, `hline`
    - `jitter`, `line`, `linerange`, `map`, `path`, `point`, `pointrange`
    - `polygon`, `quantile`, `raster`, `rect`, `ribbon`, `rug`
    - `segment`, `smooth`, `step`, `text`, `tile`, `violin`, `vline`
  - 3 common plots
    - scatter plots -- `points`, `jitter`, `abline`
    - bar plots -- `histogram`, `bar`, `errorbar`
    - line plots -- `line`
  - You can also add a nother variable, such as summary statistics in the geom_point
  - Jitter and alpha are common to deal with overplotting

### -- Scatter plots and jittering (1)

```{r}

# Shown in the viewer:
ggplot(mtcars, aes(x = cyl, y = wt)) +
  geom_point()

# Solutions:
# 1 - With geom_jitter()
ggplot(mtcars, aes(x = cyl, y = wt)) +
  geom_jitter()

```

  - Thats a little too much jitter
  - We lose the sense of separate variables
  - We can adjust the jitter width to fix this
  
```{r}
# 2 - Set width in geom_jitter()
ggplot(mtcars, aes(x = cyl, y = wt)) +
  geom_jitter(width = 0.1)

# 3 - Set position = position_jitter() in geom_point() ()
ggplot(mtcars, aes(x = cyl, y = wt)) +
  geom_point(position = position_jitter(0.1))

```

### -- Scatter plots and jittering (2)

```{r}

# Examine the structure of Vocab
library(car)
str(Vocab)

# Basic scatter plot of vocabulary (y) against education (x). Use geom_point()
ggplot(Vocab, aes(education, vocabulary)) + 
  geom_point()


# Use geom_jitter() instead of geom_point()
ggplot(Vocab, aes(education, vocabulary)) + 
  geom_jitter()

  
# Using the above plotting command, set alpha to a very low 0.2
ggplot(Vocab, aes(education, vocabulary)) + 
  geom_jitter(alpha = 0.2)

  
# Using the above plotting command, set the shape to 1
ggplot(Vocab, aes(education, vocabulary)) + 
  geom_jitter(alpha = 0.2, shape = 1)

```

## Bar Plots

### -- Histograms

```{r}

# 1 - Make a univariate histogram
ggplot(mtcars, aes(x = mpg)) +
  geom_histogram()

```

  - By default the bin width is the range/30
  
```{r}
# 2 - Plot 1, plus set binwidth to 1 in the geom layer
ggplot(mtcars, aes(x = mpg)) +
  geom_histogram(binwidth = 1)

```

  - We also have access to internal data frame for the plot
  - We can change the y aes to `..density..` 
  - This is calculated internally in the summary stats just like `count`

```{r}
# 3 - Plot 2, plus MAP ..density.. to the y aesthetic (i.e. in a second aes() function)
ggplot(mtcars, aes(x = mpg)) +
  geom_histogram(aes(y = ..density..), binwidth = 1)

# 4 - plot 3, plus SET the fill attribute to "#377EB8"
ggplot(mtcars, aes(x = mpg)) +
  geom_histogram(aes(y = ..density..), binwidth = 1, fill = "#377EB8")

```

### -- Position

```{r}

# Draw a bar plot of cyl, filled according to am
ggplot(mtcars, aes(x = cyl, fill = am)) +
  geom_bar()

```

  - `stack` is the default
  
```{r}
# Change the position argument to stack
ggplot(mtcars, aes(x = cyl, fill = am)) +
  geom_bar(position = "stack")

# Change the position argument to fill
ggplot(mtcars, aes(x = cyl, fill = am)) +
  geom_bar(position = "fill")

# Change the position argument to dodge
ggplot(mtcars, aes(x = cyl, fill = am)) +
  geom_bar(position = "dodge")

```

### -- Overlapping bar plots

```{r}

# 1 - The last plot form the previous exercise
# ggplot(mtcars, aes(x = cyl, fill = am)) + 
#   geom_bar(position = "dodge")

# 2 - Define posn_d with position_dodge()
posn_d <- position_dodge(width = 0.2)

# 3 - Change the position argument to posn_d
ggplot(mtcars, aes(x = cyl, fill = am)) + 
  geom_bar(position = posn_d)


# 4 - Use posn_d as position and adjust alpha to 0.6
ggplot(mtcars, aes(x = cyl, fill = am)) + 
  geom_bar(position = posn_d, alpha = 0.6)

```

### -- Overlapping histograms

```{r}

# A basic histogram, add coloring defined by cyl 
ggplot(mtcars, aes(mpg, fill = cyl)) +
  geom_histogram(binwidth = 1)

# Change position to identity 
ggplot(mtcars, aes(mpg, fill = cyl)) +
  geom_histogram(binwidth = 1, position = 'identity')

```

  - Now the bars are not stacked, but some are hidden behind others
  
```{r}
# Change geom to freqpoly (position is identity by default) 
ggplot(mtcars, aes(mpg, col = cyl)) +
  geom_freqpoly(binwidth = 1)

```

  - This would look much better with a more full histogram
  
### -- Bar plots with color ramp, part 1

  - Color brewer is a great package for working with colors
  - Its worth its own chapter probably. 
```{r}

# Example of how to use a brewed color palette
ggplot(mtcars, aes(x = cyl, fill = am)) +
  geom_bar() +
  scale_fill_brewer(palette = "Set1")

# Use str() on Vocab to check out the structure
Vocab$education <- as.factor(Vocab$education)
Vocab$vocabulary <- as.factor(Vocab$vocabulary)
str(Vocab)

# Plot education on x and vocabulary on fill
# Use the default brewed color palette
ggplot(Vocab, aes(x = education, fill = vocabulary)) +
  geom_bar(position = 'fill') + 
  scale_fill_brewer()

```

  - There are only 9 colors in the "Blues" palette, but we have 11 categories under vocabulary
    - So we get an error. And a weird looking chart. groups 9 and 10 are blank
    - We need to create our own palette with the same number of colors as groups
    
### -- Bar plots with color ramp, part 2

  - A quick example of how the colorRampPalette works
    - It returns a function that can be used to make a new palette
    - You can just use two colors and it will scale between them
    - Or you can use many colors or an existing palette 
    
```{r}

new_col <- colorRampPalette(c("#FFFFFF", "#0000FF"))
new_col(4) # the newly extrapolated colours
munsell::plot_hex(new_col(4)) # Quick and dirty plot

```

```{r}
library(RColorBrewer)

# Final plot of last exercise
ggplot(Vocab, aes(x = education, fill = vocabulary)) +
  geom_bar(position = "fill") +
  scale_fill_brewer()
  
# Definition of a set of blue colors
blues <- brewer.pal(9, "Blues") # from the RColorBrewer package
blues

# 1 - Make a color range using colorRampPalette() and the set of blues
blue_range <- colorRampPalette(blues)

# This is our new pallete. We can create it with as many colors as we want. 
munsell::plot_hex(blue_range(11)) 

# 2 - Use blue_range to adjust the color of the bars, use scale_fill_manual()
ggplot(Vocab, aes(x = education, fill = vocabulary)) +
  geom_bar(position = "fill") +
  scale_fill_manual(values = blue_range(11))

```
  - Nice. Thats much better. 
  
### -- Overlapping histograms (2)

```{r}

# 1 - Basic histogram plot command
ggplot(mtcars, aes(mpg)) + 
  geom_histogram(binwidth = 1)

# 2 - Plot 1, Expand aesthetics: am onto fill
ggplot(mtcars, aes(mpg, fill = am)) + 
  geom_histogram(binwidth = 1)

# 3 - Plot 2, change position = "dodge"
ggplot(mtcars, aes(mpg, fill = am)) + 
  geom_histogram(binwidth = 1, position = "dodge")

# 4 - Plot 3, change position = "fill"
## In this case, none of these positions really work well, because it's difficult to compare the distributions directly.
ggplot(mtcars, aes(mpg, fill = am)) + 
  geom_histogram(binwidth = 1, position = "fill")

# 5 - Plot 4, plus change position = "identity" and alpha = 0.4
ggplot(mtcars, aes(mpg, fill = am)) + 
  geom_histogram(binwidth = 1, 
    position = "identity",
    alpha = 0.4)

# 6 - Plot 5, plus change mapping: cyl onto fill
ggplot(mtcars, aes(mpg, fill = cyl)) + 
  geom_histogram(binwidth = 1, 
    position = "identity",
    alpha = 0.4)

```

## Line Plots - Time Series

### -- Line plots

```{r}

# Print out head of economics
head(economics)

# Plot unemploy as a function of date using a line plot
ggplot(economics, aes(x = date, y = unemploy)) +
  geom_line()
  
  
# Adjust plot to represent the fraction of total population that is unemployed
ggplot(economics, aes(x = date, y = unemploy/pop)) +
  geom_line()

```

### -- Periods of recession

```{r}
# Basic line plot
# ggplot(economics, aes(x = date, y = unemploy/pop)) +
#   geom_line()

# Expand the following command with geom_rect() to draw the recess periods
ggplot(economics, aes(x = date, y = unemploy/pop)) +
  geom_rect(data = recess,
         aes(xmin = begin, 
             xmax = end, 
             ymin = -Inf, 
             ymax = Inf),
            inherit.aes = FALSE,
            fill = "red", 
            alpha = 0.2
            ) + 
  geom_line()

```

### -- Multiple time series, part 1

```{r}

# Check the structure as a starting point
str(fish.species)

# Use gather to go from fish.species to fish.tidy
fish.tidy <- gather(fish.species, Species, Capture, -Year)

str(fish.tidy)

```

### -- Multiple time series, part 2

```{r}

# Recreate the plot shown on the right
ggplot(fish.tidy, aes(x = Year, y = Capture, col = Species)) +
  geom_line()
  
```


&nbsp; &nbsp;

***  
# qplot and wrap-up
***  

## qplot

  - qplot is sort of the base R eqivalent in terms of syntax
    - you can make charts quick and easy. It will guess the geom for you
  - But its better just to know the ggplot layers and be explicit and build your plots.
    - This is intuative and not that much more typing. 
    - Its still good to know about qplot because you will see it 

### -- Using qplot

```{r}

# The old way (shown)
plot(mpg ~ wt, data = mtcars) # formula notation
with(mtcars, plot(wt, mpg)) # x, y notation

# Using ggplot:
ggplot(mtcars, aes(wt, mpg)) +
  geom_point()

# Using qplot:
qplot(wt, mpg, data = mtcars)

```

### -- Using aesthetics

```{r}

# Categorical variable mapped onto size:
# cyl
qplot(wt, mpg, data = mtcars, size = factor(cyl))

# gear
qplot(wt, mpg, data = mtcars, size = factor(gear))

# Continuous variable mapped onto col:
# hp
qplot(wt, mpg, data = mtcars, col = hp)

# qsec
qplot(wt, mpg, data = mtcars, col = qsec)

```

### -- Choosing geoms, part 1

```{r}

# qplot() with x only
qplot(x = factor(cyl), data = mtcars)

# qplot() with x and y
qplot(x = factor(cyl), y = factor(vs), data = mtcars)

# qplot() with geom set to jitter manually
qplot(x = factor(cyl), y = factor(vs), data = mtcars, geom = 'jitter')

```

### -- Choosing geoms, part 2 - dotplot

```{r}

# cyl and am are factors, wt is numeric
class(mtcars$cyl)
class(mtcars$am)
class(mtcars$wt)

# "Basic" dot plot, with geom_point():
ggplot(mtcars, aes(cyl, wt, col = am)) +
  geom_point(position = position_jitter(0.2, 0))

# 1 - "True" dot plot, with geom_dotplot():
ggplot(mtcars, aes(cyl, wt, fill = am)) +
  geom_dotplot(binaxis = "y", stackdir = "center")

# 2 - qplot with geom "dotplot", binaxis = "y" and stackdir = "center"
qplot(
  cyl, wt, 
  data = mtcars, 
  fill = am, 
  geom = "dotplot", 
  binaxis = "y", 
  stackdir = "center"
)

```

## Wrap-up

### -- Chicken weight

```{r}

# ChickWeight is available in your workspace
# 1 - Check out the head of ChickWeight
head(ChickWeight)

# 2 - Basic line plot
ggplot(ChickWeight, aes(x = Time, y = weight)) + 
  geom_line(aes(group = Chick))

# 3 - Take plot 2, map Diet onto col.
ggplot(ChickWeight, 
    aes(x = Time, y = weight, col = Diet)) + 
  geom_line(
    aes(group = Chick))


# 4 - Take plot 3, add geom_smooth()
ggplot(ChickWeight, 
    aes(x = Time, y = weight, col = Diet)) + 
  geom_line(
    aes(group = Chick), alpha = 0.3) + 
  geom_smooth(lwd = 2, se = F)

```

### -- Titanic

```{r}

# titanic is avaliable in your workspace
# 1 - Check the structure of titanic
str(titanic)

# 2 - Use ggplot() for the first instruction
ggplot(titanic, 
    aes(x = Pclass, fill = Sex)) + 
  geom_bar(
    position = "dodge")

# 3 - Plot 2, add facet_grid() layer
ggplot(titanic, 
    aes(x = Pclass, fill = Sex)) + 
  geom_bar(
    position = "dodge") +
  facet_grid(. ~ Survived)

# 4 - Define an object for position jitterdodge, to use below
posn.jd <- position_jitterdodge(0.5, 0, 0.6)

# 5 - Plot 3, but use the position object from instruction 4
ggplot(titanic, 
    aes(x = Pclass, y = Age, col = Sex)) + 
  geom_point(
    size = 3, alpha = 0.5, position = posn.jd) +
  facet_grid(. ~ Survived)

```


