---
title: "Intermediate R"
author: "William Surles"
date: "2017-03-29"
output: 
 html_document:
  self_contained: yes
  theme: flatly
  highlight: tango
---

***
> This document was updated on **`r Sys.Date()`**.

***

```{r setup, include=FALSE}
library(shiny)
library(codetools)
knitr::opts_chunk$set(eval=T, echo=T, cache=T, message=F, warning=F)
```

# Introduction

  - This all comes from the [Intermediate R](https://www.datacamp.com/courses/intermediate-r) course on DataCamp
    - I highly recommend the [DataCamp](https://www.datacamp.com/home) courses
    - Go *do* the [course](https://www.datacamp.com/courses/intermediate-r) for yourself to get more details and videos on these topics
  - I am copying my work from the class to this document and adding comments

## Whats Covered

  - Conditionals and Control Flow
    - Relational Operators. `==` `<` `>` 
    - Logical Operators. `and` `or` `not`
    - Conditional Statements. `if` `else`
  - Loops
    - While loops
    - For loops
    - Breaks
  - Functions
    - Documentation
    - Required vs optional inputs
    - Scoping
    - Writing functions
    - Loading packages
  - Apply statements
    - lapply, sapply, vapply
  - Utilities
    - Useful base functions
    - Regular Expressions and grep/sub
    - Times and Dates
    
  
  
&nbsp; &nbsp;

***  

# Conditionals and Control Flow

## Relational Operators
  - These are used to compare objects
  - They are common in any langage and are a basis of programming

### Equality

```{r}

# Comparison of logicals
TRUE == FALSE

# Comparison of numerics
-6 * 14 != 17 - 101

# Comparison of character strings
"useR" == "user"

# Compare a logical with a numeric
TRUE == 1

```

### Greater and less than

```{r}

# Comparison of numerics
-6 * 5 + 2 >= -10 + 1

# Comparison of character strings
"raining" <= "raining dogs"

# Comparison of logicals
TRUE > FALSE

```

### Compare vectors

```{r}

# The linkedin and facebook vectors have already been created for you
linkedin <- c(16, 9, 13, 5, 2, 17, 14)
facebook <- c(17, 7, 5, 16, 8, 13, 14)

# Popular days
linkedin > 15

# Quiet days
linkedin <= 5

# LinkedIn more popular than Facebook
linkedin > facebook

```

### Compare matrices

```{r}

linkedin <- c(16, 9, 13, 5, 2, 17, 14)
facebook <- c(17, 7, 5, 16, 8, 13, 14)

views <- matrix(c(linkedin, facebook), nrow = 2, byrow = TRUE)

# When does views equal 13?
views == 13

# When is views less than or equal to 14?
views <= 14

# How often does facebook equal or exceed linkedin times two?
sum(facebook >= linkedin * 2)

```

## Logical Operators

  - These are also super common and a basis of programming
  - They let us combine multiple relational statements

### & and |

```{r}

linkedin <- c(16, 9, 13, 5, 2, 17, 14)

last <- tail(linkedin, 1)

# Is last under 5 or above 10?
last < 5 | last > 10

# Is last between 15 (exclusive) and 20 (inclusive)?
last > 15 & last <= 20

# Is last between 0 and 5 or between 10 and 15?
(last > 0 & last < 5) | (last > 10 & last < 15)

```

### & and | (2)

```{r}

linkedin <- c(16, 9, 13, 5, 2, 17, 14)
facebook <- c(17, 7, 5, 16, 8, 13, 14)

# linkedin exceeds 10 but facebook below 10
linkedin > 10 & facebook < 10

# When were one or both visited at least 12 times?
linkedin > 12 | facebook > 12

# When is views between 11 (exclusive) and 14 (inclusive)?
views > 11 & views <= 14

```

### Reverse the result: !

```{r}

!TRUE
!(5 > 3)
!!FALSE

x <- 5
y <- 7
!(!(x < 4) & !!!(y > 12))

```

### Blend it all together

```{r}

second <- c(3, 23, 18, 18, 25, 20, 17, 27, 6, 35, 17, 6, 1, 12, 15, 17, 12, 8)

# Build a logical vector, TRUE if value in second is extreme: extremes
extremes <- second > 25 | second < 5

# Count the number of TRUEs in extremes
sum(extremes)

```

## Conditional Statements

  - Used to execute statements based on result of relational statements
  - All of this is basis of programming and can be used in all the languages
  
### The if statement

```{r}

# Variables related to your last day of recordings
medium <- "LinkedIn"
num_views <- 14

# Examine the if statement for medium
if (medium == "LinkedIn") {
  print("Showing LinkedIn information")
}

# Write the if statement for num_views
if (num_views > 15) {
  print("You're popular!")
}

```

### And an else

```{r}

# Variables related to your last day of recordings
medium <- "LinkedIn"
num_views <- 14

# Control structure for medium
if (medium == "LinkedIn") {
  print("Showing LinkedIn information")
} else {
  print("Unknown medium")
}

# Control structure for num_views
if (num_views > 15) {
  print("You're popular!")
} else {
  print("Try to be more visible!")
}

```

### Customize further: else if

```{r}

# Variables related to your last day of recordings
medium <- "LinkedIn"
num_views <- 14

# Control structure for medium
if (medium == "LinkedIn") {
  print("Showing LinkedIn information")
} else if (medium == "Facebook") {
  # Add code to print correct string when condition is True
  print("Showing Facebook information")
} else {
  print("Unknown medium")
}

# Control structure for num_views
if (num_views > 15) {
  print("You're popular!")
} else if (num_views <= 15 & num_views > 10) {
  # Add code to print correct string when condition is TRUE
  print("Your number of views is average")
} else {
  print("Try to be more visible!")
}

```

### Else if 2.0

```{r}

numbers <- c(6, 100, 4, 2500)

for (number in numbers) {
  
  if (number < 10) {
    if (number < 5) {
      result <- "extra small"
    } else {
      result <- "small"
    }
  } else if (number < 100) {
    result <- "medium"
  } else {
    result <- "large"
  }
  
  print(result)
}

```

### Take control!

```{r}

# Variables related to your last day of recordings
li <- 15
fb <- 9

# Code the control-flow construct
if (li >= 15 & fb >= 15) {
  sms <- (li + fb) * 2
} else if (li < 10 & fb < 10) {
  sms <- (li + fb) / 2
} else {
  sms <- (li + fb)
}

# Print the resulting sms to the console
print(sms)

```

&nbsp; &nbsp;

***  

# Loops

## While loop

  - it runs as long as the while condition is true

### Write a while loop

```{r}

# Initialize the speed variable
speed <- 64

# Code the while loop
while (speed > 30 ) {
  print('Slow down!')
  speed <- speed - 7
}
print(speed)

```

### Throw in more conditionals

```{r}

# Initialize the speed variable
speed <- 64

# Extend/adapt the while loop
while (speed > 30) {
  print(paste("Your speed is",speed))
  if (speed > 48) {
    print("Slow down big time!")
    speed <- speed - 11
  } else {
    print("Slow down!")
    speed <- speed - 6
  }
}

```

### Stop the while loop: break

```{r}

# Initialize the speed variable
speed <- 88

while (speed > 30) {
  print(paste("Your speed is",speed))
  
  # Break the while loop when speed exceeds 80
  if (speed > 80) {
    break
  }
  
  if (speed > 48) {
    print("Slow down big time!")
    speed <- speed - 11
  } else {
    print("Slow down!")
    speed <- speed - 6
  }
}

```

### Build a while loop from scratch

```{r}

# Initialize i as 1 
i <- 1

# Code the while loop
while (i <= 10) {
  triple <- i * 3
  print(triple)
  
  if (triple %% 8 == 0) {
    break
  }
  i <- i + 1
}

```

## For loop

  - Run once for each variable in the list
    - This list can just be a sequence like 1:5 or variables in a list like a bunch of names
    - This is so fundamental to programming in any language
  - However I actualy never use loops in R
    - I always modify data arrays with array functions. They are WAY faster
    - The apply functions covered later are one way to do it. 
    - dplyr is the best way. All the loopig happens in C code so its super fast. 
    - If you are looping on an array in R, you should probably rethink your approach. 

### Loop over a vector

```{r}

# The linkedin vector has already been defined for you
linkedin <- c(16, 9, 13, 5, 2, 17, 14)

# Loop version 1
for (views in linkedin) {
  print(views)
}

# Loop version 2
for (i in 1:length(linkedin)) {
  print(linkedin[i])
}

```

### Loop over a list

```{r}

# The nyc list is already specified
nyc <- list(pop = 8405837, 
            boroughs = c("Manhattan", "Bronx", "Brooklyn", "Queens", "Staten Island"), 
            capital = FALSE)

# Loop version 1
for (value in nyc) {
  print(value)
}

# Loop version 2
for (i in 1:length(nyc)) {
  print(nyc[[i]])
}

```

### Loop over a matrix

```{r}

# The tic-tac-toe matrix has already been defined for you
ttt <- matrix(c("O", NA, "X", NA, "O", NA, "X", "O", "X"), nrow = 3, ncol = 3)

# define the double for loop
for (i in 1:nrow(ttt)) {
  for (j in 1:ncol(ttt)) {
  print(paste("On row",i,"and column",j,"the board contains",ttt[i,j]))
  }
}

```

### Mix it up with control flow

```{r}

# The linkedin vector has already been defined for you
linkedin <- c(16, 9, 13, 5, 2, 17, 14)

# Code the for loop with conditionals
for (i in 1:length(linkedin)) {
  if (linkedin[i] > 10) {
    print("You're popular!")
  } else {
    print("Be more visible!")
  }
  print(linkedin[i])
}

```

### Next, you break it

```{r}

# The linkedin vector has already been defined for you
linkedin <- c(16, 9, 13, 5, 2, 17, 14)

# Extend the for loop
for (li in linkedin) {
  if (li > 10) {
    print("You're popular!")
  } else {
    print("Be more visible!")
  }
  
  # Add code to conditionally break iteration
  if (li > 16) {
    print("This is ridiculous, I'm outta here!")
    break
  }
  
  # Add code to conditionally skip iteration
  if (li < 5) {
    print("This is too embarrassing!")
    next
  }
  
  print(li)
}

```

### Build a loop from scratch

```{r}

# Pre-defined variables
rquote <- "R's internals are irrefutably intriguing"
chars <- strsplit(rquote, split = "")[[1]]
chars

# Your solution here
rcount <- 0
for (c in chars) {
  if (c == 'u') {
    break
  }
  if (c == 'r' | c == 'R') {
    rcount <- rcount + 1
  }
}

# Print the resulting rcount variable to the console
print(rcount)

```

&nbsp; &nbsp;

***  

# Functions

### Introduction to Functions

  - You have already been using these a bunch. 

### Function documentation

```{r}

# Consult the documentation on the mean() function
# In Rstudio this will open the help window
?mean

# Inspect the arguments of the mean() function
args(mean)

```

### Use a function

```{r}

# The linkedin and facebook vectors have already been created for you
linkedin <- c(16, 9, 13, 5, 2, 17, 14)
facebook <- c(17, 7, 5, 16, 8, 13, 14)

# Calculate average number of views
avg_li <- mean(linkedin)
avg_fb <- mean(facebook)

# Inspect avg_li and avg_fb
avg_li
avg_fb


# Calculate the mean of linkedin minus facebook
mean(linkedin - facebook)

```

### Use a function (2)

```{r}

# The linkedin and facebook vectors have already been created for you
linkedin <- c(16, 9, 13, 5, 2, 17, 14)
facebook <- c(17, 7, 5, 16, 8, 13, 14)

# Calculate the mean of the sum
avg_sum <- mean(linkedin + facebook)

# Calculate the trimmed mean of the sum
avg_sum_trimmed <- mean(linkedin + facebook, trim = 0.2)

# Inspect both new variables
avg_sum
avg_sum_trimmed

```

### Use a function (3)

```{r}

# The linkedin and facebook vectors have already been created for you
linkedin <- c(16, 9, 13, 5, NA, 17, 14)
facebook <- c(17, NA, 5, 16, 8, 13, 14)

# Basic average of linkedin
mean(linkedin)
mean(facebook)

# Advanced average of facebook
mean(linkedin, na.rm = T)
mean(facebook, na.rm = T)

```

### Functions inside functions

```{r}

# The linkedin and facebook vectors have already been created for you
linkedin <- c(16, 9, 13, 5, NA, 17, 14)
facebook <- c(17, NA, 5, 16, 8, 13, 14)

# Calculate the mean absolute deviation
mean(abs(linkedin - facebook), na.rm = T)

```

### Required, or optional?

  - There are required and optional arguments to a function
  - Some (or actually most) of the optional arguments have a default value 
    - this will be set automatically
    - for example, `read.table` has `header = FALSE` by default
    
The begining of the `read.table` function definition looks like this:

`read.table(file, header = FALSE, sep = "", quote = "\"",...`

In the `read.table()` function: 

  - `file` is required
  - `header`, `sep`, and `quote` are optional arguments
    - `header` is defaulted to `FALSE`
    - `sep` is defaulted to and empty string `""`
    - `quote` is defaulted to `"`
  - You do not have to write all the argument names 
    - `read.table("myfile.txt", TRUE, "-")` will work.
  - But the order matters if you use that shortcut
    - `read.table("myfile.txt", "-", TRUE)` will throw an error.
  - You can use any order if you specify all the names and values
    - `read.table("myfile.txt", sep = "-", header = TRUE)` will work.

## Writing Functions

### Write your own functions

```{r}

# Create a function pow_two()
pow_two <- function(x) {
 result <- x^2
 return(result)
}

# Use the function 
pow_two(12)

# Create a function sum_abs()
sum_abs <- function(x, y) {
  result <- abs(x) + abs(y)
  return(result)
}

# Use the function
sum_abs(-2, 3)

```

### Write your own functions (2)

```{r}

# Define the function hello()
hello <- function() {
  print('Hi there!')
  return(TRUE)
}

# Call the function hello()
hello()

# Define the function my_filter()
my_filter <- function(x) {
  if (x >=0) {
    return(x)
  } else {
    return(NULL)
  }
}

# Call the function my_filter() twice
my_filter(5)
my_filter(-5)

```

### Write you own functions (3)

```{r}

# Extend the pow_two() function
pow_two <- function(x, print_info = T) {
  
  y <- x ^ 2
  
  if (print_info == T) {
    print(paste(x, 'to the power two equals',y))
  }
  
  return(y)
}

pow_two(5)
pow_two(5, print_info=F)

```

### Function scoping

  - Variables defined inside a function are not available outside of that function
    - Nor are the variable names given to the input
    - Calling x or y outside of this function would fail
  
```{r eval=F}

pow_two <- function(x) {
  y <- x ^ 2
  return(y)
}
pow_two(4)

```

### R passes arguments by value

  - In other words, a function won't change the original variable passed in

```{r}

triple <- function(x) {
  x <- 3*x
  x
}

a <- 5
triple(a)

a

```

### R you functional?

```{r}

# The linkedin and facebook vectors have already been created for you
linkedin <- c(16, 9, 13, 5, 2, 17, 14)
facebook <- c(17, 7, 5, 16, 8, 13, 14)

# Define the interpret function
interpret <- function(x) {

 if (x > 15) {
   print("You're popular!")
   return(x)
 } else {
   print("Try to be more visible!")
   return(0)
 }
 
}

# Call the interpret function twice
interpret(linkedin[1])
interpret(facebook[2])

```

### R you functional? (2)

```{r}

# The linkedin and facebook vectors have already been created for you
linkedin <- c(16, 9, 13, 5, 2, 17, 14)
facebook <- c(17, 7, 5, 16, 8, 13, 14)

# The interpret() can be used inside interpret_all()
interpret <- function(num_views) {
  if (num_views > 15) {
    print("You're popular!")
    return(num_views)
  } else {
    print("Try to be more visible!")
    return(0)
  }
}

# Define the interpret_all() function
interpret_all <- function(x, show_sum = T) {
  
  sum_pop_views <- 0
  
  for (i in x) {
    y <- interpret(i)
    sum_pop_views <- sum_pop_views + y
  }
  
  if (show_sum == T) {
    return(sum_pop_views)
  } else {
    return(NULL)
  }
  
}

# Call the interpret_all() function on both linkedin and facebook
interpret_all(facebook)
interpret_all(linkedin)

```

## R Packages

  - Many great functions are available through packages
  - The base package is installed with R and loaded when you start R
  - The others you need to load yourself when you want them
  - I think the reason R is so useful today is becasue it has amazing packages 
    - Especially the ones made by Hadley Wicklam or the folks at Rstudio

### Load an R Package

```{r}

# The mtcars vectors have already been prepared for you
wt <- mtcars$wt
hp <- mtcars$hp

# Request the currently attached packages
search()

# Load the ggplot2 package
library(ggplot2)

# Use the qplot() function. 
# It will fail if you try before loading ggplot2 library
qplot(wt,hp)

# Check out the currently attached packages again
search()

```

### Different ways to load a package

  - Its not picky. These all work.

```{r}

## I already have ggplot2 installed
## install.packages("ggplot2")

library(ggplot2)
library("ggplot2")
require(ggplot2)


```

&nbsp; &nbsp;

***  

# The apply family

### lapply

  - applies a function to each element of a list/vector
  - less code than writting out a whole loop
  - also faster
  - always returns a list
    - wrap in `unlist()` if you want a vector

### Use Lapply with a built-in R function

```{r}

# The vector pioneers has already been created for you
pioneers <- c("GAUSS:1777", "BAYES:1702", "PASCAL:1623", "PEARSON:1857")

# Split names from birth year: split_math
split_math <- strsplit(pioneers, ":")
split_math

# Convert to lowercase strings: split_low
split_low <- lapply(split_math, tolower)
split_low

# Take a look at the structure of split_low
str(split_low)

# For comparision, if you wrote out the whole loop
split_low2 <- list()

for (i in 1:length(split_math)) {
  split_low2[[i]] <- tolower(split_math[[i]])
}

split_low2

```

### Use lapply with your own function

```{r}

# Code from previous exercise:
pioneers <- c("GAUSS:1777", "BAYES:1702", "PASCAL:1623", "PEARSON:1857")
split <- strsplit(pioneers, split = ":")
split_low <- lapply(split, tolower)

# Write function select_first()
select_first <- function(x) {
  return(x[1])
}

# Apply select_first() over split_low: names
names <- lapply(split_low, select_first)
names

# Write function select_second()
# Write function select_first()
select_second <- function(x) {
  return(x[2])
}

# Apply select_second() over split_low: years
years <- lapply(split_low, select_second)
years

```

### lapply and anonymous functions

```{r}

# Definition of split_low
pioneers <- c("GAUSS:1777", "BAYES:1702", "PASCAL:1623", "PEARSON:1857")
split <- strsplit(pioneers, split = ":")
split_low <- lapply(split, tolower)

names <- lapply(split_low, function(x) {x[1]})
years <- lapply(split_low, function(x) {x[2]})

names
years

```

### Use lapply with additional arguments

```{r}

# Definition of split_low
pioneers <- c("GAUSS:1777", "BAYES:1702", "PASCAL:1623", "PEARSON:1857")
split <- strsplit(pioneers, split = ":")
split_low <- lapply(split, tolower)

# Replace the select_*() functions by a single function: select_el
select_el <- function(x,pos) { 
  x[pos] 
}

# Call the select_el() function twice on split_low: names and years
names <- lapply(split_low, select_el, 1)
years <- lapply(split_low, select_el, 2)

names
years

```

### Apply functions that return NULL

```{r}

lapply(list(1, "a", TRUE), str)
str(list(1,"a",TRUE))
str(TRUE)

```

## sapply

  - Stands for Simple apply
  - Like lapply, but returns a vetor if it can
  - It will name the vector by default
  - If there are multiple outputs it will return a matrix
  - If it can't return a vector or matrix it will return a list
  - Becareful with this if you expect a certain data type returned in a program

### How to use sapply

```{r}
# temp variable
temp <- list(
  c(3, 7,  9,  6, -1),
  c(6,  9, 12, 13,  5),
  c(4,  8,  3, -1, -3),
  c(1,  4,  7,  2, -2),
  c(5, 7, 9, 4, 2),
  c(-3,  5,  8,  9,  4),
  c(3, 6, 9, 4, 1)
  )

# Use lapply() to find each day's minimum temperature
lapply(temp, min)

# Use sapply() to find each day's minimum temperature
sapply(temp, min)

# Use lapply() to find each day's maximum temperature
lapply(temp, max)

# Use sapply() to find each day's maximum temperature
sapply(temp, max)

```

### sapply with your own function

```{r}

# temp is already defined in the workspace

# Define a function that calculates the average of the min and max of a vector: extremes_avg
extremes_avg <- function(x) {
 avg <- mean(c(min(x), max(x)))
 return(avg)
}

# Apply extremes_avg() over temp using sapply()
sapply(temp, extremes_avg)

# Apply extremes_avg() over temp using lapply()
lapply(temp, extremes_avg)

```

### sapply with function returning vector

```{r}

# temp is already available in the workspace

# Create a function that returns min and max of a vector: extremes

extremes <- function (x) {
  return(c(min(x),max(x)))
}

# Apply extremes() over temp with sapply()
sapply(temp, extremes)

# Apply extremes() over temp with lapply()
lapply(temp, extremes)

```

### sapply can't simplify, now what?

```{r}

# temp is already prepared for you in the workspace

# Create a function that returns all values below zero: below_zero
below_zero <- function (x) {
 val <- x[x <0]
 return(val)
}

below_zero(c(1,2,3,-1,-1,-3))

# Apply below_zero over temp using sapply(): freezing_s
freezing_s <- sapply(temp, below_zero)
freezing_s

# Apply below_zero over temp using lapply(): freezing_l
freezing_l <- lapply(temp, below_zero)
freezing_l

# Compare freezing_s to freezing_l using identical()
identical(freezing_s, freezing_l)

```

### sapply with functions that return NULL

```{r}

# temp is already available in the workspace

# Write a function that 'cat()s' out the average temperatures: print_info
print_info <- function (x) {
  cat("The average temperature is", mean(x), "\n")
}


# Apply print_info() over temp using lapply()
lapply(temp, print_info)

# Apply print_info() over temp using sapply()
sapply(temp, print_info)

```

### Reverse engineering sapply

  - this uses an anonymous function
  - the result will have 3 rows and 2 columns

```{r}
sapply(list(runif (10), runif (10)), 
       function(x) c(min = min(x), mean = mean(x), max = max(x)))
```

## vapply

  - this is safer than sapply 
    - becasue sapply can return a vector or a list (if result lengths differ)
  - you can tell it what the return data type should be

### Use vapply

```{r}

# temp is already available in the workspace

# Code the basics() function
basics <- function (x) {
  return(c(min(x), mean(x), max(x)))
}

# Apply basics() over temp using vapply()
vapply(temp, basics, numeric(3))

```

### Use vapply (2)

```{r}

# temp is already available in the workspace

# Definition of the basics() function
basics <- function(x) {
  c(min = min(x), mean = mean(x), median = median(x), max = max(x))
}

# This time there are 4 values
vapply(temp, basics, numeric(4))

```

### From sapply to vapply

```{r}

# temp is already defined in the workspace
temp

# sapply() expression
sapply(temp, max)

# Convert to vapply() expression
vapply(temp, max, numeric(1))

# sapply() expression
sapply(temp, function(x, y) { mean(x) > y }, y = 5)

# Convert to vapply() expression
vapply(temp, function(x, y) { mean(x) > y }, y = 5, logical(1))

# Definition of get_info (don't change)
get_info <- function(x, y) { 
  if (mean(x) > y) {
    return("Not too cold!")
  } else {
    return("Pretty cold!")
  }
}

# sapply() expression
sapply(temp, get_info, y = 5)

# Convert to vapply() expression
vapply(temp, get_info, y = 5, character(1))

```

&nbsp; &nbsp;

***  

# Utilities

### Useful Functions

The class video goes through some useful functions. 
I'll list them here. There are many more than this. 

  - `abs()`
    - calc the absolute value of a vector
  - `sum()`
    - calc the sum of a vector
  - `mean()`
    - calc the mean value of a vector
  - `seq()`
    - create a sequence
  - `rep()`
    - repeat an array
  - `sort()`
    - sort a vector
  - `str()`
    - see the structure of your datatype
  - `is.*()`, `as.*()`
    - check your data type of change it
  - `append()`, `rev()`
    - append and reverse vectors

### Mathematical utilities

```{r}

# The errors vector has already been defined for you
errors <- c(1.9,-2.6,4.0,-9.5,-3.4,7.3)

# Sum of absolute rounded values of errors
sum(round(abs(errors)))

```

### Find the error

```{r}

# Don't edit these two lines
vec1 <- c(1.5,2.5,8.4,3.7,6.3)
vec2 <- rev(vec1)

# Fix the error
mean(c(abs(vec1), abs(vec2)))

```

### Data Utilities

```{r}

# The linkedin and facebook vectors have already been created for you
linkedin <- list(16, 9, 13, 5, 2, 17, 14)
facebook <- list(17, 7, 5, 16, 8, 13, 14)

# Convert linkedin and facebook to a vector: li_vec and fb_vec
li_vec <- as.numeric(linkedin)
fb_vec <- as.numeric(facebook)


# Append fb_vec to li_vec: social_vec
social_vec <- append(li_vec, fb_vec)
social_vec

# Sort social_vec
sort(social_vec, decreasing=T)
```

### Find the error (2)

```{r}
# Fix me
round(sum(unlist(list(1.1,3,5))))

# Fix me
rep(seq(1, 7, by = 2), times = 7)

```

### Beat Gauss using R

```{r}

# Create first sequence: seq1
seq1 <- seq(1,500,by=3)
seq1
# Create second sequence: seq2
seq2 <- seq(1200, 900, by=-7)
seq2

# Calculate total sum of the sequences
sum(c(seq1, seq2))

```

## Regular Expressions

  - These are just sequences of characters and meta-characters that can match a pattern
  - They are used for
    - pattern existence
    - pattern replacement
    - pattern extraction
  
### grepl & grep

```{r}

# The emails vector has already been defined for you
emails <- c("john.doe@ivyleague.edu", "education@world.gov", "dalai.lama@peace.org", 
            "invalid.edu", "quant@bigdatacollege.edu", "cookie.monster@sesame.tv")

# Use grepl() to match for "edu"
grepl("edu",emails)

# Use grep() to match for "edu", save result to hits
hits <- grep("edu",emails)
hits

# Subset emails using hits
emails[hits]

```

### grepl & grep (2)

```{r}

# The emails vector has already been defined for you
emails <- c("john.doe@ivyleague.edu", "education@world.gov", "dalai.lama@peace.org", 
            "invalid.edu", "quant@bigdatacollege.edu", "cookie.monster@sesame.tv")

# Use grep() to match for .edu addresses more robustly
grep("@.*\\.edu$",emails)

# Use grepl() to match for .edu addresses more robustly, save result to hits
hits <- grepl("@.*\\.edu$",emails)
hits

# Subset emails using hits
emails[hits]

```

### sub and gsub

```{r}

# The emails vector has already been defined for you
emails <- c("john.doe@ivyleague.edu", "education@world.gov", "dalai.lama@peace.org", 
            "invalid.edu", "quant@bigdatacollege.edu", "cookie.monster@sesame.tv")

# Use sub() to convert the email domains to datacamp.edu (attempt 1)
# The @ character is removed
sub("@.*\\.edu$","datacamp.edu",emails)

# Use sub() to convert the email domains to datacamp.edu (attempt 2)
# Thats better
sub("@.*\\.edu$","@datacamp.edu",emails)

```

### sub and gsub (2)

```{r}

awards <- c("Won 1 Oscar.",
  "Won 1 Oscar. Another 9 wins & 24 nominations.",
  "1 win and 2 nominations.",
  "2 wins & 3 nominations.",
  "Nominated for 2 Golden Globes. 1 more win & 2 nominations.",
  "4 wins & 1 nomination.")

sub(".*\\s([0-9]+)\\snomination.*$", "\\1", awards)

```

## Times and Dates

### Right here, right now

```{r}

# Get the current date: today
today <- Sys.Date()

# See what today looks like under the hood
unclass(today)

# Get the current time: now
now <- Sys.time()

# See what now looks like under the hood
unclass(now)

```

### Create and format times

  - This will be helpful...
    - look up the format values in `strptime`. 
    - e.g. `?strptime`
    - There are a lot

```{r}

# Definition of character strings representing dates
str1 <- "May 23, '96"
str2 <- "2012-3-15"
str3 <- "30/January/2006"

# Convert the strings to dates: date1, date2, date3
date1 <- as.Date(str1, format = "%b %d, '%y")
date2 <- as.Date(str2, format = "%Y-%m-%d")
date3 <- as.Date(str3, format = "%d/%B/%Y")

date1
date2
date3

# Convert dates to formatted strings

# Get the day of week
format(date1, "%A")

# Get the day of the month
format(date2, "%d")

# Get the month abv and year
format(date3, "%b %Y")

```



```{r}

# Definition of character strings representing times
str1 <- "May 23, '96 hours:23 minutes:01 seconds:45"
str2 <- "2012-3-12 14:23:08"

# Convert the strings to POSIXct objects: time1, time2
time1 <- as.POSIXct(str1, format = "%B %d, '%y hours:%H minutes:%M seconds:%S")
time2 <- as.POSIXct(str2, format = "%Y-%m-%d %T")

time1
time2

# Convert times to formatted strings

# Get the minnute
format(time1, "%M")

# Get the hour:minute and AM/PM
format(time2, "%I:%M %p")

```

### Calculations with Dates

```{r}

# day1, day2, day3, day4 and day5 are already available in the workspace
day1 <- as.Date("2017-03-12")
day2 <- as.Date("2017-03-14")
day3 <- as.Date("2017-03-19")
day4 <- as.Date("2017-03-25")
day5 <- as.Date("2017-03-30")

# Difference between last and first pizza day
day5-day1

# Create vector pizza
pizza <- c(day1, day2, day3, day4, day5)
pizza

# Create differences between consecutive pizza days: day_diff
day_diff <- diff(pizza)
day_diff

# Average period between two consecutive pizza days
mean(day_diff)

```

### Calculations with Times

```{r}

# login and logout are already defined in the workspace
login <- c(as.POSIXct("2017-03-16 10:18:04 UTC"), 
            as.POSIXct("2017-03-21 09:14:18 UTC"),
            as.POSIXct("2017-03-21 12:21:51 UTC"), 
            as.POSIXct("2017-03-21 12:37:24 UTC"),
            as.POSIXct("2017-03-23 21:37:55 UTC"))

logout <- c(as.POSIXct("2017-03-16 10:56:29 UTC"),
            as.POSIXct("2017-03-21 09:14:52 UTC"),
            as.POSIXct("2017-03-21 12:35:48 UTC"), 
            as.POSIXct("2017-03-21 13:17:22 UTC"),
            as.POSIXct("2017-03-23 22:08:47 UTC"))
 
# Calculate the difference between login and logout: time_online
time_online <- logout - login

# Inspect the variable time_online
time_online

# Calculate the total time online
sum(time_online)

# Calculate the average time online
mean(time_online)

```

### Time is of the essence

```{r}

astro <- c("20-Mar-2015", "25-Jun-2015", "23-Sep-2015", "22-Dec-2015")
names(astro) <- c("spring", "summer","fall","winter") 
astro

meteo <- c("March 1, 15", "June 1, 15", "September 1, 15", "December 1, 15")
names(meteo) <- c("spring", "summer", "fall", "winter")
meteo
    
# Convert astro to vector of Date objects: astro_dates
astro_dates <- as.Date(astro, "%d-%b-%Y")
str(astro)
str(astro_dates)

# Convert meteo to vector of Date objects: meteo_dates
meteo_dates <- as.Date(meteo, "%B %d, %y")
str(meteo)
str(meteo_dates)

# Calculate the maximum absolute difference between astro_dates and meteo_dates
max(abs(astro_dates - meteo_dates))

```

&nbsp; &nbsp;

***  

## The End

  - Woof, thats a lot!
    - And thats just a sample of some key programming topics in R
  - Theres a lot more to learn in each of those topics
    - Especially the funtions and packages
    - There are a lot of key packages to work with dataframes or times
    - There are packages for everything really
  - This doc is useful to me as a reference
    - The class was best forlearning these things
    - Now I just need to see it again a few times to really get it
    - And I can look back here if ever need a little refresher in the future
    
&nbsp; &nbsp;

***  