---
title: "Importing Data in R (Part 2)"
author: "William Surles"
date: "2017-07-06"
output: 
 html_document:
  self_contained: yes
  theme: flatly
  highlight: tango
---

***
> This document was updated on **`r Sys.Date()`**.

***

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval=T, echo=T, cache=T, message=F, warning=F)

library(haven)
```

# Introduction

  - Course notes for [Importing Data in R (Part 2)](https://www.datacamp.com/courses/importing-data-in-r-part-2)

## Whats Covered

&nbsp; &nbsp;

***  

## quiz stuff


```{r}

sugar <- read_stata('http://assets.datacamp.com/production/course_1478/datasets/trade.dta')

plot(sugar$Import, sugar$Weight_I)

```

They are relatively positively correlated

## Do you know your data?

```{r}

## How many observations/individuals from Bulgaria have an income above 1000?
# nrow(subset(edu_equal_1, ethnicity_head == 'Bulgaria' & income > 1000))

```

## Excursion: Correlation

```{r}

## What is the correlation coefficient for the two numerical variables gdp and f_illit (female illiteracy rate)?
# cor(demo$gdp, demo$f_illit)

```

&nbsp; &nbsp;

***  
# Importing data from databases (Part 1)
***  

## Establish a connection
## Inspect the connection
## Import table data
## List table data
## List the database tables
## Import users
## Import all tables
## How do the tables relate?

&nbsp; &nbsp;

***  
# Importing data from databases (Part 2)
***  

## SQL Queries from inside R
## Query tweater (1)
## Query tweater (2)
## Query tweater (3)
## Query tweater (4)
## Join the query madness!
## DBI internals
## Send - Fetch - Clear
## Be polite and ...

&nbsp; &nbsp;

***  
# Importing data from the web (Part 1)
***  

## Import flat files from the web
## Secure importing
## Downloading files
## Import Excel files form the web
## Downloading any file, secure or not
## Reading a txt file from the web
## HTTP? httr! (1)
## HTTP? httr! (2)

&nbsp; &nbsp;

***  
# Importing data from the web (Part 2)
***  

## APIs & JSON
## From JSON to R
## Quandl API
## OMDb API
## JSON & jsonlite
## JSON practice (1)
## JSON practice (2)
## toJSON() 
## Minify and prettify

&nbsp; &nbsp;

***  
# Importing data from statistical software packages
***  

## haven
## Import SAS data with haven
## Import STATA data with haven
## What does the graphic tell?
## Import SPSS data with haven
## Factorize, round two
## foreign
## Import STATA data with foreign (1)
## Import STATA data with foreign (2)
## Do you know your data?
## Import SPSS data with foreign (1)
## Excursion: COrrelation
## Import SPSS data with foreign (2)
