---
title: "Exploratory Data Analysis in R: Case Study"
author: "William Surles"
date: "2017-08-014"
output: 
 html_document:
  self_contained: yes
  theme: flatly
  highlight: tango
  toc: true
  toc_float: true
  toc_depth: 3
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(eval=T, echo=T, cache=T, message=F, warning=F)
```

# Introduction

  - Course notes from the [Exploratory Data Analysis in R: Case Study](https://www.datacamp.com/courses/exploratory-data-analysis-in-r-case-study) course on DataCamp
  
## Whats Covered

  - Data cleaning and summarizing with dplyr
  - Data visualization with ggplot2
  - Tidy modeling with broom
  - Joining and tidying

## Libraries and Data

```{r} 

library(dplyr)
library(tidyr)
library(purrr)
library(ggplot2)
library(broom)
library(countrycode)
library(stringr)
library(data.table)

votes <- readRDS('data/votes.rds')
descriptions <- readRDS('data/descriptions.rds')

```


&nbsp; &nbsp;

***  
# Data cleaning and summarizing with dplyr
***  

## The United Nations voting dataset

### -- Filtering rows

  - 1 = Yes
  - 2 = Abstain
  - 3 = No
  - 8 = Not present
  - 9 = Not a member

```{r}

# Load the dplyr package

# Print the votes dataset
votes

# Filter for votes that are "yes", "abstain", or "no"
votes %>%
  filter(vote <= 3)

```

### -- Adding a year column

  - first session was in 1946. That is session 1.
  - There is one session per year. 
  - So, `year = session + 1945`
  
```{r}

# Add another %>% step to add a year column
votes %>%
  filter(vote <= 3) %>%
  mutate(year = session + 1945)

```

### -- Adding a country column

  - The class country codes had the "United States" for some reason 
  - I need to drop "of America" from the country code here so the class code works for the rest of the doc
  
```{r}

# Load the countrycode package

# Convert country code 100
countrycode(100, "cown", "country.name")

# Add a country column within the mutate: votes_processed
votes_processed <- votes %>%
  filter(vote <= 3) %>%
  mutate(
    year = session + 1945,
    country = countrycode(ccode, "cown", "country.name"),
    country = recode(country, 
                     'United States of America' =  'United States',
                     'United Kingdom of Great Britain and Northern Ireland' = 'United Kingdom')
    )


```

## Grouping and summarizing

### -- Summarizing the full dataset

```{r}

# Print votes_processed
votes_processed

# Find total and fraction of "yes" votes
votes_processed %>%
  summarise(
    total = n(),
    percent_yes = mean(vote == 1)
    )

```

### -- Summarizing by year

```{r}

# Change this code to summarize by year
votes_processed %>%
  group_by(year) %>%
  summarize(total = n(),
            percent_yes = mean(vote == 1))

```

### -- Summarizing by country

```{r}

# Summarize by country: by_country
by_country <- votes_processed %>%
  group_by(country) %>%
  summarize(total = n(),
            percent_yes = mean(vote == 1))

by_country

```

## Sorting and filtering summarized data

### -- Sorting by percentage of "yes" votes

```{r}

# You have the votes summarized by country
by_country <- votes_processed %>%
  group_by(country) %>%
  summarize(total = n(),
            percent_yes = mean(vote == 1))

# Print the by_country dataset
by_country

# Sort in ascending order of percent_yes
by_country %>%
  arrange(percent_yes)

# Now sort in descending order
by_country %>%
  arrange(desc(percent_yes))

```

### -- Filtering summarized output

```{r}

# Filter out countries with fewer than 100 votes
by_country %>%
  arrange(percent_yes) %>%
  filter(total >= 100)
  
```


&nbsp; &nbsp;

***  
# Data visualization with ggplot2
***  

## Visualization with ggplot2

### -- Plotting a line over time

```{r}

# Define by_year
by_year <- votes_processed %>%
  group_by(year) %>%
  summarize(total = n(),
            percent_yes = mean(vote == 1))

# Load the ggplot2 package
library(ggplot2)

# Create line plot
ggplot(by_year, aes(x = year, y = percent_yes)) +
  geom_line()

```

### -- Other ggplot2 layers

```{r}

# Change to scatter plot and add smoothing curve
ggplot(by_year, aes(year, percent_yes)) +
  geom_point() +
  geom_smooth()

```

## Visualizing by country

### -- Summarizing by year and country

```{r}

# Group by year and country: by_year_country
by_year_country <- votes_processed %>%
  group_by(year, country) %>%
  summarize(total = n(),
            percent_yes = mean(vote == 1))

```

### -- Plotting just the UK over time

```{r}

# Start with by_year_country dataset
by_year_country <- votes_processed %>%
  group_by(year, country) %>%
  summarize(total = n(),
            percent_yes = mean(vote == 1))

# Print by_year_country
by_year_country

# Create a filtered version: UK_by_year
UK_by_year <- by_year_country %>%
  filter(country == 'United Kingdom')

# Line plot of percent_yes over time for UK only
ggplot(UK_by_year, aes(x = year, y = percent_yes)) +
  geom_line()

```

### -- Plotting multiple countries

```{r}

# Vector of four countries to examine
countries <- c("United States", "United Kingdom",
               "France", "India")

# Filter by_year_country: filtered_4_countries
filtered_4_countries <- by_year_country %>%
  filter(country %in% countries)

# Line plot of % yes in four countries
ggplot(filtered_4_countries, aes(x = year, y = percent_yes, color = country)) +
  geom_line()

```

## Faceting

### -- Faceting by country

```{r}

# Vector of six countries to examine
countries <- c("United States", "United Kingdom",
               "France", "Japan", "Brazil", "India")

# Filtered by_year_country: filtered_6_countries
filtered_6_countries <- by_year_country %>%
  filter(country %in% countries)

# Line plot of % yes over time faceted by country
ggplot(filtered_6_countries, aes(x = year, y = percent_yes)) +
  geom_line() + 
  facet_wrap(~ country)

```

### -- Faceting with free y-axis

```{r}

# Vector of six countries to examine
countries <- c("United States", "United Kingdom",
               "France", "Japan", "Brazil", "India")

# Filtered by_year_country: filtered_6_countries
filtered_6_countries <- by_year_country %>%
  filter(country %in% countries)

# Line plot of % yes over time faceted by country
ggplot(filtered_6_countries, aes(year, percent_yes)) +
  geom_line() +
  facet_wrap(~ country, scales = "free_y")

```

### -- Choose your own countries

```{r}

# Add three more countries to this list
countries <- c("United States", "United Kingdom",
               "France", "Germany", "China", "India", "Argentina", "Canada","Uruguay")

# Filtered by_year_country: filtered_countries
filtered_countries <- by_year_country %>%
  filter(country %in% countries)

# Line plot of % yes over time faceted by country
ggplot(filtered_countries, aes(year, percent_yes)) +
  geom_line() +
  facet_wrap(~ country)

```


&nbsp; &nbsp;

***  
# Tidy modeling with broom
***  

## Linear regression

  > Visualization can surprise you, but it doesn't scale well. Modeling scales well, but it can't surprise you. - Hadley Wickham

### -- Linear regression on the United States

```{r}

# Percentage of yes votes from the US by year: US_by_year
US_by_year <- by_year_country %>%
  filter(country == "United States")

# Print the US_by_year data
US_by_year

# Perform a linear regression of percent_yes by year: US_fit
US_fit <- lm(percent_yes ~ year, data = US_by_year)

# Perform summary() on the US_fit object
summary(US_fit)

```

### -- Finding the slope of a linear regression

  - What is th estimated slope of this relationship? Said differently, what's the estimated change each year of the probability of the US voting "yes"?
    - -0.0006

### -- Finding the p-value of a linear regression

  - In this linear model, what is the p-value of the relationship between year and percent_yes?
    - 1.37e-07
    
## Tidying models with broom

### -- Tidying a linear regression model

```{r}

# Load the broom package

# Call the tidy() function on the US_fit object
tidy(US_fit)

```

### -- Combining models for multiple countries

```{r}

# Linear regression of percent_yes by year for US
US_by_year <- by_year_country %>%
  filter(country == "United States")

US_fit <- lm(percent_yes ~ year, US_by_year)

# Fit model for the United Kingdom
UK_by_year <- by_year_country %>%
  filter(country == "United Kingdom")

by_year_country %>% 
  filter(country %like% c('United'))

UK_fit <- lm(percent_yes ~ year, UK_by_year)

# Create US_tidied and UK_tidied
US_tidied <- tidy(US_fit)
UK_tidied <- tidy(UK_fit)

# Combine the two tidied models
bind_rows(US_tidied, UK_tidied)

```

## Nesting for multiple models

### -- Nesting a data frame

```{r}

# Load the tidyr package
library(tidyr)

# Nest all columns besides country
by_year_country %>%
  nest(-country)

```

### -- List columns

```{r}

# All countries are nested besides country
nested <- by_year_country %>%
  nest(-country)

nested$data[[1]]
nested$data[[1]]$percent_yes

# Print the nested data for Brazil
nested$data[nested$country == 'Brazil']

```

### -- Unnesting

```{r}

# All countries are nested besides country
nested <- by_year_country %>%
  nest(-country)

# Unnest the data column to return it to its original form
nested %>%
  unnest()

```

## Fitting multiple models

```{r}

v <- list(1,2,3)
map(v, ~ . * 10)

```

### -- Performing linear regression on each nested dataset

  - Oddly, when I just used `nest(-country)` the year column was not in the tibble
  - using group by first is a little more explicit and maintains all of the columns other than country
  
```{r}

# Load tidyr and purrr
library(tidyr)
library(purrr)

# Perform a linear regression on each item in the data column
by_year_country %>%
  group_by(country) %>%
  nest() %>%
  mutate(
    model = map(data, ~ lm(percent_yes ~ year, data = .))
    )
    
```

### -- Tidy each linear regression model

```{r}

# Load the broom package

# Add another mutate that applies tidy() to each model
by_year_country %>%
  group_by(country) %>%
  nest() %>%
  mutate(
    model = map(data, ~ lm(percent_yes ~ year, data = .)),
    tidied = map(model, tidy)
    )

```

### -- Unnesting a data frame

```{r}

# Add one more step that unnests the tidied column
country_coefficients <- by_year_country %>%
  group_by(country) %>%
  nest() %>%
  mutate(model = map(data, ~ lm(percent_yes ~ year, data = .)),
         tidied = map(model, tidy)) %>%
  unnest(tidied)

# Print the resulting country_coefficients variable
country_coefficients
```

## Working with many tidy models

  - We need to do a multiple hypotesis correction
    - As some p-values ca be less than 0.5 by chance
    - This concept is outside the scope of this course
    - For now, know that R has a built in funciton `p.adjust` that can do this

### -- Filtering model terms

```{r}

# Print the country_coefficients dataset
country_coefficients

# Filter for only the slope terms
country_coefficients %>%
  filter(term == 'year')

```

### -- Filtering for significant countries

```{r}

# Filter for only the slope terms
slope_terms <- country_coefficients %>%
  filter(term == "year")

# Add p.adjusted column, then filter
slope_terms %>%
  mutate(p.adjusted = p.adjust(p.value)) %>% 
  data.frame() %>% head()
  
slope_terms %>%
  mutate(p.adjusted = p.adjust(p.value)) %>% 
  filter(p.adjusted < 0.05)

```

### -- Sorting by slope

```{r}

# Filter by adjusted p-values
filtered_countries <- country_coefficients %>%
  filter(term == "year") %>%
  mutate(p.adjusted = p.adjust(p.value)) %>%
  filter(p.adjusted < .05)

# Sort for the countries increasing most quickly
filtered_countries %>%
  arrange(desc(estimate))

# Sort for the countries decreasing most quickly
filtered_countries %>%
  arrange(estimate)

```

&nbsp; &nbsp;

***  
# Joining and tidying
***  

## Joining datasets

### -- Joining datasets with inner_join

```{r}

# Load dplyr package

# Print the votes_processed dataset
votes_processed

# Print the descriptions dataset
descriptions

# Join them together based on the "rcid" and "session" columns
votes_joined <- inner_join(votes_processed, descriptions, by = c("rcid", "session"))

```

### -- Filtering the joined dataset

  - me: Palestinian conflict
  - nu: Nuclear weapons and nuclear material
  - di: Arms control and disarmament
  - hr: Human rights
  - co: Colonialism
  - ec: Economic development
  
```{r}

# Filter for votes related to colonialism
votes_joined %>% 
  filter(co == 1)

```

### -- Visualizing colonialism votes

```{r}

# Load the ggplot2 package
library(ggplot2)

# Filter, then summarize by year: US_co_by_year
US_co_by_year <- votes_joined %>%
  filter(
    co == 1,
    country == "United States"
    ) %>%
  group_by(year) %>%
  summarize(percent_yes = mean(vote == 1))

# Graph the % of "yes" votes over time
ggplot(US_co_by_year, aes(x = year, y = percent_yes)) +
  geom_line()

```

## Tidy data

### -- Using gather to tidy a dataset

```{r}

# Load the tidyr package
library(tidyr)

# Gather the six me/nu/di/hr/co/ec columns
votes_joined %>%
  gather(topic, has_topic, one_of("me","nu","di","hr","co","ec"))

# Perform gather again, then filter
votes_gathered <- votes_joined %>%
  gather(topic, has_topic, one_of("me","nu","di","hr","co","ec")) %>%
  filter(has_topic == 1)

```

### -- Recoding the topics

  - `recode` is awesome

```{r}

example <- c("apple", "banana", "apple", "orange")
recode(example,
       apple = "plum",
       banana = "grape")

```

  
```{r}

# Replace the two-letter codes in topic: votes_tidied
votes_tidied <- votes_gathered %>%
  mutate(topic = recode(topic,
                        me = "Palestinian conflict",
                        nu = "Nuclear weapons and nuclear material",
                        di = "Arms control and disarmament",
                        hr = "Human rights",
                        co = "Colonialism",
                        ec = "Economic development"))

```

### -- Summarize by country, year, and topic

```{r}

# Print votes_tidied
votes_tidied

# Summarize the percentage "yes" per country-year-topic
by_country_year_topic <- votes_tidied %>%
  group_by(country, year, topic) %>%
  summarize(
    total = n(),
    percent_yes = mean(vote == 1)
    ) %>%
  ungroup()

# Print by_country_year_topic
by_country_year_topic

```

### -- Visualizing trends in topics for one country

```{r}

# Load the ggplot2 package
library(ggplot2)

# Filter by_country_year_topic for just the US
US_by_country_year_topic <- by_country_year_topic %>%
  filter(country == "United States")

# Plot % yes over time for the US, faceting by topic
ggplot(US_by_country_year_topic, aes(x = year, y = percent_yes)) + 
  geom_line() +
  facet_wrap(~ topic)

```

## Tidy modeling by topic and country

### -- Nesting by topic and country

```{r}

# Load purrr, tidyr, and broom
library(purrr)
library(tidyr)
library(broom)

# Print by_country_year_topic
by_country_year_topic

# Fit model on the by_country_year_topic dataset
country_topic_coefficients <- by_country_year_topic %>%
  nest(-country, -topic) %>%
  mutate(
    model = map(data, ~ lm(percent_yes ~ year, data = .)),
    tidied = map(model, tidy)
    ) %>%
  unnest(tidied)
    

# Print country_topic_coefficients
country_topic_coefficients

```

### -- Interpreting tidy models

```{r}

# Create country_topic_filtered
country_topic_filtered <- country_topic_coefficients %>% 
  filter(term == 'year') %>%
  mutate(p.adjusted = p.adjust(p.value)) %>%
  filter(p.adjusted < 0.05)

country_topic_filtered

```

### -- Steepest trends by topic

  - Which combination of country and topic has the steepes downward trend?
  
```{r}

country_topic_filtered %>% arrange(estimate)

```

  - Vanuata on the Palestinian conflict
  
### -- Checking models visually

```{r}

# Create vanuatu_by_country_year_topic
vanuatu_by_country_year_topic <- by_country_year_topic %>%
  filter(country == "Vanuatu")

# Plot of percentage "yes" over time, faceted by topic
ggplot(vanuatu_by_country_year_topic, aes(x = year, y = percent_yes)) +
  geom_line() + 
  facet_wrap(~ topic)

```

## Conclusion

  - Cool class. I learned some neat tricks.
  - It just barely touches the process or working with models in dplyr and broom
  - It seems like there could be another class on more advanced modeling and use cases of dplyr and broom. 
  
### -- A chart I just wanted to make

```{r, fig.width = 9}

filtered_by_country_year_topic <- by_country_year_topic %>%
  filter(country %in% c("United States", "United Kingdom", "France", "Germany"))

# Plot of percentage "yes" over time, faceted by topic
ggplot(filtered_by_country_year_topic, aes(x = year, y = percent_yes, color = country)) +
  geom_point() + 
  geom_smooth() + 
  facet_wrap(~ topic)

```

