---
title: "Data Visualization with ggplot2 (Part 3-C)"
author: "William Surles"
date: "2017-08-02"
output: 
 html_document:
  self_contained: yes
  theme: flatly
  highlight: tango
  toc: true
  toc_float: true
  toc_depth: 3
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(eval=T, echo=T, cache=T, message=F, warning=F)

```

# Introduction

  - Course notes from the [Data Vizualization with ggplot2 (Part 3)](https://www.datacamp.com/courses/data-visualization-with-ggplot2-part-3) course on DataCamp
  
## Whats Covered

Part A:

  - Statistical plots
    - Aesthetics review, 
    - box plots, density plots 
    - multiple groups/variables
  - Plots for specific data types (Part 1)
    - graphics of large data
    - Ternary plots
    - Network plots
    - Diagnostic plots

Part B:

  - Plots for specific data types (Part 2)
    - choropleths
    - cartographic maps
    - animations
    
**Part C:**

  - ggplot2 internals
    - grid graphics, grid grapshics in ggplot2
    - ggplot objects
    - gridExtra
  - Data Munging and Visualization Case Study
    - Bag plot case study, weather case study

## Libraries and Data

```{r} 

source("create_datasets.R")
load('data/test_datasets.RData')

library(readr)
library(dplyr)
library(ggplot2)
library(purrr)

library(grid)
library(gtable)
library(aplpack)

```


&nbsp; &nbsp;

***  
# ggplot2 internals
***  

## Grid Graphics

### -- Viewport basics (1)

```{r}
# the grid library is loaded

# Draw rectangle in null viewport
grid.rect(gp = gpar(fill = "grey90"))

# Write text in null viewport
grid.text("null viewport")

# Draw a line
grid.lines(x = c(0, 0.75), y = c(0.25, 1),
          gp = gpar(lty = 2, col = "red"))

```

### -- Viewport basics (2)

```{r}

# Populate null viewport
grid.rect(gp = gpar(fill = "grey90"))
grid.text("null viewport")
grid.lines(x = c(0,0.75), y = c(0.25, 1),
           gp = gpar(lty = 2, col = "red"))

# Create new viewport: vp
vp <- viewport(x = 0.5, y = 0.5, width = 0.5, height = 0.5, just = "center")

# Push vp
pushViewport(vp)

# Populate new viewport with rectangle
grid.rect(gp = gpar(fill = "blue"))

```

### -- Build a plot from scratch (1)

```{r}

# 1 - Create plot viewport: pvp
mar <- c(5, 4, 2, 2)
pvp <- plotViewport(mar)

# 2 - Push pvp
pushViewport(pvp)

# 3 - Add rectangle
grid.rect(gp = gpar(fill = "grey80"))

# Create data viewport: dvp
dvp <- dataViewport(xData = mtcars$wt, yData = mtcars$mpg)

# 4 - Push dvp
pushViewport(dvp)

# Add two axes
grid.xaxis()
grid.yaxis()

```

### -- Build a plot from scratch (2)

```{r}

# Work from before
pushViewport(plotViewport(c(5, 4, 2, 2)))
grid.rect(gp = gpar())
pushViewport(dataViewport(xData = mtcars$wt, yData = mtcars$mpg))
grid.xaxis()
grid.yaxis()

# 1 - Add text to x axis
grid.text("Weight", y = unit(-3, "lines"))

# 2 - Add text to y axis
grid.text("MPG", x = unit(-3, "lines"), rot = 90)

# 3 - Add points
grid.points(x = mtcars$wt, y = mtcars$mpg, pch = 16)

```

### -- Modifying a plot with grid.edit

```{r}

# Work from before
pushViewport(plotViewport(c(5, 4, 2, 2)))
grid.rect(gp = gpar())
pushViewport(dataViewport(xData = mtcars$wt, yData = mtcars$mpg))
grid.xaxis()
grid.yaxis()

# Work from before - add names
grid.text("Weight", y = unit(-3, "lines"), name = "xaxis")
grid.text("MPG", x = unit(-3, "lines"), rot = 90, name = "yaxis")
grid.points(x = mtcars$wt, y = mtcars$mpg, pch = 16, name = "datapoints")

# Edit "xaxis"
grid.edit("xaxis", label = "Weight (1000 lbs)")

# Edit "yaxis"
grid.edit("yaxis", label = "Miles/(US) gallon")

# Edit "datapoints"
grid.edit("datapoints", 
  gp = gpar(col = "#C3212766", cex = 2))

```

## Grid Graphics in ggplot2

### -- Exploring the gTable

```{r}

# A simple plot p
p <- ggplot(mtcars, aes(x = wt, y = mpg, col = factor(cyl))) + geom_point()

# Create gtab with ggplotGrob()
gtab <- ggplotGrob(p)

# Print out gtab
gtab

# Extract the grobs from gtab: gtab
g <- gtab$grob
 
# Draw only the legend
legend_index <- which(vapply(g, inherits, what = "gtable", logical(1)))
grid::grid.draw(g[[legend_index]])

```

### -- Modifying the gTable

  - This seems to not work in the notebook. Its very weird. Something I will need to dig into more later
    - It seems to work fine in Rstudio though
  - Also, the `grid` library seems really flaky. None of its functions were working in this code chunk
    - They work fine in some of the above code chuncks
    - I reloaded it here and that seems to help. 
  - Over messing with the grid graphics seems like flaky business. 
  
```{r}

library(grid)

# Code from before
p <- ggplot(mtcars, aes(x = wt, y = mpg, col = factor(cyl))) + geom_point()
gtab <- ggplotGrob(p)
g <- gtab$grobs
legend_index <- which(vapply(g, inherits, what = "gtable", logical(1)))
grid::grid.draw(g[[legend_index]])

# the gtable library is loaded

# 1 - Show layout of legend grob
## this is not printing in the html but it shows fine in the console. 
gtable_show_layout(g[[legend_index]])

# Create text grob
my_text <- grid::textGrob(label = "Motor Trend, 1974", gp = gpar(fontsize = 7, col = "gray25"))

# 2 - Use gtable_add_grob to modify original gtab
new_legend <- gtable_add_grob(
  gtab$grobs[[legend_index]], my_text, 3, 2)

# 3 - Update in gtab
gtab$grobs[[legend_index]] <- new_legend

# 4 - Draw gtab
grid::grid.draw(gtab)

```

## ggplot Objects

### -- Exploring ggplot objects

```{r}

# Simple plot p
p <- ggplot(mtcars, aes(x = wt, y = mpg, col = factor(cyl))) + geom_point()

# Examine class() and names()
class(p)
names(p)


# Print the scales sub-list
p$scales$scales

# Update p
p <- p +
  scale_x_continuous("Length", limits = c(4, 8), expand = c(0, 0)) +
  scale_y_continuous("Width", limits = c(2, 4.5), expand = c(0, 0))

# Print the scales sub-list
p$scales$scales

```

### -- ggplot_build and ggplot_gtable

```{r}

# Box plot of mtcars: p
p <- ggplot(mtcars, aes(x = factor(cyl), y = wt)) + geom_boxplot()

# Create pbuild
pbuild <- ggplot_build(p)

# a list of 3 elements
names(pbuild)

# Print out each element in pbuild
# In the exercise they have panel as an option. But its `plot now`
pbuild$data
pbuild$layout
## This should show the plot. Not working in the markdown doc though. 
pbuild$plot

# Create gtab from pbuild
gtab <- ggplot_gtable(pbuild)

# Draw gtab
grid::grid.draw(gtab)

```

### -- Extracting details

  - This is cool becasue we can get exact values of the box plot. 
  - There are many ways to calculate quartiles and if we do it ourselves outside the plot, its likely we will have different values
  
```{r}

# Box plot of mtcars: p
p <- ggplot(mtcars, aes(x = factor(cyl), y = wt)) + geom_boxplot()

# Build pdata
pdata <- ggplot_build(p)$data

# confirm that the first element of the list is a data frame
class(pdata[[1]])

# Isolate this data frame
my_df <- pdata[[1]]

# The x labels
my_df$group <- ggplot_build(p)$layout$panel_ranges[[1]]$x.labels

# Print out specific variables
my_df[c(1:6, 11)]

```

## gridExtra

### -- Arranging plots (1)

```{r}

# Add a theme (legend at the bottom)
g1 <- ggplot(mtcars, aes(wt, mpg, col = cyl)) +
  geom_point(alpha = 0.5) +
  theme(legend.position = "bottom")

# Add a theme (no legend)
g2 <- ggplot(mtcars, aes(disp, fill = cyl)) +
  geom_histogram(position = "identity", alpha = 0.5, binwidth = 20) +
  theme(legend.position = "none")

# Load gridExtra
library(gridExtra)

# Call grid.arrange()
grid.arrange(g1, g2, ncol = 2)

```

### -- Arranging plots (2)

```{r}

# ggplot2, grid and gridExtra have been loaded for you
# Definitions of g1 and g2
g1 <- ggplot(mtcars, aes(wt, mpg, col = cyl)) +
  geom_point() +
  theme(legend.position = "bottom")

g2 <- ggplot(mtcars, aes(disp, fill = cyl)) +
  geom_histogram(binwidth = 20) +
  theme(legend.position = "none")

# Extract the legend from g1
my_legend <- ggplotGrob(g1)$grobs[[legend_index]]  

# Create g1_noleg
g1_noleg <- g1 + 
    theme(legend.position = "none")

# Calculate the height: legend_height
legend_height <- sum(my_legend$heights)

# Arrange g1_noleg, g2 and my_legend
grid.arrange(
  g1_noleg, g2, my_legend,
  layout_matrix = matrix(
    c(1,3,2,3), 
    ncol = 2),
    heights = unit.c(unit(1, "npc") - legend_height, legend_height))

```

&nbsp; &nbsp;

***  
# Data Munging and Visualization Case Study
***  

## Case Study I - Bag Plot

### -- Base package bag plot

```{r}

# test_datasets.RData has been loaded

# Call bagplot() on test_data
# The aplpack library has been loaded
head(ch5_test_data)
bagplot(ch5_test_data)

# Call compute.bagplot on test_data, assign to bag
bag <- compute.bagplot(ch5_test_data)

# Display information
bag$hull.loop
bag$hull.bag
bag$pxy.outlier

# Highlight components
points(bag$hull.loop, col = "green", pch = 16)
points(bag$hull.bag, col = "orange", pch = 16)
points(bag$pxy.outlier, col = "purple", pch = 16)

```

### -- Multilayer ggplot2 bag plot

```{r}

# bag and ch5_test_data are available

# Create data frames from matrices
hull.loop <- data.frame(x = bag$hull.loop[,1], y = bag$hull.loop[,2])
hull.bag <- data.frame(x = bag$hull.bag[,1], y = bag$hull.bag[,2])
pxy.outlier <- data.frame(x = bag$pxy.outlier[,1], y = bag$pxy.outlier[,2])

# Finish the ggplot command
ggplot(ch5_test_data, aes(x = x,  y = y)) +
  geom_polygon(data = hull.loop, fill = "green") +
  geom_polygon(data = hull.bag, fill = "orange") +
  geom_point(data = pxy.outlier, col = "purple", pch = 16, cex = 1.5)

```

### -- Creating ggproto functions

```{r}

# ggproto for StatLoop (hull.loop)
StatLoop <- ggproto(
  "StatLoop", 
  Stat,
  required_aes = c("x", "y"),
  compute_group = function(data, scales) {
    bag <- compute.bagplot(x = data$x, y = data$y)
    data.frame(x = bag$hull.loop[,1], y = bag$hull.loop[,2])
  })

# ggproto for StatBag (hull.bag)
StatBag <- ggproto(
  "StatBag", 
  Stat,
  required_aes = c("x", "y"),
  compute_group = function(data, scales) {
   bag <- compute.bagplot(x = data$x, y = data$y)
   data.frame(x = bag$hull.bag[,1], y = bag$hull.bag[,2])
  })

# ggproto for StatOut (pxy.outlier)
StatOut <- ggproto(
  "StatOut", 
  Stat,
  required_aes = c("x", "y"),
  compute_group = function(data, scales) {
   bag <- compute.bagplot(x = data$x, y = data$y)
   data.frame(x = bag$pxy.outlier[,1], y = bag$pxy.outlier[,2])
  })

```

### -- Creating stat_bag()

  - This all works fine in the class console. 
  - But locally I get the most vague error possible. 
  
```{r}

# StatLoop, StatBag and StatOut are available

# Combine ggproto objects in layers to build stat_bag()
stat_bag <- function(mapping = NULL, data = NULL, geom = polygon,
                     position = "identity", na.rm = FALSE, show.legend = NA,
                     inherit.aes = TRUE, loop = FALSE, ...) {
  list(
    # StatLoop layer
    layer(
      stat = StatLoop, data = data, mapping = mapping, geom = geom, 
      position = position, show.legend = show.legend, inherit.aes = inherit.aes,
      params = list(na.rm = na.rm, alpha = 0.35, col = NA, ...)
    ),
    # StatBag layer
    layer(
      stat = StatBag, data = data, mapping = mapping, geom = geom, 
      position = position, show.legend = show.legend, inherit.aes = inherit.aes,
      params = list(na.rm = na.rm, alpha = 0.35, col = NA, ...)
    ),
    # StatOut layer
    layer(
      stat = StatOut, data = data, mapping = mapping, geom = "point", 
      position = position, show.legend = show.legend, inherit.aes = inherit.aes,
      params = list(na.rm = na.rm, alpha = 0.7, col = NA, shape = 21, ...)
    )
  )
}

```

### -- Use stat_bag()

  - The stat_bag function is not working locally
  - I get a `Error: object of type 'closure' is not subsettable
    - This is the most vague error. I have no idea what is wrong in the multiple prior steps. 
  - Maybe I should gind a tutorial on this later when I want to use something like this and hope I can get it working right. 
  - After reading a bunch of discussions, it looks like some libraries may be out of sync. 
  
```{r, error=T}

# hull.loop, hull.bag and pxy.outlier are available
# stat_bag, ch5_test_data and ch5_test_data2 are available

# Previous method
ggplot(ch5_test_data, aes(x = x,  y = y)) +
  geom_polygon(data = hull.loop, fill = "green") +
  geom_polygon(data = hull.bag, fill = "orange") +
  geom_point(data = pxy.outlier, col = "purple", pch = 16, cex = 1.5)

# stat_bag
ggplot(data = ch5_test_data, aes(x = x, y = y)) +
  stat_bag(ch5_test_data, fill = 'black')

# stat_bag on test_data2
ggplot(ch5_test_data2, aes(x = x, y = y, fill = treatment)) +
  stat_bag()

```

## Case Study II - Weather (Part 1)

### -- Step 1: Read in data and examine

```{r eval=F}

# Import weather data
weather_data_urls <- c(
  'https://assets.datacamp.com/production/course_862/datasets/NYNEWYOR.txt',
  'https://assets.datacamp.com/production/course_862/datasets/FRPARIS.txt',
  'https://assets.datacamp.com/production/course_862/datasets/ILREYKJV.txt',
  'https://assets.datacamp.com/production/course_862/datasets/UKLONDON.txt'
)
weather_data_files <- c("data/NYNEWYOR.txt","data/FRPARIS.txt", "data/ILREYKJV.txt", "data/UKLONDON.txt")

download.file(weather_data_urls, weather_data_files)  

```



```{r}
# Check out the NY weather data
weather <- read.fwf('data/NYNEWYOR.txt',
                    header = F,
                    col.names = c("month", "day", "year", "temp"),
                    widths = c(14, 14, 13, 4))

# Check structure of weather
str(weather)

# Create past with two filter() calls
past <- weather %>%
  filter(!(month == 2 & day == 29)) %>%
  filter(year != 2016)
  
# Check structure of past
str(past)
```

### -- Step 2: Summarize history

```{r}

# Create new version of past
past_summ <- past %>%
  group_by(year) %>%
  mutate(yearday = 1:length(day)) %>%
  ungroup() %>%
  filter(temp != -99) %>%
  group_by(yearday) %>%
  mutate(max = max(temp),
         min = min(temp),
         avg = mean(temp),
         CI_lower = Hmisc::smean.cl.normal(temp)[2],
         CI_upper = Hmisc::smean.cl.normal(temp)[3]) %>%
  ungroup()

# Structure of past_summ
str(past_summ)
head(past_summ)

Hmisc::smean.cl.normal(past$temp)[2]
table(past$temp)
table(past_summ$temp)

```

### -- Step 3: Plot history

```{r}

# Adapt historical plot
ggplot(past_summ, aes(x = yearday, y = temp)) +
  geom_point(
    col = "#EED8AE", 
    alpha = 0.3,
    shape = 16) +
  geom_linerange(
    aes(ymin = CI_lower, ymax = CI_upper), 
    col = "#8B7E66")

```

### -- Step 4: Plot present

```{r}

# weather and past are available in your workspace

# Create present
present <- weather %>%
  filter(!(month == 2 & day == 29)) %>%
  filter(year == max(year)) %>%
  group_by(year) %>%
  mutate(yearday = 1:length(day)) %>%
  ungroup() %>%
  filter(temp != -99)

# Add geom_line to ggplot command
ggplot(past_summ, aes(x = yearday, y = temp)) + 
  geom_point(
    col = "#EED8AE", 
    alpha = 0.3, 
    shape = 16) +
  geom_linerange(
    aes(ymin = CI_lower, ymax = CI_upper), 
    col = "#8B7E66") + 
  geom_line(data = present)

```

### -- Step 5: Find new record highs

```{r}

# Create past_highs
past_highs <- past_summ %>%
  group_by(yearday) %>%
  summarise(past_high = max(temp))

# Create record_high
record_high <- present %>%
  left_join(past_highs) %>%
  filter(temp > past_high)
  
record_high

# Add record_high information to plot
ggplot(past_summ, aes(x = yearday, y = temp)) + 
  geom_point(col = "#EED8AE", alpha = 0.3, shape = 16) +
  geom_linerange(aes(ymin = CI_lower, ymax = CI_upper), col = "#8B7E66") +
  geom_line(data = present) +
  geom_point(data = record_high, col = "#CD2626")

```

### -- Step 6: Efficiently calculate record highs and lows

```{r}

# Create past_extremes
past_extremes <- past_summ %>%
  group_by(yearday) %>%
  summarise(past_low = min(temp),
            past_high = max(temp))

# Create record_high_low
record_high_low <- present %>%
  left_join(past_extremes) %>%
  mutate(record = ifelse(temp < past_low, 
                         "#0000CD",
                         ifelse(temp > past_high, 
                                "#CD2626", 
                                "#00000000")))

# Structure of record_high_low
str(record_high_low)
head(record_high_low)

# Add point layer of record_high_low
p <- ggplot(past_summ, aes(x = yearday, y = temp)) + 
  geom_point(col = "#EED8AE", alpha = 0.3, shape = 16) +
  geom_linerange(aes(ymin = CI_lower, ymax = CI_upper), col = "#8B7E66") +
  geom_line(data = present) +
  geom_point(data = record_high_low, aes(col = record)) +
  scale_color_identity()
p

```

### -- Step 7: Custom legend

```{r}

# Finish the function draw_pop_legend
draw_pop_legend <- function(x = 0.6, y = 0.2, width = 0.2, height = 0.2, fontsize = 10) {
  
  # Finish viewport() function
  pushViewport(viewport(x = x, y = y, width = width, height = height, just = "center"))

  legend_labels <- c("Past record high",
                     "95% CI range",
                     "Current year",
                     "Past years",
                     "Past record low")

  legend_position <- c(0.9, 0.7, 0.5, 0.2, 0.1)
  
  # Finish grid.text() function
  grid.text(label = legend_labels, x = 0.12, y = legend_position, 
            just = "left", 
            gp = gpar(fontsize = fontsize, col = "grey20"))
  
  # Position dots, rectangle and line
  point_position_y <- c(0.1, 0.2, 0.9)
  point_position_x <- rep(0.06, length(point_position_y))
  grid.points(x = point_position_x, y = point_position_y, pch = 16,
              gp = gpar(col = c("#0000CD", "#EED8AE", "#CD2626")))
  grid.rect(x = 0.06, y = 0.5, width = 0.06, height = 0.4,
            gp = gpar(col = NA, fill = "#8B7E66"))
  grid.lines(x = c(0.03, 0.09), y = c(0.5, 0.5),
             gp = gpar(col = "black", lwd = 3))
  
  # Add popViewport() for bookkeeping
  popViewport()
}

# Plotting object p, from previous exercise
p

# Call draw_pop_legend()
draw_pop_legend()

```

## Case Study II - Weather (Part 2)

### -- Step 1: clean_weather()

```{r}

# Finish the clean_weather function
clean_weather <- function(file) {
  weather <- read.fwf(file,
                      header = FALSE,
                      col.names = c("month", "day", "year", "temp"),
                      widths = c(14, 14, 13, 4))
  weather %>%
    filter(!(month == 2 & day == 29)) %>%
    group_by(year) %>%
    mutate(yearday = 1:length(day)) %>%
    ungroup() %>%
    filter(temp != -99)
}

# Import NYNEWYOR.txt: my_data
my_data <- clean_weather('data/NYNEWYOR.txt')

```

```{r}
ny_weather_data_url <- 'https://assets.datacamp.com/production/course_862/datasets/NYNEWYOR.txt'
download.file(ny_weather_data_url, 'data/NYNEWYOR.txt')

```

### -- Step 2: Historical data

```{r}

# Create the stats object
StatHistorical <- ggproto("StatHistorical", Stat,
                    compute_group = function(data, scales, params) {
                      data <- data %>%
                        filter(year != max(year)) %>%
                        group_by(x) %>%
                        mutate(ymin = Hmisc::smean.cl.normal(y)[3],
                               ymax = Hmisc::smean.cl.normal(y)[2]) %>%
                        ungroup()
                    },
                    required_aes = c("x", "y", "year"))

# Create the layer
stat_historical <- function(mapping = NULL, data = NULL, geom = "point",
                            position = "identity", na.rm = FALSE, show.legend = NA, 
                            inherit.aes = TRUE, ...) {
  list(
    layer(
      stat = "identity", data = data, mapping = mapping, geom = geom,
      position = position, show.legend = show.legend, inherit.aes = inherit.aes,
      params = list(na.rm = na.rm, col = "#EED8AE", alpha = 0.3, shape = 16, ...)
    ),
    layer(
      stat = StatHistorical, data = data, mapping = mapping, geom = "linerange",
      position = position, show.legend = show.legend, inherit.aes = inherit.aes,
      params = list(na.rm = na.rm, col = "#8B7E66", ...)
    )
  )
}

# Build the plot
my_data <- clean_weather("data/NYNEWYOR.txt")
ggplot(my_data, aes(x = yearday, y = temp, year = year)) +
  stat_historical()

```

### -- Step 3: Present data

```{r}

# Create the stats object
StatPresent <- ggproto("StatPresent", Stat,
                       compute_group = function(data, scales, params) {
                         data <- filter(data, year == max(year))
                       },
                       required_aes = c("x", "y", "year"))

# Create the layer
stat_present <- function(mapping = NULL, data = NULL, geom = "line",
                       position = "identity", na.rm = FALSE, show.legend = NA, 
                       inherit.aes = TRUE, ...) {
  layer(
    stat = StatPresent, data = data, mapping = mapping, geom = geom,
    position = position, show.legend = show.legend, inherit.aes = inherit.aes,
    params = list(na.rm = na.rm, ...)
  )
}

# Build the plot
my_data <- clean_weather("data/NYNEWYOR.txt")
ggplot(my_data, aes(x = yearday, y = temp, year = year)) +
  stat_historical() +
  stat_present()

```

### -- Step 4: Extremes

```{r}

# Create the stats object
StatExtremes <- ggproto("StatExtremes", Stat,
                        compute_group = function(data, scales, params) {
                          
                          present <- data %>%
                            filter(year == max(year)) 
                          
                          past <- data %>%
                            filter(year != max(year)) 
                          
                          past_extremes <- past %>%
                            group_by(x) %>%
                            summarise(past_low = min(y),
                                      past_high = max(y))
                          
                          # transform data to contain extremes
                          data <- present %>%
                            left_join(past_extremes) %>%
                            mutate(record = ifelse(y < past_low, 
                                                   "#0000CD", 
                                                   ifelse(y > past_high, 
                                                          "#CD2626", 
                                                          "#00000000")))
                        },
                        required_aes = c("x", "y", "year"))

# Create the layer
stat_extremes <- function(mapping = NULL, data = NULL, geom = "point",
                          position = "identity", na.rm = FALSE, show.legend = NA, 
                          inherit.aes = TRUE, ...) {
  layer(
    stat = StatExtremes, data = data, mapping = mapping, geom = geom,
    position = position, show.legend = show.legend, inherit.aes = inherit.aes,
    params = list(na.rm = na.rm, ...)
  )
}

# Build the plot
my_data <- clean_weather("data/NYNEWYOR.txt")
ggplot(my_data, aes(x = yearday, y = temp, year = year)) +
  stat_historical() +
  stat_present() +
  stat_extremes(aes(col = ..record..)) +
  scale_color_identity() # Colour specification

```

### -- Step 5: Re-use plotting style

```{r}

# File paths of all datasets
my_files <- c("data/NYNEWYOR.txt","data/FRPARIS.txt", "data/ILREYKJV.txt", "data/UKLONDON.txt")

# Build my_data with a for loop
my_data <- NULL
for (file in my_files) {
  temp <- clean_weather(file)
  temp$id <- sub(".txt", "", file)
  my_data <- rbind(my_data, temp)
}

str(my_data)

# Build the final plot, from scratch!
ggplot(my_data, aes(x = yearday, y = temp, year = year)) +
  stat_historical() +
  stat_present() +
  stat_extremes(aes(col = ..record..)) +
  scale_color_identity() +  # specify colour here
  facet_wrap(~ id, ncol = 2)

```
