---
title: "Data Visualization with ggplot2 (Part 3-A)"
author: "William Surles"
date: "2017-08-02"
output: 
 html_document:
  self_contained: yes
  theme: flatly
  highlight: tango
  toc: true
  toc_float: true
  toc_depth: 3
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(eval=T, echo=T, cache=T, message=F, warning=F)

```

# Introduction

  - Course notes from the [Data Vizualization with ggplot2 (Part 3)](https://www.datacamp.com/courses/data-visualization-with-ggplot2-part-3) course on DataCamp
  
## Whats Covered

**Part A:**

  - Statistical plots
    - Aesthetics review, 
    - box plots, density plots 
    - multiple groups/variables
  - Plots for specific data types (Part 1)
    - graphics of large data
    - Ternary plots
    - Network plots
    - Diagnostic plots

Part B:

  - Plots for specific data types (Part 2)
    - choropleths
    - cartographic maps
    - animations
    
Part C:

  - ggplot2 internals
    - grid graphics, grid grapshics in ggplot2
    - ggplot objects
    - gridExtra
  - Data Munging and Visualization Case Study
    - Bag plot case study, weather case study

## Libraries and Data

```{r} 

source("create_datasets.R")
load('data/test_datasets.RData')

library(readr)
library(dplyr)
library(ggplot2)
library(purrr)

library(ggplot2movies)
library(viridis)
library(GGally)
library(ggtern)
library(ggthemes)
library(geomnet)
library(ggmap)
library(ggfortify)

```

&nbsp; &nbsp;

***  
# Statistical plots
***  

## Introduction

### -- Refresher (1)

```{r}

# Create movies_small
# library(ggplot2movies)
set.seed(123)
movies_small <- movies[sample(nrow(movies), 1000), ]
movies_small$rating <- factor(round(movies_small$rating))

# Explore movies_small with str()
str(movies_small)

# Build a scatter plot with mean and 95% CI
ggplot(movies_small, aes(x = rating, y = votes)) +
  geom_point() +
  stat_summary(fun.data = "mean_cl_normal",
               geom = "crossbar",
               width = 0.2,
               col = "red") +
  scale_y_log10()

```

### -- Refresher (2)

```{r}

str(diamonds)

# Reproduce the plot
ggplot(diamonds, aes(x = carat, y = price, col = color)) +
  geom_point(alpha = 0.5, size = 0.5, shape = 16) +
  scale_x_log10(expression(log[10](Carat)), limits = c(0.1, 10)) +
  scale_y_log10(expression(log[10](Price)), limits = c(100, 100000)) +
  scale_color_brewer(palette = "YlOrRd") +
  coord_equal() +
  theme_classic()
  
```

### -- Refresher (3)

```{r}

# Add smooth layer and facet the plot
ggplot(diamonds, aes(x = carat, y = price, col = color)) +
  stat_smooth(method = "lm") + 
  scale_x_log10(expression(log[10](Carat)), limits = c(0.1,10)) +
  scale_y_log10(expression(log[10](Price)), limits = c(100,100000)) +
  scale_color_brewer(palette = "YlOrRd") +
  coord_equal() +
  theme_classic()

```

## Box Plots

### -- Transformations

```{r}

# movies_small is available

# Add a boxplot geom
d <- ggplot(movies_small, aes(x = rating, y = votes)) +
  geom_point() +
  geom_boxplot() +
  stat_summary(fun.data = "mean_cl_normal",
               geom = "crossbar",
               width = 0.2,
               col = "red")

# Untransformed plot
d

# Transform the scale
d + scale_y_log10()

# Transform the coordinates
## coord_trans is different to scale transformations in that it occurs after statistical transformation and will affect the visual appearance of geoms - there is no guarantee that straight lines will continue to be straight.

## This does not work in my case. 
## It likely is from the statitics having a zero value.
## d + coord_trans(y = "log10")

# It works fine without the stats layer
ggplot(movies_small, aes(x = rating, y = votes)) +
  geom_point() +
  geom_boxplot() +
  coord_trans(y = "log10")

```

This is the example from the documentation which actually works

```{r}
# Three ways of doing transformation in ggplot:

#  * by transforming the data
ggplot(diamonds, aes(log10(carat), log10(price))) +
  geom_point()

#  * by transforming the scales
ggplot(diamonds, aes(carat, price)) +
  geom_point() +
  scale_x_log10() +
  scale_y_log10()

#  * by transforming the coordinate system:
ggplot(diamonds, aes(carat, price)) +
  geom_point() +
  coord_trans(x = "log10", y = "log10")

# The difference between transforming the scales and
# transforming the coordinate system is that scale
# transformation occurs BEFORE statistics, and coordinate
# transformation afterwards.  Coordinate transformation also
# changes the shape of geoms:

d <- subset(diamonds, carat > 0.5)

ggplot(d, aes(carat, price)) +
  geom_point() +
  geom_smooth(method = "lm") +
  scale_x_log10() +
  scale_y_log10()

ggplot(d, aes(carat, price)) +
  geom_point() +
  geom_smooth(method = "lm") +
  coord_trans(x = "log10", y = "log10")

```

### -- Cut it up!

```{r}

# Plot object p
p <- ggplot(diamonds, aes(x = carat, y = price))

# Use cut_interval
p + geom_boxplot(aes(group = cut_interval(carat, n = 10)))

# Use cut_number
p + geom_boxplot(aes(group = cut_number(carat, n = 10)))

# Use cut_width
p + geom_boxplot(aes(group = cut_width(carat, width = 0.25)))

```

  - If you only have continuous variables, you can convert them into ordinal variables using any of the following functions:
    - `cut_interval(x, n)` makes n groups from vector x with equal range.
    - `cut_number(x, n)` makes n groups from vector x with (approximately) equal numbers of observations.
    - `cut_width(x, width)` makes groups of width width from vector x.

### -- Understanding quartiles

  - Notice how the IQR becomes more consistent across methods as the sample size increases

```{r}
plot_quart <- function(n) {
  set.seed(123)
  playData <- data.frame(raw.values = rnorm(n, 1, 6))

  quan.summary <- data.frame(t(sapply(1:9, function(x) quantile(playData$raw.values, type = x))))
  names(quan.summary) <- c("Min", "Q1", "Median", "Q3", "Max")
  quan.summary$Type <- as.factor(1:9)

  library(reshape2)
  quan.summary <- melt(quan.summary, id = "Type")
  quan.summary <- list(quartiles = quan.summary, values = playData)

  ggplot(quan.summary$quartiles, aes(x = Type, y = value, col = variable)) +
    geom_point() +
    geom_rug(data = quan.summary$values, aes(y = raw.values), sides = "l", inherit.aes = F)
}

plot_quart(4)
plot_quart(10)
plot_quart(50)
plot_quart(100)
```
  
## Density Plots

### -- geom_density()

```{r}

# test_datasets.RData has been loaded

str(ch1_test_data)

# Calculating density: d
d <- density(ch1_test_data$norm, bw = "nrd0", kernel = "gaussian")

# Use which.max() to calculate mode
mode <- d$x[which.max(d$y)]

# Finish the ggplot call
ggplot(ch1_test_data, aes(x = norm)) +
  geom_density() +
  geom_rug() +
  geom_vline(xintercept = mode, col = "red")

```

### -- Combine density plots and histogram

```{r}

# ch1_test_data is available

# Arguments you'll need later on
fun_args <- list(mean = mean(ch1_test_data$norm), sd = sd(ch1_test_data$norm))

# Finish the ggplot
ggplot(ch1_test_data, aes(x = norm)) + 
  geom_histogram(aes(y = ..density..)) + 
  geom_density(col = "red") + 
  stat_function(
    fun = dnorm, 
    args = fun_args, 
    col = "blue")

```

### -- Adjusting density plots

There are three parameters that you may be tempted to adjust in a density plot:

  - bw - the smoothing bandwidth to be used, see ?density for details
  - adjust - adjustment of the bandwidth, see density for details
  - kernel - kernel used for density estimation, defined as
    - "g" = gaussian
    - "r" = rectangular
    - "t" = triangular
    - "e" = epanechnikov
    - "b" = biweight
    - "c" = cosine
    - "o" = optcosine

```{r}

# small_data is available
small_data <- structure(list(x = c(-3.5, 0, 0.5, 6)), .Names = "x", row.names = c(NA, 
-4L), class = "data.frame")

# Get the bandwith
get_bw <- density(small_data$x)$bw

# Basic plotting object
p <- ggplot(small_data, aes(x = x)) +
  geom_rug() +
  coord_cartesian(ylim = c(0,0.5))

# Create three plots
p + geom_density()
p + geom_density(adjust = 0.25)
p + geom_density(bw = 0.25 * get_bw)

# Create two plots
## rectangular kernel
p + geom_density(kernel = "r")
## epanechnikov kernel
p + geom_density(kernel = "e")

```

## Multiple Groups/Variables

### -- Box plots with varying width

  - One way to represent the sample size, *n*, is to use variable witdths for the boxes. 

```{r, fig.width=9}

# Finish the plot
ggplot(diamonds, aes(x = cut, y = price, col = color)) + 
  geom_boxplot(varwidth = T) + 
  facet_grid(. ~ color)

```

### -- Mulitple density plots

```{r}

# ch1_test_data and ch1_test_data2 are available
str(ch1_test_data)
str(ch1_test_data2)

# Plot with ch1_test_data
ggplot(ch1_test_data, aes(x = norm)) +
  geom_rug() + 
  geom_density()

# Plot two distributions with ch1_test_data2
ggplot(ch1_test_data2, aes(x = value, fill = dist, col = dist)) +
  geom_rug(alpha = 0.6) + 
  geom_density(alpha = 0.6)

```

### -- Multiple density plots (2)

```{r}

# Individual densities
ggplot(mammals[mammals$vore == "Insectivore", ], 
    aes(x = sleep_total, fill = vore)) +
  geom_density(col = NA, alpha = 0.35) +
  scale_x_continuous(limits = c(0, 24)) +
  coord_cartesian(ylim = c(0, 0.3))

# With faceting
ggplot(mammals, aes(x = sleep_total, fill = vore)) +
  geom_density(col = NA, alpha = 0.35) +
  scale_x_continuous(limits = c(0, 24)) +
  coord_cartesian(ylim = c(0, 0.3)) +
  facet_wrap( ~ vore, nrow = 2)

# Note that by default, the x ranges fill the scale
ggplot(mammals, aes(x = sleep_total, fill = vore)) +
  geom_density(col = NA, alpha = 0.35) +
  scale_x_continuous(limits = c(0, 24)) +
  coord_cartesian(ylim = c(0, 0.3))

# Trim each density plot individually
ggplot(mammals, aes(x = sleep_total, fill = vore)) +
  geom_density(col = NA, alpha = 0.35, trim = T) +
  scale_x_continuous(limits=c(0,24)) +
  coord_cartesian(ylim = c(0, 0.3))

```

### -- Weighted density plots

  - When plotting a single variable, the density plots (and their bandwidths) are calculated separate for each variable (see the plot from the previous exercise, provided).
  - However, when you compare several variables (such as eating habits) it's useful to see the density of each subset in relation to the whole data set. 

```{r}

# Unweighted density plot from before
ggplot(mammals, aes(x = sleep_total, fill = vore)) +
  geom_density(col = NA, alpha = 0.35) +
  scale_x_continuous(limits = c(0, 24)) +
  coord_cartesian(ylim = c(0, 0.3))

# Unweighted violin plot
ggplot(mammals, aes(x = vore, y = sleep_total, fill = vore)) +
  geom_violin()

# Calculate weighting measure
library(dplyr)
mammals2 <- mammals %>%
  group_by(vore) %>%
  mutate(n = n() / nrow(mammals)) -> mammals

str(mammals2, give.attr = F)
str(mammals, give.attr = F)

# Weighted density plot
## I remove the ylim because the y scale changes here
ggplot(mammals, aes(x = sleep_total, fill = vore)) +
  geom_density(aes(weight = n), col = NA, alpha = 0.35) +
  scale_x_continuous(limits = c(0, 24))

# Weighted violin plot
ggplot(mammals, aes(x = vore, y = sleep_total, fill = vore)) +
  geom_violin(aes(weight = n), col = NA)

```

### -- 2D density plots (1)

```{r}

# Base layers
p <- ggplot(faithful, aes(x = waiting, y = eruptions)) +
  scale_y_continuous(limits = c(1, 5.5), expand = c(0, 0)) +
  scale_x_continuous(limits = c(40, 100), expand = c(0, 0)) +
  coord_fixed(60 / 4.5)

# 1 - Use geom_density_2d()
p + geom_density_2d()

# 2 - Use stat_density_2d() with arguments
p + stat_density_2d(aes(col = ..level..), h = c(5, 0.5))

```

### -- 2D density plots (2)

```{r}

# Load in the viridis package
library(viridis)

# Add viridis color scale
ggplot(faithful, aes(x = waiting, y = eruptions)) +
  scale_y_continuous(limits = c(1, 5.5), expand = c(0,0)) +
  scale_x_continuous(limits = c(40, 100), expand = c(0,0)) +
  coord_fixed(60/4.5) +
  stat_density_2d(geom = "tile", aes(fill = ..density..), h=c(5,.5), contour = FALSE) +
  scale_fill_viridis()

```


&nbsp; &nbsp;

***  
# Plots for specific data types (Part 1)
***  

## Graphics of Large Data

### -- Pair plots and correlation matrices

```{r}

# pairs
pairs(iris[1:4])

# chart.Correlation
library(PerformanceAnalytics)

chart.Correlation(iris[1:4])

# ggpairs
# library(GGally)

mtcars_fact <- mtcars %>%
  mutate(
    cyl = as.factor(cyl),
    vs = as.factor(vs),
    am = as.factor(am),
    gear = as.factor(gear),
    carb = as.factor(carb)
  )

ggpairs(mtcars_fact[1:3])

```

### -- Create a correlation matrix in ggplot2

```{r}

library(ggplot2)
library(reshape2)

cor_list <- function(x) {
  L <- M <- cor(x)
  
  M[lower.tri(M, diag = TRUE)] <- NA
  M <- melt(M)
  names(M)[3] <- "points"
  
  L[upper.tri(L, diag = TRUE)] <- NA
  L <- melt(L)
  names(L)[3] <- "labels"
  
  merge(M, L)
}

# Calculate xx with cor_list
library(dplyr)

xx <- iris %>%
  group_by(Species) %>%
  do(cor_list(.[1:4])) 

# Finish the plot
ggplot(xx, aes(x = Var1, y = Var2)) +
  geom_point(
    aes(col = points, size = abs(points)), 
    shape = 16
    ) +
  geom_text(
    aes(col = labels, size = abs(labels), label = round(labels, 2))
    ) +
  scale_size(range = c(0, 6)) +
  scale_color_gradient2("r", limits = c(-1, 1)) +
  scale_y_discrete("", limits = rev(levels(xx$Var1))) +
  scale_x_discrete("") +
  guides(size = FALSE) +
  geom_abline(slope = -1, intercept = nlevels(xx$Var1) + 1) +
  coord_fixed() +
  facet_grid(. ~ Species) +
  theme(axis.text.y = element_text(angle = 45, hjust = 1),
        axis.text.x = element_text(angle = 45, hjust = 1),
        strip.background = element_blank())

```

## Ternary Plots

### -- Proportional/stacked bar plots

```{r}

# Explore africa
load('data/africa.RData')
africa_sample <- sample_n(africa, 50)
str(africa_sample)
head(africa_sample)

# Add an ID column from the row.names
africa_sample$ID <- row.names(africa_sample)

# Gather africa_sample
library(tidyr)
africa_sample_tidy <- gather(africa_sample, key, value, -ID)
head(africa_sample_tidy)

# Finish the ggplot command
ggplot(africa_sample_tidy, aes(x = factor(ID), y = value, fill = key)) +
  geom_col() +
  coord_flip()

```

### -- Producing ternary plots

```{r}

# The ggtern library is loaded

# Build ternary plot
str(africa)
ggtern(africa, aes(x = Sand, y = Silt, z = Clay)) +
  geom_point(shape = 16, alpha = 0.2)

```

### -- Adjusting ternary plots

```{r}

# ggtern and ggplot2 are loaded

# Plot 1
ggtern(africa, aes(x = Sand, y = Silt, z = Clay)) +
  geom_density_tern()

# Plot 2
ggtern(africa, aes(x = Sand, y = Silt, z = Clay)) +
  stat_density_tern(geom = 'polygon', aes(fill = ..level.., alpha = ..level..)) +
  guides(fill = FALSE)

```

  - Just playing around and trying to make my own pretty plot
    - I want something like the faithful density but on the tern
    - This is a start
    
```{r}

## I want to see all the points on there
ggtern(africa, aes(x = Sand, y = Silt, z = Clay)) +
  geom_point(alpha = 0.1, color = "navyblue", size = .5) + 
  stat_density_tern(
    geom = 'polygon', 
    aes(fill = ..level.., alpha = ..level..),
    bins = 100
    ) +
  guides(alpha = FALSE) + 
  scale_fill_viridis()

```

## Network Plots

### -- Build the network (1)

```{r}

# Load geomnet & examine structure of madmen
# The geomnet library is loaded
str(madmen)

## This is a much better way to see whats in each list. Love it. 
## library(purrr)
madmen %>% purrr::map(head)

# Merge edges and vertices
mmnet <- merge(madmen$edges, madmen$vertices,
               by.x = "Name1", by.y = "label",
               all = TRUE)

# Examine structure of mmnet
head(mmnet)
str(mmnet)

```

### -- Build the network (2)

```{r}

# geomnet is pre-loaded

# Merge edges and vertices
mmnet <- merge(madmen$edges, madmen$vertices,
               by.x = "Name1", by.y = "label",
               all = TRUE)

head(mmnet)

# Finish the ggplot command
ggplot(data = mmnet, aes(from_id = Name1, to_id = Name2)) +
  geom_net(
    aes(col = Gender),
    size = 6,
    linewidth = 1,
    labelon = T,
    fontsize = 3,
    labelcolour = "black")

```

### -- Adjusting the network

```{r}

# geomnet is pre-loaded
# ggmap is already loaded

# Merge edges and vertices
mmnet <- merge(madmen$edges, madmen$vertices,
               by.x = "Name1", by.y = "label",
               all = TRUE)

# Tweak the network plot
ggplot(data = mmnet, aes(from_id = Name1, to_id = Name2)) +
  geom_net(
    aes(col = Gender),
    size = 6,
    linewidth = 1,
    labelon = TRUE,
    fontsize = 3,
    labelcolour = "black",
    directed = T) +
  scale_color_manual(values = c("#FF69B4", "#0099ff")) +
  xlim(c(-0.05, 1.05)) +
  theme_nothing() +
  theme(legend.key = element_blank())

```

## Diagnostic Plots

### -- Autoplot on linear models

```{r, fig.width = 9}

# Create linear model: res
res <- lm(Volume ~ Girth, data = trees)

# Plot res
plot(res)

# Import ggfortify and use autoplot()
# library(ggfortify)
autoplot(res, ncol = 2)

```

### -- ggfortify - time series

```{r}

# ggfortify and Canada are available

# Inspect structure of Canada
str(Canada)
head(Canada)

# Call plot() on Canada
plot(Canada)

# Call autoplot() on Canada
# autoplot is from the ggfortify library
autoplot(Canada)

```

### -- Distance matrices and Multi-Dimensional Scaling (MDS)

  - The cmdscale() function from the stats package performs Classical Multi-Dimensional Scaling and returns point coodinates as a matrix. 
  - Although autoplot() will work on this object, it will produce a heatmap, and not a scatter plot. 
  - However, if either eig = TRUE, add = TRUE or x.ret = TRUE is specified, cmdscale() will return a list instead of matrix. 
  - In these cases, the list method for autoplot() in the ggfortify package can deal with the output. 
  - Specifics on multi-dimensional scaling is beyond the scope of this course, however details on the method and these arguments can be found in the help pages `?cmdscale`.

```{r}

# ggfortify and eurodist are available
str(eurodist)

# Autoplot + ggplot2 tweaking
autoplot(eurodist) + 
  coord_fixed()

# Autoplot of MDS
autoplot(cmdscale(eurodist, eig = TRUE), 
         label = TRUE, 
         label.size = 3, 
         size = 0)

```

### -- Plotting K-means clustering

  - You must explicitly pass the original data to the autoplot function via the data argument, since kmeans objects don't contain the original data.
  - This kmeans cluster is wrong. Its very differnt from what I saw in the datacamp exercise. Not sure why.

```{r}

# Perform clustering
iris_k <- kmeans(iris[-5], 3)

# Autoplot: color according to cluster
autoplot(iris_k, data = iris, frame = T)

# Autoplot: above, plus shape according to species
autoplot(iris_k, data = iris, frame = T, shape = 'Species')

```

